### 2018-06-25 phrases.Rmd
references added, R code some, paper template confirmed

---
title: "Some usefull words/expressions/sentences"
output: html_document
---

# Words

- discernable [di'sə:nəbl, -'zə:-]

```
===== 辞典翻译: discernable ======
adj. 可辨别的；可认识的
============ 网络释义 ============
-------- discernable ---------
可辨别的
方向
分辨
-- discernable recognizable --
可辨别的
--- discernable visible ----
可辨别的
```

-   abstinence ['æbstinəns]

```
====== 辞典翻译: abstinence ======
  n. 节制；节欲；戒酒；禁食
============ 网络释义 ============
--------- abstinence ---------
  节制
  禁欲
  禁戒
----- alcohol abstinence -----
  酒戒断
----- Abstinence theory ------
  节欲论
  弃权
  忍欲说
```

-  exhaustive [iɡ'zɔ:stiv]

```
====== 辞典翻译: exhaustive ======
  adj. 详尽的；彻底的；消耗的
============ 网络释义 ============
--------- Exhaustive ---------
  无遗
  详尽的
  全部
----- Exhaustive search ------
  穷举搜索
  穷举搜索完全搜索
  全程搜索
----- exhaustive voting ------
  淘汰投票
  消耗性投票
```

-   temperament ['tempərəmənt]

```
===== 辞典翻译: temperament ======
  n. 气质，性情，性格；急躁
============ 网络释义 ============
-------- temperament ---------
  气质
  气质 (心理学)
  性格
------ Well Temperament ------
  Well temperament
  Well Temperament
  平均律
---- Musical temperament -----
  律学
  一份音乐的气质
  音乐性
```

# Expressions

# Sentences

#### terry2017discontinuous - Discontinuous Patterns of Cigarette Smoking From Ages 18 to 50 in the United States: A Repeated-Measures Latent Class Analysis

- RMLCA models were fitted in SAS 9.4 using PROC LCA. Parameters were estimated by maximum likelihood using the EM algorithm.

- To ascertain if the same latent class structure was observed for males and females, multiple-group RMLCA models by sex were fitted in PROC LCA using the GROUPS statement (both with and without imposing measurement invariance across males and females).

- Model selection (ie, the number of latent classes specified) was determined by model fit, parsimony, and stability.

- Simulations have shown that the Bayesian information criterion (BIC) and sample size-adjusted BIC (a-BIC) perform particularly well at selecting the “correct” latent class model.

- Improvement in both BIC and consistent Akaike information criterion (CAIC) values continued only through the 12-class model; thus, the 12-class model was selected as optimal.

#### lanza2007proc - PROC LCA: A SAS procedure for latent class analysis

- In traditional LCA, two sets of parameters are estimated: class membership probabilities and item-response probabilities conditional on class membership.

- Latent class models usually involve categorical indicators (although a version of LCA involving continuous indicators called latent profile analysis [Gibson, 1959] is being used increasingly).

- When categorical data are used, the latent class model has the advantage of making no assumptions about the distributions of the indicators other than that of local independence; that is, the assumption that within a latent class the indicators are independent.