### 2018-07-05 9thJuly/9thJuly.Rmd
LCA analysis by day

---
title: "Preliminary results"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

# Data description

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
setwd("~/Documents/LSHTMproject/Rcode")

load("LCA.Rdata")

library(epiDisplay)
library(plyr)
library(tidyverse)
library(poLCA)
```

## Number of subjects per survey year

```{r message=FALSE, warning=FALSE, paged.print=FALSE}

with(NDNS, tab1(SurveyYear, graph = FALSE, decimal = 2))

```

## Number of subjects per servey year by gender where Men == 1, Women == 2

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
with(NDNS, tabpct(SurveyYear, Sex,  graph = FALSE, decimal = 2))
```

## Summary of their age 
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
NDNS %>% 
  group_by(SurveyYear, Sex) %>% 
  summarise(N = n(), MeanAge = mean(Age), SDAge = sd(Age), minAge = min(Age), maxAge = max(Age))
```

## Distribution of the dietary data (by Day of Week)

### Day1

```{r}
rm(list=ls(all=TRUE))
load("~/Documents/LSHTMproject/Rcode/NDNSday1_4.Rdata")
with(dta_d1_wide, tab1(DayofWeek, graph = FALSE, decimal = 2))
```

### Day2

```{r}
with(dta_d2_wide, tab1(DayofWeek, graph = FALSE, decimal = 2))
```

### Day3

```{r}
with(dta_d3_wide, tab1(DayofWeek, graph = FALSE, decimal = 2))
```

### Day4

```{r}
with(dta_d4_wide, tab1(DayofWeek, graph = FALSE, decimal = 2))
```

## Distribution of the dietary data (by DayNo)

### Monday

```{r}
rm(list=ls(all=TRUE))
load("~/Documents/LSHTMproject/Rcode/NDNSMon_Sun.Rdata")
with(dta_Mon_wide, tab1(DayNo, graph = FALSE, decimal = 2))
```

### Tuesday

```{r}
with(dta_Tue_wide, tab1(DayNo, graph = FALSE, decimal = 2))
```

### Wednesday

```{r}
with(dta_Wed_wide, tab1(DayNo, graph = FALSE, decimal = 2))
```

### Thursday

```{r}
with(dta_Thurs_wide, tab1(DayNo, graph = FALSE, decimal = 2))
```

### Friday

```{r}
with(dta_Fri_wide, tab1(DayNo, graph = FALSE, decimal = 2))
```

### Saturday

```{r}
with(dta_Sat_wide, tab1(DayNo, graph = FALSE, decimal = 2))
```

### Sunday

```{r}
with(dta_Sun_wide, tab1(DayNo, graph = FALSE, decimal = 2))
```

**The problem of analysing data by day of the week would be that every subject only contributed 3-4 days' data. Therefore, we cannot see one subject's dietary change during a Mon-Sun week.**

## Dietary data by day

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
rm(list=ls(all=TRUE))
load("~/Documents/LSHTMproject/Rcode/NDNSday1_4.Rdata")

vecid <- unique(dfs3$id)

vecid1<-unique(dta_day1$id) # n = 6153
vecid2<-unique(dta_day2$id) # n = 6153
vecid3<-unique(dta_day3$id) # n = 6151
vecid4<-unique(dta_day4$id) # n = 6026

setdiff(vecid, vecid1) # two subjects did not have day 1 data

setdiff(vecid, vecid2) # two subjects did not have day 2 data

setdiff(vecid, vecid3) # four subjects did not have day 3 data

setdiff(vecid, vecid4) # 129 subjects did not have day 4 data
```

# LCA analyses by day

## Day 1

```{r eval=FALSE}
set.seed(01012)
max_II <- 100000
lc1<-poLCA(f, data=dta_d1_wide, nclass=1, na.rm = FALSE, nrep=20, maxiter=max_II) #Loglinear independence model.
lc2<-poLCA(f, data=dta_d1_wide, nclass=2, na.rm = FALSE, nrep=20, maxiter=max_II)
lc3<-poLCA(f, data=dta_d1_wide, nclass=3, na.rm = FALSE, nrep=20, maxiter=max_II)
lc4<-poLCA(f, data=dta_d1_wide, nclass=4, na.rm = FALSE, nrep=20, maxiter=max_II) 
lc5<-poLCA(f, data=dta_d1_wide, nclass=5, na.rm = FALSE, nrep=20, maxiter=max_II)
lc6<-poLCA(f, data=dta_d1_wide, nclass=6, na.rm = FALSE, nrep=20, maxiter=max_II)
lc7<-poLCA(f, data=dta_d1_wide, nclass=7, na.rm = FALSE, nrep=20, maxiter=max_II)
lc8<-poLCA(f, data=dta_d1_wide, nclass=8, na.rm = FALSE, nrep=20, maxiter=max_II)
```

### Model comparison and selection

```{r echo=FALSE}
rm(list=ls(all=TRUE))
load("~/Documents/LSHTMproject/Rcode/day1LCA1_6.Rdata")

library(kableExtra)
# dt <- read.csv("../Tablecsv/Modelfitd01.csv", header = T)
# names(dt) <- c("N of classes", "N of parameters", "G^2", "df", "AIC", "BIC", "log-likelihood")
lca_results[1,9] <- "---"
kable(lca_results, digits = 3, row.names = FALSE, align = "c",
              format = "html", caption = "Model Comparison. (Day 1, n = 6153)") %>%
 kable_styling(font_size = 16, 
        bootstrap_options = c("striped", "hover", "condensed"),
        position = "center", full_width = FALSE) %>% 
  row_spec(c(4,6), bold = T, color = "white", background = "#D7261E") %>% 
  footnote(general = "Abbreviation: N, number; resid.df, residual degree of freedom; AIC, Akaike information criterion; BIC, Bayesian information criterion; aBIC, adjusted BIC; cAIC, consistent AIC; Entropy, a pseudo-r-squared index.")
```

### Day 1 data (Fig. 4 classes)

```{r echo=FALSE, fig.height=5, fig.width=14, message=FALSE, warning=FALSE}

library(ggthemr)

# lc5 <- poLCA(f, dta_d1_wide, nclass = 5, graphs = FALSE, maxiter = 10000)

# a <- (data.frame(lc5$probs))
a <- (data.frame(lc4$probs))
# a <- (data.frame(lc6$probs))

# a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5")
a$Class <- c("Class1", "Class2", "Class3", "Class4")

# a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5", "Class6")

a_long <- a %>% 
  gather(Hour, Prob, -Class) %>% 
  separate(Hour, into = c("HourN", "Carbo"), sep = "\\.")

a_long$HourN <- factor(a_long$HourN, levels = c("H0","H1" ,"H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10","H11","H12","H13","H14","H15","H16","H17","H18","H19","H20","H21","H22","H23"))

a_long$Carbo <- factor(a_long$Carbo, levels = c("Low_carb", "Med_carb", "High_carb"))

a_long <- a_long %>% 
  filter(Carbo != "Not_eating")

ggthemr("fresh", layout = "scientific")

ggplot(a_long[a_long$Class == "Class1", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
  geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 1: 38.80%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class2", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 2:  28.25%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class3", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 3:  10.81%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class4", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 4:  23.14%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

# 
# ggplot(a_long[a_long$Class == "Class5", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
#  geom_point(size = 3) +
#   geom_line(size = 1.5) +
#   theme(axis.title = element_text(size = 18),
#         axis.text = element_text(size = 18),
#         axis.line = element_line(colour = "black"),
#         panel.border = element_blank(),
#         panel.background = element_blank(),
#         legend.text = element_text(size = 15),
#         legend.title = element_text(size = 15)
#         # legend.position = "bottom",
#         # legend.direction = "horizontal"
#         ) +
#   labs(title = "Class 5:  18.52%", x = "Hour of the day", y = "Probability",
#        color = "Carbohydrate\nintake") +
#   scale_color_discrete(labels = c("Low", "Medium", "High")) +
#   ylim(c(0,1))
# 
# ggplot(a_long[a_long$Class == "Class6", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
#  geom_point(size = 3) +
#   geom_line(size = 1.5) +
#   theme(axis.title = element_text(size = 18),
#         axis.text = element_text(size = 18),
#         axis.line = element_line(colour = "black"),
#         panel.border = element_blank(),
#         panel.background = element_blank(),
#         legend.text = element_text(size = 15),
#         legend.title = element_text(size = 15)
#         # legend.position = "bottom",
#         # legend.direction = "horizontal"
#         ) +
#   labs(title = "Class 6:  10.02%", x = "Hour of the day", y = "Probability",
#        color = "Carbohydrate\nintake") +
#   scale_color_discrete(labels = c("Low", "Medium", "High")) +
#   ylim(c(0,1))
```

### Day 1 data (Fig. 6 classes)

```{r echo=FALSE, fig.height=5, fig.width=14, message=FALSE, warning=FALSE}

library(ggthemr)

# lc5 <- poLCA(f, dta_d1_wide, nclass = 5, graphs = FALSE, maxiter = 10000)

# a <- (data.frame(lc5$probs))
# a <- (data.frame(lc4$probs))
a <- (data.frame(lc6$probs))

# a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5")
# a$Class <- c("Class1", "Class2", "Class3", "Class4")

a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5", "Class6")

a_long <- a %>% 
  gather(Hour, Prob, -Class) %>% 
  separate(Hour, into = c("HourN", "Carbo"), sep = "\\.")

a_long$HourN <- factor(a_long$HourN, levels = c("H0","H1" ,"H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10","H11","H12","H13","H14","H15","H16","H17","H18","H19","H20","H21","H22","H23"))

a_long$Carbo <- factor(a_long$Carbo, levels = c("Low_carb", "Med_carb", "High_carb"))

a_long <- a_long %>% 
  filter(Carbo != "Not_eating")

ggthemr("fresh", layout = "scientific")

ggplot(a_long[a_long$Class == "Class1", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
  geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 1: 26.79%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class2", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 2:  11.72%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class3", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 3:  7.37%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class4", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 4:  25.58%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class5", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 5:  18.52%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class6", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 6:  10.02%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))
```

## Day 2

```{r eval=FALSE}
set.seed(01012)
max_II <- 100000
lc1<-poLCA(f, data=dta_d1_wide, nclass=1, na.rm = FALSE, nrep=20, maxiter=max_II) #Loglinear independence model.
lc2<-poLCA(f, data=dta_d1_wide, nclass=2, na.rm = FALSE, nrep=20, maxiter=max_II)
lc3<-poLCA(f, data=dta_d1_wide, nclass=3, na.rm = FALSE, nrep=20, maxiter=max_II)
lc4<-poLCA(f, data=dta_d1_wide, nclass=4, na.rm = FALSE, nrep=20, maxiter=max_II) 
lc5<-poLCA(f, data=dta_d1_wide, nclass=5, na.rm = FALSE, nrep=20, maxiter=max_II)
lc6<-poLCA(f, data=dta_d1_wide, nclass=6, na.rm = FALSE, nrep=20, maxiter=max_II)
lc7<-poLCA(f, data=dta_d1_wide, nclass=7, na.rm = FALSE, nrep=20, maxiter=max_II)
lc8<-poLCA(f, data=dta_d1_wide, nclass=8, na.rm = FALSE, nrep=20, maxiter=max_II)

# running time:  31.60514 mins
```

### Model comparison and selection

```{r echo=FALSE}
rm(list=ls(all=TRUE))
load("~/Documents/LSHTMproject/Rcode/day2LCA1_6.Rdata")

library(kableExtra)
# dt <- read.csv("../Tablecsv/Modelfitd01.csv", header = T)
# names(dt) <- c("N of classes", "N of parameters", "G^2", "df", "AIC", "BIC", "log-likelihood")
lca_results[1,9] <- "---"
kable(lca_results, digits = 3, row.names = FALSE, align = "c",
              format = "html", caption = "Model Comparison. (Day 2, n = 6153)") %>%
 kable_styling(font_size = 16, 
        bootstrap_options = c("striped", "hover", "condensed"),
        position = "center", full_width = FALSE) %>% 
  row_spec(c(4,6), bold = T, color = "white", background = "#D7261E") %>% 
  footnote(general = "Abbreviation: N, number; resid.df, residual degree of freedom; AIC, Akaike information criterion; BIC, Bayesian information criterion; aBIC, adjusted BIC; cAIC, consistent AIC; Entropy, a pseudo-r-squared index.")
```

### Day 2 data (Fig. 4 classes)

```{r echo=FALSE, fig.height=5, fig.width=14, message=FALSE, warning=FALSE}

library(ggthemr)

# lc5 <- poLCA(f, dta_d1_wide, nclass = 5, graphs = FALSE, maxiter = 10000)

# a <- (data.frame(lc5$probs))
a <- (data.frame(lc4$probs))
# a <- (data.frame(lc6$probs))

# a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5")
a$Class <- c("Class1", "Class2", "Class3", "Class4")

# a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5", "Class6")

a_long <- a %>% 
  gather(Hour, Prob, -Class) %>% 
  separate(Hour, into = c("HourN", "Carbo"), sep = "\\.")

a_long$HourN <- factor(a_long$HourN, levels = c("H0","H1" ,"H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10","H11","H12","H13","H14","H15","H16","H17","H18","H19","H20","H21","H22","H23"))

a_long$Carbo <- factor(a_long$Carbo, levels = c("Low_carb", "Med_carb", "High_carb"))

a_long <- a_long %>% 
  filter(Carbo != "Not_eating")

ggthemr("fresh", layout = "scientific")

ggplot(a_long[a_long$Class == "Class1", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
  geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 1: 9.77%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class2", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 2:  24.68%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class3", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 3:  38.28%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class4", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 4:  27.27%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

# 
# ggplot(a_long[a_long$Class == "Class5", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
#  geom_point(size = 3) +
#   geom_line(size = 1.5) +
#   theme(axis.title = element_text(size = 18),
#         axis.text = element_text(size = 18),
#         axis.line = element_line(colour = "black"),
#         panel.border = element_blank(),
#         panel.background = element_blank(),
#         legend.text = element_text(size = 15),
#         legend.title = element_text(size = 15)
#         # legend.position = "bottom",
#         # legend.direction = "horizontal"
#         ) +
#   labs(title = "Class 5:  18.52%", x = "Hour of the day", y = "Probability",
#        color = "Carbohydrate\nintake") +
#   scale_color_discrete(labels = c("Low", "Medium", "High")) +
#   ylim(c(0,1))
# 
# ggplot(a_long[a_long$Class == "Class6", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
#  geom_point(size = 3) +
#   geom_line(size = 1.5) +
#   theme(axis.title = element_text(size = 18),
#         axis.text = element_text(size = 18),
#         axis.line = element_line(colour = "black"),
#         panel.border = element_blank(),
#         panel.background = element_blank(),
#         legend.text = element_text(size = 15),
#         legend.title = element_text(size = 15)
#         # legend.position = "bottom",
#         # legend.direction = "horizontal"
#         ) +
#   labs(title = "Class 6:  10.02%", x = "Hour of the day", y = "Probability",
#        color = "Carbohydrate\nintake") +
#   scale_color_discrete(labels = c("Low", "Medium", "High")) +
#   ylim(c(0,1))
```

### Day 2 data (Fig. 6 classes)

```{r echo=FALSE, fig.height=5, fig.width=14, message=FALSE, warning=FALSE}

library(ggthemr)

# lc5 <- poLCA(f, dta_d1_wide, nclass = 5, graphs = FALSE, maxiter = 10000)

# a <- (data.frame(lc5$probs))
# a <- (data.frame(lc4$probs))
a <- (data.frame(lc6$probs))

# a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5")
# a$Class <- c("Class1", "Class2", "Class3", "Class4")

a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5", "Class6")

a_long <- a %>% 
  gather(Hour, Prob, -Class) %>% 
  separate(Hour, into = c("HourN", "Carbo"), sep = "\\.")

a_long$HourN <- factor(a_long$HourN, levels = c("H0","H1" ,"H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10","H11","H12","H13","H14","H15","H16","H17","H18","H19","H20","H21","H22","H23"))

a_long$Carbo <- factor(a_long$Carbo, levels = c("Low_carb", "Med_carb", "High_carb"))

a_long <- a_long %>% 
  filter(Carbo != "Not_eating")

ggthemr("fresh", layout = "scientific")

ggplot(a_long[a_long$Class == "Class1", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
  geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 1: 23.54%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class2", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 2:  17.43%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class3", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 3:  7.50%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class4", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 4:  23.53%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class5", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 5:  18.29%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class6", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 6:  9.70%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))
```

## Day 3

```{r eval=FALSE}
set.seed(01012)
max_II <- 100000
lc1 <- poLCA(f, data=dta_d3_wide, nclass=1, na.rm = FALSE, nrep=20, maxiter=max_II) #Loglinear independence model.
lc2 <- poLCA(f, data=dta_d3_wide, nclass=2, na.rm = FALSE, nrep=20, maxiter=max_II)
lc3 <- poLCA(f, data=dta_d3_wide, nclass=3, na.rm = FALSE, nrep=20, maxiter=max_II)
lc4 <- poLCA(f, data=dta_d3_wide, nclass=4, na.rm = FALSE, nrep=20, maxiter=max_II) 
lc5 <- poLCA(f, data=dta_d3_wide, nclass=5, na.rm = FALSE, nrep=20, maxiter=max_II)
lc6 <- poLCA(f, data=dta_d3_wide, nclass=6, na.rm = FALSE, nrep=20, maxiter=max_II)
lc7 <- poLCA(f, data=dta_d3_wide, nclass=7, na.rm = FALSE, nrep=20, maxiter=max_II)
lc8 <- poLCA(f, data=dta_d3_wide, nclass=8, na.rm = FALSE, nrep=20, maxiter=max_II)

# running time:  31.65244 mins
```

### Model comparison and selection

```{r echo=FALSE}
rm(list=ls(all=TRUE))
load("~/Documents/LSHTMproject/Rcode/day3LCA1_6.Rdata")

library(kableExtra)
# dt <- read.csv("../Tablecsv/Modelfitd01.csv", header = T)
# names(dt) <- c("N of classes", "N of parameters", "G^2", "df", "AIC", "BIC", "log-likelihood")
lca_results[1,9] <- "---"
kable(lca_results, digits = 3, row.names = FALSE, align = "c",
              format = "html", caption = "Model Comparison. (Day 3, n = 6151)") %>%
 kable_styling(font_size = 16, 
        bootstrap_options = c("striped", "hover", "condensed"),
        position = "center", full_width = FALSE) %>% 
  row_spec(c(4,6), bold = T, color = "white", background = "#D7261E") %>% 
  footnote(general = "Abbreviation: N, number; resid.df, residual degree of freedom; AIC, Akaike information criterion; BIC, Bayesian information criterion; aBIC, adjusted BIC; cAIC, consistent AIC; Entropy, a pseudo-r-squared index.")
```

### Day 3 data (Fig. 4 classes)

```{r echo=FALSE, fig.height=5, fig.width=14, message=FALSE, warning=FALSE}

library(ggthemr)

# lc5 <- poLCA(f, dta_d1_wide, nclass = 5, graphs = FALSE, maxiter = 10000)

# a <- (data.frame(lc5$probs))
a <- (data.frame(lc4$probs))
# a <- (data.frame(lc6$probs))

# a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5")
a$Class <- c("Class1", "Class2", "Class3", "Class4")

# a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5", "Class6")

a_long <- a %>% 
  gather(Hour, Prob, -Class) %>% 
  separate(Hour, into = c("HourN", "Carbo"), sep = "\\.")

a_long$HourN <- factor(a_long$HourN, levels = c("H0","H1" ,"H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10","H11","H12","H13","H14","H15","H16","H17","H18","H19","H20","H21","H22","H23"))

a_long$Carbo <- factor(a_long$Carbo, levels = c("Low_carb", "Med_carb", "High_carb"))

a_long <- a_long %>% 
  filter(Carbo != "Not_eating")

ggthemr("fresh", layout = "scientific")

ggplot(a_long[a_long$Class == "Class1", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
  geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 1: 8.79%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class2", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 2:  39.49%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class3", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 3:  25.02%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class4", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 4:  26.69%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

# 
# ggplot(a_long[a_long$Class == "Class5", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
#  geom_point(size = 3) +
#   geom_line(size = 1.5) +
#   theme(axis.title = element_text(size = 18),
#         axis.text = element_text(size = 18),
#         axis.line = element_line(colour = "black"),
#         panel.border = element_blank(),
#         panel.background = element_blank(),
#         legend.text = element_text(size = 15),
#         legend.title = element_text(size = 15)
#         # legend.position = "bottom",
#         # legend.direction = "horizontal"
#         ) +
#   labs(title = "Class 5:  18.52%", x = "Hour of the day", y = "Probability",
#        color = "Carbohydrate\nintake") +
#   scale_color_discrete(labels = c("Low", "Medium", "High")) +
#   ylim(c(0,1))
# 
# ggplot(a_long[a_long$Class == "Class6", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
#  geom_point(size = 3) +
#   geom_line(size = 1.5) +
#   theme(axis.title = element_text(size = 18),
#         axis.text = element_text(size = 18),
#         axis.line = element_line(colour = "black"),
#         panel.border = element_blank(),
#         panel.background = element_blank(),
#         legend.text = element_text(size = 15),
#         legend.title = element_text(size = 15)
#         # legend.position = "bottom",
#         # legend.direction = "horizontal"
#         ) +
#   labs(title = "Class 6:  10.02%", x = "Hour of the day", y = "Probability",
#        color = "Carbohydrate\nintake") +
#   scale_color_discrete(labels = c("Low", "Medium", "High")) +
#   ylim(c(0,1))
```

### Day 3 data (Fig. 6 classes)

```{r echo=FALSE, fig.height=5, fig.width=14, message=FALSE, warning=FALSE}

library(ggthemr)

# lc5 <- poLCA(f, dta_d1_wide, nclass = 5, graphs = FALSE, maxiter = 10000)

# a <- (data.frame(lc5$probs))
# a <- (data.frame(lc4$probs))
a <- (data.frame(lc6$probs))

# a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5")
# a$Class <- c("Class1", "Class2", "Class3", "Class4")

a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5", "Class6")

a_long <- a %>% 
  gather(Hour, Prob, -Class) %>% 
  separate(Hour, into = c("HourN", "Carbo"), sep = "\\.")

a_long$HourN <- factor(a_long$HourN, levels = c("H0","H1" ,"H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10","H11","H12","H13","H14","H15","H16","H17","H18","H19","H20","H21","H22","H23"))

a_long$Carbo <- factor(a_long$Carbo, levels = c("Low_carb", "Med_carb", "High_carb"))

a_long <- a_long %>% 
  filter(Carbo != "Not_eating")

ggthemr("fresh", layout = "scientific")

ggplot(a_long[a_long$Class == "Class1", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
  geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 1: 19.41%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class2", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 2:  6.65%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class3", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 3:  15.12%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class4", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 4:  15.43%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class5", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 5:  20.84%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class6", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 6:  22.56%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))
```

## Day 4

```{r eval=FALSE}
set.seed(01012)
max_II <- 100000
lc1<-poLCA(f, data=dta_d4_wide, nclass=1, na.rm = FALSE, nrep=20, maxiter=max_II) #Loglinear independence model.
lc2<-poLCA(f, data=dta_d4_wide, nclass=2, na.rm = FALSE, nrep=20, maxiter=max_II)
lc3<-poLCA(f, data=dta_d4_wide, nclass=3, na.rm = FALSE, nrep=20, maxiter=max_II)
lc4<-poLCA(f, data=dta_d4_wide, nclass=4, na.rm = FALSE, nrep=20, maxiter=max_II) 
lc5<-poLCA(f, data=dta_d4_wide, nclass=5, na.rm = FALSE, nrep=20, maxiter=max_II)
lc6<-poLCA(f, data=dta_d4_wide, nclass=6, na.rm = FALSE, nrep=20, maxiter=max_II)
lc7<-poLCA(f, data=dta_d4_wide, nclass=7, na.rm = FALSE, nrep=20, maxiter=max_II)
lc8<-poLCA(f, data=dta_d4_wide, nclass=8, na.rm = FALSE, nrep=20, maxiter=max_II)

# running time:  27.12666 mins
```

### Model comparison and selection

```{r echo=FALSE}
rm(list=ls(all=TRUE))
load("~/Documents/LSHTMproject/Rcode/day4LCA1_6.Rdata")

library(kableExtra)
# dt <- read.csv("../Tablecsv/Modelfitd01.csv", header = T)
# names(dt) <- c("N of classes", "N of parameters", "G^2", "df", "AIC", "BIC", "log-likelihood")
lca_results[1,9] <- "---"
kable(lca_results, digits = 3, row.names = FALSE, align = "c",
              format = "html", caption = "Model Comparison. (Day 4, n = 6026)") %>%
 kable_styling(font_size = 16, 
        bootstrap_options = c("striped", "hover", "condensed"),
        position = "center", full_width = FALSE) %>% 
  row_spec(c(4,6), bold = T, color = "white", background = "#D7261E") %>% 
  footnote(general = "Abbreviation: N, number; resid.df, residual degree of freedom; AIC, Akaike information criterion; BIC, Bayesian information criterion; aBIC, adjusted BIC; cAIC, consistent AIC; Entropy, a pseudo-r-squared index.")
```

### Day 4 data (Fig. 4 classes)

```{r echo=FALSE, fig.height=5, fig.width=14, message=FALSE, warning=FALSE}

library(ggthemr)

# lc5 <- poLCA(f, dta_d1_wide, nclass = 5, graphs = FALSE, maxiter = 10000)

# a <- (data.frame(lc5$probs))
a <- (data.frame(lc4$probs))
# a <- (data.frame(lc6$probs))

# a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5")
a$Class <- c("Class1", "Class2", "Class3", "Class4")

# a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5", "Class6")

a_long <- a %>% 
  gather(Hour, Prob, -Class) %>% 
  separate(Hour, into = c("HourN", "Carbo"), sep = "\\.")

a_long$HourN <- factor(a_long$HourN, levels = c("H0","H1" ,"H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10","H11","H12","H13","H14","H15","H16","H17","H18","H19","H20","H21","H22","H23"))

a_long$Carbo <- factor(a_long$Carbo, levels = c("Low_carb", "Med_carb", "High_carb"))

a_long <- a_long %>% 
  filter(Carbo != "Not_eating")

ggthemr("fresh", layout = "scientific")

ggplot(a_long[a_long$Class == "Class1", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
  geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 1: 10.11%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class2", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 2:  29.94%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class3", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 3:  24.40%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class4", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 4:  35.55%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

# 
# ggplot(a_long[a_long$Class == "Class5", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
#  geom_point(size = 3) +
#   geom_line(size = 1.5) +
#   theme(axis.title = element_text(size = 18),
#         axis.text = element_text(size = 18),
#         axis.line = element_line(colour = "black"),
#         panel.border = element_blank(),
#         panel.background = element_blank(),
#         legend.text = element_text(size = 15),
#         legend.title = element_text(size = 15)
#         # legend.position = "bottom",
#         # legend.direction = "horizontal"
#         ) +
#   labs(title = "Class 5:  18.52%", x = "Hour of the day", y = "Probability",
#        color = "Carbohydrate\nintake") +
#   scale_color_discrete(labels = c("Low", "Medium", "High")) +
#   ylim(c(0,1))
# 
# ggplot(a_long[a_long$Class == "Class6", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
#  geom_point(size = 3) +
#   geom_line(size = 1.5) +
#   theme(axis.title = element_text(size = 18),
#         axis.text = element_text(size = 18),
#         axis.line = element_line(colour = "black"),
#         panel.border = element_blank(),
#         panel.background = element_blank(),
#         legend.text = element_text(size = 15),
#         legend.title = element_text(size = 15)
#         # legend.position = "bottom",
#         # legend.direction = "horizontal"
#         ) +
#   labs(title = "Class 6:  10.02%", x = "Hour of the day", y = "Probability",
#        color = "Carbohydrate\nintake") +
#   scale_color_discrete(labels = c("Low", "Medium", "High")) +
#   ylim(c(0,1))
```

### Day 4 data (Fig. 6 classes)

```{r echo=FALSE, fig.height=5, fig.width=14, message=FALSE, warning=FALSE}

library(ggthemr)

# lc5 <- poLCA(f, dta_d1_wide, nclass = 5, graphs = FALSE, maxiter = 10000)

# a <- (data.frame(lc5$probs))
# a <- (data.frame(lc4$probs))
a <- (data.frame(lc6$probs))

# a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5")
# a$Class <- c("Class1", "Class2", "Class3", "Class4")

a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5", "Class6")

a_long <- a %>% 
  gather(Hour, Prob, -Class) %>% 
  separate(Hour, into = c("HourN", "Carbo"), sep = "\\.")

a_long$HourN <- factor(a_long$HourN, levels = c("H0","H1" ,"H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10","H11","H12","H13","H14","H15","H16","H17","H18","H19","H20","H21","H22","H23"))

a_long$Carbo <- factor(a_long$Carbo, levels = c("Low_carb", "Med_carb", "High_carb"))

a_long <- a_long %>% 
  filter(Carbo != "Not_eating")

ggthemr("fresh", layout = "scientific")

ggplot(a_long[a_long$Class == "Class1", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
  geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 1: 7.51%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class2", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 2:  27.19%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class3", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 3:  25.77%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class4", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 4:  16.17%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class5", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 5:  7.87%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class6", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 6:  15.49%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))
```