### 2018-07-08 9thJuly/9thJuly.Rmd
mixed results added

---
title: "Preliminary results"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    nature:
      highlightLines: true
---

# Data description

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
setwd("~/Documents/LSHTMproject/Rcode")

load("LCA.Rdata")

library(epiDisplay)
library(plyr)
library(tidyverse)
library(poLCA)
```

## Number of subjects per survey year

```{r message=FALSE, warning=FALSE, paged.print=FALSE}

with(NDNS, tab1(SurveyYear, graph = FALSE, decimal = 2))

```

## Number of subjects per servey year by gender where Men == 1, Women == 2

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
with(NDNS, tabpct(SurveyYear, Sex,  graph = FALSE, decimal = 2))
```

## Summary of their age 
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
NDNS %>% 
  group_by(SurveyYear, Sex) %>% 
  summarise(N = n(), MeanAge = mean(Age), SDAge = sd(Age), minAge = min(Age), maxAge = max(Age))
```

## Distribution of the dietary data (by Day of Week)

### Day1

```{r}
rm(list=ls(all=TRUE))
load("~/Documents/LSHTMproject/Rcode/NDNSday1_4.Rdata")
with(dta_d1_wide, tab1(DayofWeek, graph = FALSE, decimal = 2))
```

### Day2

```{r}
with(dta_d2_wide, tab1(DayofWeek, graph = FALSE, decimal = 2))
```

### Day3

```{r}
with(dta_d3_wide, tab1(DayofWeek, graph = FALSE, decimal = 2))
```

### Day4

```{r}
with(dta_d4_wide, tab1(DayofWeek, graph = FALSE, decimal = 2))
```

## Distribution of the dietary data (by DayNo)

### Monday

```{r}
rm(list=ls(all=TRUE))
load("~/Documents/LSHTMproject/Rcode/NDNSMon_Sun.Rdata")
with(dta_Mon_wide, tab1(DayNo, graph = FALSE, decimal = 2))
```

### Tuesday

```{r}
with(dta_Tue_wide, tab1(DayNo, graph = FALSE, decimal = 2))
```

### Wednesday

```{r}
with(dta_Wed_wide, tab1(DayNo, graph = FALSE, decimal = 2))
```

### Thursday

```{r}
with(dta_Thurs_wide, tab1(DayNo, graph = FALSE, decimal = 2))
```

### Friday

```{r}
with(dta_Fri_wide, tab1(DayNo, graph = FALSE, decimal = 2))
```

### Saturday

```{r}
with(dta_Sat_wide, tab1(DayNo, graph = FALSE, decimal = 2))
```

### Sunday

```{r}
with(dta_Sun_wide, tab1(DayNo, graph = FALSE, decimal = 2))
```

**The problem of analysing data by day of the week would be that every subject only contributed 2-4 days' data. Therefore, we cannot see one subject's dietary change during a Mon-Sun week.**

## Dietary data by day

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
rm(list=ls(all=TRUE))
load("~/Documents/LSHTMproject/Rcode/NDNSday1_4.Rdata")

vecid <- unique(dfs3$id)

vecid1<-unique(dta_day1$id) # n = 6153
vecid2<-unique(dta_day2$id) # n = 6153
vecid3<-unique(dta_day3$id) # n = 6151
vecid4<-unique(dta_day4$id) # n = 6026

setdiff(vecid, vecid1) # two subjects did not have day 1 data

setdiff(vecid, vecid2) # two subjects did not have day 2 data

setdiff(vecid, vecid3) # four subjects did not have day 3 data

setdiff(vecid, vecid4) # 129 subjects did not have day 4 data
```

# Definition of carbohydrates intake groups:

In this analysis, subjects have 2-4 days' dietary data: their food during the participation were self-recorded. We were able to calculate their energy intake each hour, then we also calculated the percentage of energy coming from carbohydrates that contributed to each hour when there is food consumption recorded. **Four categories were identified:**

1. Not eating; 
2. Eating low carbohydrate food (energy contribution less than or equal to 25%); 
3. Eating medium carbohydrate food (energy contribution between 26% and 75%); 
4. Eating high carbohydrate food (energy contribution higher or equal to 75%).

# LCA analyses by day

## Day 1

```{r eval=FALSE}
set.seed(01012)
max_II <- 100000
lc1<-poLCA(f, data=dta_d1_wide, nclass=1, na.rm = FALSE, nrep=20, maxiter=max_II) #Loglinear independence model.
lc2<-poLCA(f, data=dta_d1_wide, nclass=2, na.rm = FALSE, nrep=20, maxiter=max_II)
lc3<-poLCA(f, data=dta_d1_wide, nclass=3, na.rm = FALSE, nrep=20, maxiter=max_II)
lc4<-poLCA(f, data=dta_d1_wide, nclass=4, na.rm = FALSE, nrep=20, maxiter=max_II) 
lc5<-poLCA(f, data=dta_d1_wide, nclass=5, na.rm = FALSE, nrep=20, maxiter=max_II)
lc6<-poLCA(f, data=dta_d1_wide, nclass=6, na.rm = FALSE, nrep=20, maxiter=max_II)
lc7<-poLCA(f, data=dta_d1_wide, nclass=7, na.rm = FALSE, nrep=20, maxiter=max_II)
lc8<-poLCA(f, data=dta_d1_wide, nclass=8, na.rm = FALSE, nrep=20, maxiter=max_II)
```

### Model comparison and selection

```{r echo=FALSE, cache=TRUE}
rm(list=ls(all=TRUE))
load("~/Documents/LSHTMproject/Rcode/day1LCA1_6.Rdata")

library(kableExtra)
# dt <- read.csv("../Tablecsv/Modelfitd01.csv", header = T)
# names(dt) <- c("N of classes", "N of parameters", "G^2", "df", "AIC", "BIC", "log-likelihood")
lca_results[1,9] <- "---"
kable(lca_results, digits = 3, row.names = FALSE, align = "c",
              format = "html", caption = "Model Comparison. (Day 1, n = 6153)") %>%
 kable_styling(font_size = 16, 
        bootstrap_options = c("striped", "hover", "condensed"),
        position = "center", full_width = FALSE) %>% 
  row_spec(7, bold = T, color = "white", background = "#D7261E") %>%
  footnote(general = "Abbreviation: N, number; resid.df, residual degree of freedom; AIC, Akaike information criterion; BIC, Bayesian information criterion; aBIC, adjusted BIC; cAIC, consistent AIC; Entropy, a pseudo-r-squared index.")
```

### Day 1 data (Fig. 7 classes)

```{r echo=FALSE, fig.height=5, fig.width=14, message=FALSE, warning=FALSE, cache=TRUE}

library(ggthemr)

# lc5 <- poLCA(f, dta_d1_wide, nclass = 5, graphs = FALSE, maxiter = 10000)

# a <- (data.frame(lc5$probs))
# a <- (data.frame(lc4$probs))
# a <- (data.frame(lc6$probs))
 a <- (data.frame(lc7$probs))

# a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5")
# a$Class <- c("Class1", "Class2", "Class3", "Class4")

# a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5", "Class6")

a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5", "Class6", "Class7")

a_long <- a %>% 
  gather(Hour, Prob, -Class) %>% 
  separate(Hour, into = c("HourN", "Carbo"), sep = "\\.")

a_long$HourN <- factor(a_long$HourN, levels = c("H0","H1" ,"H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10","H11","H12","H13","H14","H15","H16","H17","H18","H19","H20","H21","H22","H23"))

a_long$Carbo <- factor(a_long$Carbo, levels = c("Low_carb", "Med_carb", "High_carb"))

a_long <- a_long %>% 
  filter(Carbo != "Not_eating")

ggthemr("fresh", layout = "scientific")

ggplot(a_long[a_long$Class == "Class1", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
  geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 1: 18.67%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class2", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 2:  11.81%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class3", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 3:  10.66%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class4", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 4:  6.78%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class5", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 5:  17.45%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class6", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 6:  25.50%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class7", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 7:  9.15%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))
```

<!-- ### Day 1 data (Fig. 6 classes)  -->

<!-- ```{r echo=FALSE, fig.height=5, fig.width=14, message=FALSE, warning=FALSE} -->

<!-- library(ggthemr) -->

<!-- # lc5 <- poLCA(f, dta_d1_wide, nclass = 5, graphs = FALSE, maxiter = 10000) -->

<!-- # a <- (data.frame(lc5$probs)) -->
<!-- # a <- (data.frame(lc4$probs)) -->
<!-- a <- (data.frame(lc6$probs)) -->

<!-- # a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5") -->
<!-- # a$Class <- c("Class1", "Class2", "Class3", "Class4") -->

<!-- a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5", "Class6") -->

<!-- a_long <- a %>%  -->
<!--   gather(Hour, Prob, -Class) %>%  -->
<!--   separate(Hour, into = c("HourN", "Carbo"), sep = "\\.")  -->

<!-- a_long$HourN <- factor(a_long$HourN, levels = c("H0","H1" ,"H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10","H11","H12","H13","H14","H15","H16","H17","H18","H19","H20","H21","H22","H23")) -->

<!-- a_long$Carbo <- factor(a_long$Carbo, levels = c("Low_carb", "Med_carb", "High_carb")) -->

<!-- a_long <- a_long %>%  -->
<!--   filter(Carbo != "Not_eating") -->

<!-- ggthemr("fresh", layout = "scientific") -->

<!-- ggplot(a_long[a_long$Class == "Class1", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +  -->
<!--   geom_point(size = 3) +  -->
<!--   geom_line(size = 1.5) +  -->
<!--   theme(axis.title = element_text(size = 18),  -->
<!--         axis.text = element_text(size = 18),  -->
<!--         axis.line = element_line(colour = "black"),  -->
<!--         panel.border = element_blank(),  -->
<!--         panel.background = element_blank(),  -->
<!--         legend.text = element_text(size = 15),  -->
<!--         legend.title = element_text(size = 15)  -->
<!--         # legend.position = "bottom",  -->
<!--         # legend.direction = "horizontal" -->
<!--         ) +  -->
<!--   labs(title = "Class 1: 26.79%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") +  -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) +  -->
<!--   ylim(c(0,1)) -->

<!-- ggplot(a_long[a_long$Class == "Class2", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +  -->
<!--  geom_point(size = 3) +  -->
<!--   geom_line(size = 1.5) +  -->
<!--   theme(axis.title = element_text(size = 18),  -->
<!--         axis.text = element_text(size = 18),  -->
<!--         axis.line = element_line(colour = "black"),  -->
<!--         panel.border = element_blank(),  -->
<!--         panel.background = element_blank(),  -->
<!--         legend.text = element_text(size = 15),  -->
<!--         legend.title = element_text(size = 15)  -->
<!--         # legend.position = "bottom",  -->
<!--         # legend.direction = "horizontal" -->
<!--         ) +  -->
<!--   labs(title = "Class 2:  11.72%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") +  -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) +  -->
<!--   ylim(c(0,1)) -->

<!-- ggplot(a_long[a_long$Class == "Class3", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +  -->
<!--  geom_point(size = 3) +  -->
<!--   geom_line(size = 1.5) +  -->
<!--   theme(axis.title = element_text(size = 18),  -->
<!--         axis.text = element_text(size = 18),  -->
<!--         axis.line = element_line(colour = "black"),  -->
<!--         panel.border = element_blank(),  -->
<!--         panel.background = element_blank(),  -->
<!--         legend.text = element_text(size = 15),  -->
<!--         legend.title = element_text(size = 15)  -->
<!--         # legend.position = "bottom",  -->
<!--         # legend.direction = "horizontal" -->
<!--         ) +  -->
<!--   labs(title = "Class 3:  7.37%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") +  -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) +  -->
<!--   ylim(c(0,1)) -->

<!-- ggplot(a_long[a_long$Class == "Class4", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +  -->
<!--  geom_point(size = 3) +  -->
<!--   geom_line(size = 1.5) +  -->
<!--   theme(axis.title = element_text(size = 18),  -->
<!--         axis.text = element_text(size = 18),  -->
<!--         axis.line = element_line(colour = "black"),  -->
<!--         panel.border = element_blank(),  -->
<!--         panel.background = element_blank(),  -->
<!--         legend.text = element_text(size = 15),  -->
<!--         legend.title = element_text(size = 15)  -->
<!--         # legend.position = "bottom",  -->
<!--         # legend.direction = "horizontal" -->
<!--         ) +  -->
<!--   labs(title = "Class 4:  25.58%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") +  -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) +  -->
<!--   ylim(c(0,1)) -->

<!-- ggplot(a_long[a_long$Class == "Class5", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + -->
<!--  geom_point(size = 3) + -->
<!--   geom_line(size = 1.5) + -->
<!--   theme(axis.title = element_text(size = 18), -->
<!--         axis.text = element_text(size = 18), -->
<!--         axis.line = element_line(colour = "black"), -->
<!--         panel.border = element_blank(), -->
<!--         panel.background = element_blank(), -->
<!--         legend.text = element_text(size = 15), -->
<!--         legend.title = element_text(size = 15) -->
<!--         # legend.position = "bottom", -->
<!--         # legend.direction = "horizontal" -->
<!--         ) + -->
<!--   labs(title = "Class 5:  18.52%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") + -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) + -->
<!--   ylim(c(0,1)) -->

<!-- ggplot(a_long[a_long$Class == "Class6", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + -->
<!--  geom_point(size = 3) + -->
<!--   geom_line(size = 1.5) + -->
<!--   theme(axis.title = element_text(size = 18), -->
<!--         axis.text = element_text(size = 18), -->
<!--         axis.line = element_line(colour = "black"), -->
<!--         panel.border = element_blank(), -->
<!--         panel.background = element_blank(), -->
<!--         legend.text = element_text(size = 15), -->
<!--         legend.title = element_text(size = 15) -->
<!--         # legend.position = "bottom", -->
<!--         # legend.direction = "horizontal" -->
<!--         ) + -->
<!--   labs(title = "Class 6:  10.02%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") + -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) + -->
<!--   ylim(c(0,1)) -->
<!-- ``` -->

## Day 2

```{r eval=FALSE}
set.seed(01012)
max_II <- 100000
lc1<-poLCA(f, data=dta_d1_wide, nclass=1, na.rm = FALSE, nrep=20, maxiter=max_II) #Loglinear independence model.
lc2<-poLCA(f, data=dta_d1_wide, nclass=2, na.rm = FALSE, nrep=20, maxiter=max_II)
lc3<-poLCA(f, data=dta_d1_wide, nclass=3, na.rm = FALSE, nrep=20, maxiter=max_II)
lc4<-poLCA(f, data=dta_d1_wide, nclass=4, na.rm = FALSE, nrep=20, maxiter=max_II) 
lc5<-poLCA(f, data=dta_d1_wide, nclass=5, na.rm = FALSE, nrep=20, maxiter=max_II)
lc6<-poLCA(f, data=dta_d1_wide, nclass=6, na.rm = FALSE, nrep=20, maxiter=max_II)
lc7<-poLCA(f, data=dta_d1_wide, nclass=7, na.rm = FALSE, nrep=20, maxiter=max_II)
lc8<-poLCA(f, data=dta_d1_wide, nclass=8, na.rm = FALSE, nrep=20, maxiter=max_II)

# running time:  31.60514 mins
```

### Model comparison and selection

```{r echo=FALSE, cache=TRUE}
rm(list=ls(all=TRUE))
load("~/Documents/LSHTMproject/Rcode/day2LCA1_6.Rdata")

library(kableExtra)
# dt <- read.csv("../Tablecsv/Modelfitd01.csv", header = T)
# names(dt) <- c("N of classes", "N of parameters", "G^2", "df", "AIC", "BIC", "log-likelihood")
lca_results[1,9] <- "---"
kable(lca_results, digits = 3, row.names = FALSE, align = "c",
              format = "html", caption = "Model Comparison. (Day 2, n = 6153)") %>%
 kable_styling(font_size = 16, 
        bootstrap_options = c("striped", "hover", "condensed"),
        position = "center", full_width = FALSE) %>% 
  row_spec(7, bold = T, color = "white", background = "#D7261E") %>% 
  footnote(general = "Abbreviation: N, number; resid.df, residual degree of freedom; AIC, Akaike information criterion; BIC, Bayesian information criterion; aBIC, adjusted BIC; cAIC, consistent AIC; Entropy, a pseudo-r-squared index.")
```

### Day 2 data (Fig. 7 classes)

```{r echo=FALSE, fig.height=5, fig.width=14, message=FALSE, warning=FALSE, cache=TRUE}

library(ggthemr)

# lc5 <- poLCA(f, dta_d1_wide, nclass = 5, graphs = FALSE, maxiter = 10000)

# a <- (data.frame(lc5$probs))
# a <- (data.frame(lc4$probs))
# a <- (data.frame(lc6$probs))
a <- (data.frame(lc7$probs))

# a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5")
# a$Class <- c("Class1", "Class2", "Class3", "Class4")

# a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5", "Class6")
a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5", "Class6", "Class7")

a_long <- a %>% 
  gather(Hour, Prob, -Class) %>% 
  separate(Hour, into = c("HourN", "Carbo"), sep = "\\.")

a_long$HourN <- factor(a_long$HourN, levels = c("H0","H1" ,"H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10","H11","H12","H13","H14","H15","H16","H17","H18","H19","H20","H21","H22","H23"))

a_long$Carbo <- factor(a_long$Carbo, levels = c("Low_carb", "Med_carb", "High_carb"))

a_long <- a_long %>% 
  filter(Carbo != "Not_eating")

ggthemr("fresh", layout = "scientific")

ggplot(a_long[a_long$Class == "Class1", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
  geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 1: 18.67%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class2", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 2:  11.81%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class3", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 3:  10.66%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class4", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 4:  6.78%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class5", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 5:  17.45%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class6", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 6:  25.50%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class7", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 7:  9.15%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))
```

<!-- ### Day 2 data (Fig. 6 classes)  -->

<!-- ```{r echo=FALSE, fig.height=5, fig.width=14, message=FALSE, warning=FALSE} -->

<!-- library(ggthemr) -->

<!-- # lc5 <- poLCA(f, dta_d1_wide, nclass = 5, graphs = FALSE, maxiter = 10000) -->

<!-- # a <- (data.frame(lc5$probs)) -->
<!-- # a <- (data.frame(lc4$probs)) -->
<!-- a <- (data.frame(lc6$probs)) -->

<!-- # a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5") -->
<!-- # a$Class <- c("Class1", "Class2", "Class3", "Class4") -->

<!-- a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5", "Class6") -->

<!-- a_long <- a %>%  -->
<!--   gather(Hour, Prob, -Class) %>%  -->
<!--   separate(Hour, into = c("HourN", "Carbo"), sep = "\\.")  -->

<!-- a_long$HourN <- factor(a_long$HourN, levels = c("H0","H1" ,"H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10","H11","H12","H13","H14","H15","H16","H17","H18","H19","H20","H21","H22","H23")) -->

<!-- a_long$Carbo <- factor(a_long$Carbo, levels = c("Low_carb", "Med_carb", "High_carb")) -->

<!-- a_long <- a_long %>%  -->
<!--   filter(Carbo != "Not_eating") -->

<!-- ggthemr("fresh", layout = "scientific") -->

<!-- ggplot(a_long[a_long$Class == "Class1", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +  -->
<!--   geom_point(size = 3) +  -->
<!--   geom_line(size = 1.5) +  -->
<!--   theme(axis.title = element_text(size = 18),  -->
<!--         axis.text = element_text(size = 18),  -->
<!--         axis.line = element_line(colour = "black"),  -->
<!--         panel.border = element_blank(),  -->
<!--         panel.background = element_blank(),  -->
<!--         legend.text = element_text(size = 15),  -->
<!--         legend.title = element_text(size = 15)  -->
<!--         # legend.position = "bottom",  -->
<!--         # legend.direction = "horizontal" -->
<!--         ) +  -->
<!--   labs(title = "Class 1: 23.54%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") +  -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) +  -->
<!--   ylim(c(0,1)) -->

<!-- ggplot(a_long[a_long$Class == "Class2", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +  -->
<!--  geom_point(size = 3) +  -->
<!--   geom_line(size = 1.5) +  -->
<!--   theme(axis.title = element_text(size = 18),  -->
<!--         axis.text = element_text(size = 18),  -->
<!--         axis.line = element_line(colour = "black"),  -->
<!--         panel.border = element_blank(),  -->
<!--         panel.background = element_blank(),  -->
<!--         legend.text = element_text(size = 15),  -->
<!--         legend.title = element_text(size = 15)  -->
<!--         # legend.position = "bottom",  -->
<!--         # legend.direction = "horizontal" -->
<!--         ) +  -->
<!--   labs(title = "Class 2:  17.43%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") +  -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) +  -->
<!--   ylim(c(0,1)) -->

<!-- ggplot(a_long[a_long$Class == "Class3", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +  -->
<!--  geom_point(size = 3) +  -->
<!--   geom_line(size = 1.5) +  -->
<!--   theme(axis.title = element_text(size = 18),  -->
<!--         axis.text = element_text(size = 18),  -->
<!--         axis.line = element_line(colour = "black"),  -->
<!--         panel.border = element_blank(),  -->
<!--         panel.background = element_blank(),  -->
<!--         legend.text = element_text(size = 15),  -->
<!--         legend.title = element_text(size = 15)  -->
<!--         # legend.position = "bottom",  -->
<!--         # legend.direction = "horizontal" -->
<!--         ) +  -->
<!--   labs(title = "Class 3:  7.50%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") +  -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) +  -->
<!--   ylim(c(0,1)) -->

<!-- ggplot(a_long[a_long$Class == "Class4", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +  -->
<!--  geom_point(size = 3) +  -->
<!--   geom_line(size = 1.5) +  -->
<!--   theme(axis.title = element_text(size = 18),  -->
<!--         axis.text = element_text(size = 18),  -->
<!--         axis.line = element_line(colour = "black"),  -->
<!--         panel.border = element_blank(),  -->
<!--         panel.background = element_blank(),  -->
<!--         legend.text = element_text(size = 15),  -->
<!--         legend.title = element_text(size = 15)  -->
<!--         # legend.position = "bottom",  -->
<!--         # legend.direction = "horizontal" -->
<!--         ) +  -->
<!--   labs(title = "Class 4:  23.53%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") +  -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) +  -->
<!--   ylim(c(0,1)) -->

<!-- ggplot(a_long[a_long$Class == "Class5", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + -->
<!--  geom_point(size = 3) + -->
<!--   geom_line(size = 1.5) + -->
<!--   theme(axis.title = element_text(size = 18), -->
<!--         axis.text = element_text(size = 18), -->
<!--         axis.line = element_line(colour = "black"), -->
<!--         panel.border = element_blank(), -->
<!--         panel.background = element_blank(), -->
<!--         legend.text = element_text(size = 15), -->
<!--         legend.title = element_text(size = 15) -->
<!--         # legend.position = "bottom", -->
<!--         # legend.direction = "horizontal" -->
<!--         ) + -->
<!--   labs(title = "Class 5:  18.29%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") + -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) + -->
<!--   ylim(c(0,1)) -->

<!-- ggplot(a_long[a_long$Class == "Class6", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + -->
<!--  geom_point(size = 3) + -->
<!--   geom_line(size = 1.5) + -->
<!--   theme(axis.title = element_text(size = 18), -->
<!--         axis.text = element_text(size = 18), -->
<!--         axis.line = element_line(colour = "black"), -->
<!--         panel.border = element_blank(), -->
<!--         panel.background = element_blank(), -->
<!--         legend.text = element_text(size = 15), -->
<!--         legend.title = element_text(size = 15) -->
<!--         # legend.position = "bottom", -->
<!--         # legend.direction = "horizontal" -->
<!--         ) + -->
<!--   labs(title = "Class 6:  9.70%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") + -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) + -->
<!--   ylim(c(0,1)) -->
<!-- ``` -->

## Day 3

```{r eval=FALSE}
set.seed(01012)
max_II <- 100000
lc1 <- poLCA(f, data=dta_d3_wide, nclass=1, na.rm = FALSE, nrep=20, maxiter=max_II) #Loglinear independence model.
lc2 <- poLCA(f, data=dta_d3_wide, nclass=2, na.rm = FALSE, nrep=20, maxiter=max_II)
lc3 <- poLCA(f, data=dta_d3_wide, nclass=3, na.rm = FALSE, nrep=20, maxiter=max_II)
lc4 <- poLCA(f, data=dta_d3_wide, nclass=4, na.rm = FALSE, nrep=20, maxiter=max_II) 
lc5 <- poLCA(f, data=dta_d3_wide, nclass=5, na.rm = FALSE, nrep=20, maxiter=max_II)
lc6 <- poLCA(f, data=dta_d3_wide, nclass=6, na.rm = FALSE, nrep=20, maxiter=max_II)
lc7 <- poLCA(f, data=dta_d3_wide, nclass=7, na.rm = FALSE, nrep=20, maxiter=max_II)
lc8 <- poLCA(f, data=dta_d3_wide, nclass=8, na.rm = FALSE, nrep=20, maxiter=max_II)

# running time:  31.65244 mins
```

### Model comparison and selection

```{r echo=FALSE, cache=TRUE}
rm(list=ls(all=TRUE))
load("~/Documents/LSHTMproject/Rcode/day3LCA1_6.Rdata")

library(kableExtra)
# dt <- read.csv("../Tablecsv/Modelfitd01.csv", header = T)
# names(dt) <- c("N of classes", "N of parameters", "G^2", "df", "AIC", "BIC", "log-likelihood")
lca_results[1,9] <- "---"
kable(lca_results, digits = 3, row.names = FALSE, align = "c",
              format = "html", caption = "Model Comparison. (Day 3, n = 6151)") %>%
 kable_styling(font_size = 16, 
        bootstrap_options = c("striped", "hover", "condensed"),
        position = "center", full_width = FALSE) %>% 
  row_spec(7, bold = T, color = "white", background = "#D7261E") %>% 
  footnote(general = "Abbreviation: N, number; resid.df, residual degree of freedom; AIC, Akaike information criterion; BIC, Bayesian information criterion; aBIC, adjusted BIC; cAIC, consistent AIC; Entropy, a pseudo-r-squared index.")
```

### Day 3 data (Fig. 7 classes)

```{r echo=FALSE, fig.height=5, fig.width=14, message=FALSE, warning=FALSE, cache=TRUE}

library(ggthemr)

# lc5 <- poLCA(f, dta_d1_wide, nclass = 5, graphs = FALSE, maxiter = 10000)

# a <- (data.frame(lc5$probs))
# a <- (data.frame(lc4$probs))
# a <- (data.frame(lc6$probs))
a <- (data.frame(lc7$probs))

# a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5")
# a$Class <- c("Class1", "Class2", "Class3", "Class4")

# a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5", "Class6")

a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5", "Class6", "Class7")

a_long <- a %>% 
  gather(Hour, Prob, -Class) %>% 
  separate(Hour, into = c("HourN", "Carbo"), sep = "\\.")

a_long$HourN <- factor(a_long$HourN, levels = c("H0","H1" ,"H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10","H11","H12","H13","H14","H15","H16","H17","H18","H19","H20","H21","H22","H23"))

a_long$Carbo <- factor(a_long$Carbo, levels = c("Low_carb", "Med_carb", "High_carb"))

a_long <- a_long %>% 
  filter(Carbo != "Not_eating")

ggthemr("fresh", layout = "scientific")

ggplot(a_long[a_long$Class == "Class1", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
  geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 1: 6.13%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class2", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 2:  13.32%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class3", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 3:  17.95%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class4", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 4:  8.16%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class5", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 5:  20.48%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class6", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 6:  20.58%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class7", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 7:  13.38%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))
```

<!-- ### Day 3 data (Fig. 6 classes)  -->

<!-- ```{r echo=FALSE, fig.height=5, fig.width=14, message=FALSE, warning=FALSE} -->

<!-- library(ggthemr) -->

<!-- # lc5 <- poLCA(f, dta_d1_wide, nclass = 5, graphs = FALSE, maxiter = 10000) -->

<!-- # a <- (data.frame(lc5$probs)) -->
<!-- # a <- (data.frame(lc4$probs)) -->
<!-- a <- (data.frame(lc6$probs)) -->

<!-- # a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5") -->
<!-- # a$Class <- c("Class1", "Class2", "Class3", "Class4") -->

<!-- a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5", "Class6") -->

<!-- a_long <- a %>%  -->
<!--   gather(Hour, Prob, -Class) %>%  -->
<!--   separate(Hour, into = c("HourN", "Carbo"), sep = "\\.")  -->

<!-- a_long$HourN <- factor(a_long$HourN, levels = c("H0","H1" ,"H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10","H11","H12","H13","H14","H15","H16","H17","H18","H19","H20","H21","H22","H23")) -->

<!-- a_long$Carbo <- factor(a_long$Carbo, levels = c("Low_carb", "Med_carb", "High_carb")) -->

<!-- a_long <- a_long %>%  -->
<!--   filter(Carbo != "Not_eating") -->

<!-- ggthemr("fresh", layout = "scientific") -->

<!-- ggplot(a_long[a_long$Class == "Class1", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +  -->
<!--   geom_point(size = 3) +  -->
<!--   geom_line(size = 1.5) +  -->
<!--   theme(axis.title = element_text(size = 18),  -->
<!--         axis.text = element_text(size = 18),  -->
<!--         axis.line = element_line(colour = "black"),  -->
<!--         panel.border = element_blank(),  -->
<!--         panel.background = element_blank(),  -->
<!--         legend.text = element_text(size = 15),  -->
<!--         legend.title = element_text(size = 15)  -->
<!--         # legend.position = "bottom",  -->
<!--         # legend.direction = "horizontal" -->
<!--         ) +  -->
<!--   labs(title = "Class 1: 19.41%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") +  -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) +  -->
<!--   ylim(c(0,1)) -->

<!-- ggplot(a_long[a_long$Class == "Class2", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +  -->
<!--  geom_point(size = 3) +  -->
<!--   geom_line(size = 1.5) +  -->
<!--   theme(axis.title = element_text(size = 18),  -->
<!--         axis.text = element_text(size = 18),  -->
<!--         axis.line = element_line(colour = "black"),  -->
<!--         panel.border = element_blank(),  -->
<!--         panel.background = element_blank(),  -->
<!--         legend.text = element_text(size = 15),  -->
<!--         legend.title = element_text(size = 15)  -->
<!--         # legend.position = "bottom",  -->
<!--         # legend.direction = "horizontal" -->
<!--         ) +  -->
<!--   labs(title = "Class 2:  6.65%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") +  -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) +  -->
<!--   ylim(c(0,1)) -->

<!-- ggplot(a_long[a_long$Class == "Class3", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +  -->
<!--  geom_point(size = 3) +  -->
<!--   geom_line(size = 1.5) +  -->
<!--   theme(axis.title = element_text(size = 18),  -->
<!--         axis.text = element_text(size = 18),  -->
<!--         axis.line = element_line(colour = "black"),  -->
<!--         panel.border = element_blank(),  -->
<!--         panel.background = element_blank(),  -->
<!--         legend.text = element_text(size = 15),  -->
<!--         legend.title = element_text(size = 15)  -->
<!--         # legend.position = "bottom",  -->
<!--         # legend.direction = "horizontal" -->
<!--         ) +  -->
<!--   labs(title = "Class 3:  15.12%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") +  -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) +  -->
<!--   ylim(c(0,1)) -->

<!-- ggplot(a_long[a_long$Class == "Class4", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +  -->
<!--  geom_point(size = 3) +  -->
<!--   geom_line(size = 1.5) +  -->
<!--   theme(axis.title = element_text(size = 18),  -->
<!--         axis.text = element_text(size = 18),  -->
<!--         axis.line = element_line(colour = "black"),  -->
<!--         panel.border = element_blank(),  -->
<!--         panel.background = element_blank(),  -->
<!--         legend.text = element_text(size = 15),  -->
<!--         legend.title = element_text(size = 15)  -->
<!--         # legend.position = "bottom",  -->
<!--         # legend.direction = "horizontal" -->
<!--         ) +  -->
<!--   labs(title = "Class 4:  15.43%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") +  -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) +  -->
<!--   ylim(c(0,1)) -->

<!-- ggplot(a_long[a_long$Class == "Class5", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + -->
<!--  geom_point(size = 3) + -->
<!--   geom_line(size = 1.5) + -->
<!--   theme(axis.title = element_text(size = 18), -->
<!--         axis.text = element_text(size = 18), -->
<!--         axis.line = element_line(colour = "black"), -->
<!--         panel.border = element_blank(), -->
<!--         panel.background = element_blank(), -->
<!--         legend.text = element_text(size = 15), -->
<!--         legend.title = element_text(size = 15) -->
<!--         # legend.position = "bottom", -->
<!--         # legend.direction = "horizontal" -->
<!--         ) + -->
<!--   labs(title = "Class 5:  20.84%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") + -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) + -->
<!--   ylim(c(0,1)) -->

<!-- ggplot(a_long[a_long$Class == "Class6", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + -->
<!--  geom_point(size = 3) + -->
<!--   geom_line(size = 1.5) + -->
<!--   theme(axis.title = element_text(size = 18), -->
<!--         axis.text = element_text(size = 18), -->
<!--         axis.line = element_line(colour = "black"), -->
<!--         panel.border = element_blank(), -->
<!--         panel.background = element_blank(), -->
<!--         legend.text = element_text(size = 15), -->
<!--         legend.title = element_text(size = 15) -->
<!--         # legend.position = "bottom", -->
<!--         # legend.direction = "horizontal" -->
<!--         ) + -->
<!--   labs(title = "Class 6:  22.56%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") + -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) + -->
<!--   ylim(c(0,1)) -->
<!-- ``` -->

## Day 4

```{r eval=FALSE}
set.seed(01012)
max_II <- 100000
lc1<-poLCA(f, data=dta_d4_wide, nclass=1, na.rm = FALSE, nrep=20, maxiter=max_II) #Loglinear independence model.
lc2<-poLCA(f, data=dta_d4_wide, nclass=2, na.rm = FALSE, nrep=20, maxiter=max_II)
lc3<-poLCA(f, data=dta_d4_wide, nclass=3, na.rm = FALSE, nrep=20, maxiter=max_II)
lc4<-poLCA(f, data=dta_d4_wide, nclass=4, na.rm = FALSE, nrep=20, maxiter=max_II) 
lc5<-poLCA(f, data=dta_d4_wide, nclass=5, na.rm = FALSE, nrep=20, maxiter=max_II)
lc6<-poLCA(f, data=dta_d4_wide, nclass=6, na.rm = FALSE, nrep=20, maxiter=max_II)
lc7<-poLCA(f, data=dta_d4_wide, nclass=7, na.rm = FALSE, nrep=20, maxiter=max_II)
lc8<-poLCA(f, data=dta_d4_wide, nclass=8, na.rm = FALSE, nrep=20, maxiter=max_II)

# running time:  27.12666 mins
```

### Model comparison and selection

```{r echo=FALSE, cache=TRUE}
rm(list=ls(all=TRUE))
load("~/Documents/LSHTMproject/Rcode/day4LCA1_6.Rdata")

library(kableExtra)
# dt <- read.csv("../Tablecsv/Modelfitd01.csv", header = T)
# names(dt) <- c("N of classes", "N of parameters", "G^2", "df", "AIC", "BIC", "log-likelihood")
lca_results[1,9] <- "---"
kable(lca_results, digits = 3, row.names = FALSE, align = "c",
              format = "html", caption = "Model Comparison. (Day 4, n = 6026)") %>%
 kable_styling(font_size = 16, 
        bootstrap_options = c("striped", "hover", "condensed"),
        position = "center", full_width = FALSE) %>% 
  row_spec(8, bold = T, color = "white", background = "#D7261E") %>% 
  footnote(general = "Abbreviation: N, number; resid.df, residual degree of freedom; AIC, Akaike information criterion; BIC, Bayesian information criterion; aBIC, adjusted BIC; cAIC, consistent AIC; Entropy, a pseudo-r-squared index.")
```

### Day 4 data (Fig. 8 classes)

```{r echo=FALSE, fig.height=5, fig.width=14, message=FALSE, warning=FALSE, cache=TRUE}

library(ggthemr)

# lc5 <- poLCA(f, dta_d1_wide, nclass = 5, graphs = FALSE, maxiter = 10000)

# a <- (data.frame(lc5$probs))
# a <- (data.frame(lc4$probs))
# a <- (data.frame(lc6$probs))
a <- (data.frame(lc8$probs))

# a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5")
# a$Class <- c("Class1", "Class2", "Class3", "Class4")

# a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5", "Class6")

a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5", "Class6", "Class7", "Class8")

a_long <- a %>% 
  gather(Hour, Prob, -Class) %>% 
  separate(Hour, into = c("HourN", "Carbo"), sep = "\\.")

a_long$HourN <- factor(a_long$HourN, levels = c("H0","H1" ,"H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10","H11","H12","H13","H14","H15","H16","H17","H18","H19","H20","H21","H22","H23"))

a_long$Carbo <- factor(a_long$Carbo, levels = c("Low_carb", "Med_carb", "High_carb"))

a_long <- a_long %>% 
  filter(Carbo != "Not_eating")

ggthemr("fresh", layout = "scientific")

ggplot(a_long[a_long$Class == "Class1", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
  geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 1: 6.67%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class2", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 2:  10.61%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class3", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 3:  14.93%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class4", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 4:  20.54%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class5", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 5:  5.76%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class6", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 6:  16.97%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class7", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 7:  12.17%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))

ggplot(a_long[a_long$Class == "Class8", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 8:  12.36%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))
```

<!-- ### Day 4 data (Fig. 6 classes)  -->

<!-- ```{r echo=FALSE, fig.height=5, fig.width=14, message=FALSE, warning=FALSE} -->

<!-- library(ggthemr) -->

<!-- # lc5 <- poLCA(f, dta_d1_wide, nclass = 5, graphs = FALSE, maxiter = 10000) -->

<!-- # a <- (data.frame(lc5$probs)) -->
<!-- # a <- (data.frame(lc4$probs)) -->
<!-- a <- (data.frame(lc6$probs)) -->

<!-- # a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5") -->
<!-- # a$Class <- c("Class1", "Class2", "Class3", "Class4") -->

<!-- a$Class <- c("Class1", "Class2", "Class3", "Class4", "Class5", "Class6") -->

<!-- a_long <- a %>%  -->
<!--   gather(Hour, Prob, -Class) %>%  -->
<!--   separate(Hour, into = c("HourN", "Carbo"), sep = "\\.")  -->

<!-- a_long$HourN <- factor(a_long$HourN, levels = c("H0","H1" ,"H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10","H11","H12","H13","H14","H15","H16","H17","H18","H19","H20","H21","H22","H23")) -->

<!-- a_long$Carbo <- factor(a_long$Carbo, levels = c("Low_carb", "Med_carb", "High_carb")) -->

<!-- a_long <- a_long %>%  -->
<!--   filter(Carbo != "Not_eating") -->

<!-- ggthemr("fresh", layout = "scientific") -->

<!-- ggplot(a_long[a_long$Class == "Class1", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +  -->
<!--   geom_point(size = 3) +  -->
<!--   geom_line(size = 1.5) +  -->
<!--   theme(axis.title = element_text(size = 18),  -->
<!--         axis.text = element_text(size = 18),  -->
<!--         axis.line = element_line(colour = "black"),  -->
<!--         panel.border = element_blank(),  -->
<!--         panel.background = element_blank(),  -->
<!--         legend.text = element_text(size = 15),  -->
<!--         legend.title = element_text(size = 15)  -->
<!--         # legend.position = "bottom",  -->
<!--         # legend.direction = "horizontal" -->
<!--         ) +  -->
<!--   labs(title = "Class 1: 7.51%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") +  -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) +  -->
<!--   ylim(c(0,1)) -->

<!-- ggplot(a_long[a_long$Class == "Class2", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +  -->
<!--  geom_point(size = 3) +  -->
<!--   geom_line(size = 1.5) +  -->
<!--   theme(axis.title = element_text(size = 18),  -->
<!--         axis.text = element_text(size = 18),  -->
<!--         axis.line = element_line(colour = "black"),  -->
<!--         panel.border = element_blank(),  -->
<!--         panel.background = element_blank(),  -->
<!--         legend.text = element_text(size = 15),  -->
<!--         legend.title = element_text(size = 15)  -->
<!--         # legend.position = "bottom",  -->
<!--         # legend.direction = "horizontal" -->
<!--         ) +  -->
<!--   labs(title = "Class 2:  27.19%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") +  -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) +  -->
<!--   ylim(c(0,1)) -->

<!-- ggplot(a_long[a_long$Class == "Class3", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +  -->
<!--  geom_point(size = 3) +  -->
<!--   geom_line(size = 1.5) +  -->
<!--   theme(axis.title = element_text(size = 18),  -->
<!--         axis.text = element_text(size = 18),  -->
<!--         axis.line = element_line(colour = "black"),  -->
<!--         panel.border = element_blank(),  -->
<!--         panel.background = element_blank(),  -->
<!--         legend.text = element_text(size = 15),  -->
<!--         legend.title = element_text(size = 15)  -->
<!--         # legend.position = "bottom",  -->
<!--         # legend.direction = "horizontal" -->
<!--         ) +  -->
<!--   labs(title = "Class 3:  25.77%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") +  -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) +  -->
<!--   ylim(c(0,1)) -->

<!-- ggplot(a_long[a_long$Class == "Class4", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +  -->
<!--  geom_point(size = 3) +  -->
<!--   geom_line(size = 1.5) +  -->
<!--   theme(axis.title = element_text(size = 18),  -->
<!--         axis.text = element_text(size = 18),  -->
<!--         axis.line = element_line(colour = "black"),  -->
<!--         panel.border = element_blank(),  -->
<!--         panel.background = element_blank(),  -->
<!--         legend.text = element_text(size = 15),  -->
<!--         legend.title = element_text(size = 15)  -->
<!--         # legend.position = "bottom",  -->
<!--         # legend.direction = "horizontal" -->
<!--         ) +  -->
<!--   labs(title = "Class 4:  16.17%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") +  -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) +  -->
<!--   ylim(c(0,1)) -->

<!-- ggplot(a_long[a_long$Class == "Class5", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + -->
<!--  geom_point(size = 3) + -->
<!--   geom_line(size = 1.5) + -->
<!--   theme(axis.title = element_text(size = 18), -->
<!--         axis.text = element_text(size = 18), -->
<!--         axis.line = element_line(colour = "black"), -->
<!--         panel.border = element_blank(), -->
<!--         panel.background = element_blank(), -->
<!--         legend.text = element_text(size = 15), -->
<!--         legend.title = element_text(size = 15) -->
<!--         # legend.position = "bottom", -->
<!--         # legend.direction = "horizontal" -->
<!--         ) + -->
<!--   labs(title = "Class 5:  7.87%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") + -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) + -->
<!--   ylim(c(0,1)) -->

<!-- ggplot(a_long[a_long$Class == "Class6", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + -->
<!--  geom_point(size = 3) + -->
<!--   geom_line(size = 1.5) + -->
<!--   theme(axis.title = element_text(size = 18), -->
<!--         axis.text = element_text(size = 18), -->
<!--         axis.line = element_line(colour = "black"), -->
<!--         panel.border = element_blank(), -->
<!--         panel.background = element_blank(), -->
<!--         legend.text = element_text(size = 15), -->
<!--         legend.title = element_text(size = 15) -->
<!--         # legend.position = "bottom", -->
<!--         # legend.direction = "horizontal" -->
<!--         ) + -->
<!--   labs(title = "Class 6:  15.49%", x = "Hour of the day", y = "Probability", -->
<!--        color = "Carbohydrate\nintake") + -->
<!--   scale_color_discrete(labels = c("Low", "Medium", "High")) + -->
<!--   ylim(c(0,1)) -->
<!-- ``` -->

# How to deal with within-individual variation?

1. LTA (latent transit analysis): this method can help calculate the probabilities of transition from one class to another on condition of the previous (occasion's) class. I don't think we can apply this method since every subject in the NDNS data set has a different four days of Week:
    - For example subject A provided dietary data for Monday (day1), Tuesday (day2), Wednesday (day3), and Thursday (day4); 
    - However, subject B provided dietary data on Thursday (day1), Friday (day2), Saturday (day3), and Sunday (day4); 
    - Using day 1 (or any day from the four days) data alone would be not appropriate since different people started from different day, their day 1 might be weekday or weekends. 
    - So using LTA analysis would be a good choice when all of the subjects were measured on the same day of the week. Here, in our case, it may not be a good option.

2. LCA with random effect: this approach is actually using all of the data that we have, taking the personal variation into account. By saying random effect here, we mean in the data set, some observations are not independent from each other (every 2-4 observations are from an exact same person), in other words, **they are nested within the person**. Then we should tell software that it is the persons rather than the observations are independent from each other. This is more reasonable than enforcing a subject to stay in the same class for all four days, because it is certainly possible that one person can belong to a class for a few days, and jump to another class on the other day(s) within the survey.

Therefore, we chose the 2nd method in the analyses. This can be done by [installing the LCA macro in the SAS program (https://methodology.psu.edu/downloads/proclcalta)](https://methodology.psu.edu/downloads/proclcalta), and using the `PROC LCA` command.

# Mixed effect LCA (or shall we call it multilevel LCA?) results

## Input and Output example when fitting a three class LCA model

### SAS code in running the three classes mixed effect LCA analysis on the NDNS data

```{r  eval=FALSE}
PROC LCA DATA=store.NDNS OUTPOST=STORE.NDNS_PP3 OUTEST=STORE.NDNS_EST3
            OUTPARAM=STORE.NDNS_PARAM3; /* input data and output (results, estimation, posterior probabilities) dataset names */
NCLASS 3; /* number of classes fitting*/                                  
ID ID_DAY ID SEX DAYOFWEEK AGE; /* variable names want to keep in the output datasets */
ITEMS H0 H1 H2 H3 H4 H5 H6 H7 H8 H9 H10 H11 H12 H13 H14 H15 H16 H17 H18 H19 H20 H21 H22 H23;
CATEGORIES 4 4 4 4 4 4 4 4 4 4 4 4 4 4 4 4 4 4 4 4 4 4 4 4;
CLUSTERS ID; /* this is the key, to specify ID as the cluster variable */ 
MAXITER 100000;
SEED 100000;
RUN;
```

### SAS output of running the mixed effect LCA analysis

The following output shows that there are 24483 observations in the NDNS data set, the number of cluster is 6155 (which matches the sample size what we described in the description section). Smallest/largest cluster size is 2/4 indicating the number of observations within individual varies between 2 to 4.

```{r eval=FALSE}
Data Summary, Model Information, and Fit Statistics (EM Algorithm)

Number of subjects in dataset:       24483 
Number of subjects in analysis:      24483
Number of clusters in analysis:       6155 
Smallest cluster size:                   2 
Largest cluster size:                    4

Number of measurement items:            24
Response categories per item:            4 4 4 4 4 4 4 4 4 4 4 4 4 4 4 4 4 4 4 4 4 4 4 4
Number of groups in the data:            1
Number of latent classes:                3

The analysis includes clusters.
Cluster variable name:  id
Rho starting values were randomly generated (seed = 100000).

No parameter restrictions were specified (freely estimated).

The model converged in 298 iterations.

Maximum number of iterations: 100000
Convergence method: maximum absolute deviation (MAD)
Convergence criterion:  0.000001000

=============================================
Fit statistics:
=============================================

Log-likelihood:   -380275.38
G-squared:         275662.53
AIC:               276098.53
BIC:               277865.58
CAIC:              278083.58
Adjusted BIC:      277172.79
Entropy:                0.85
(Based on the pseudo-likelihood.)
```

### Example of the posterior probabilities estimated by the model

Where:

- **POSTLC1**: posterior probability of belonging to latent class 1 given the subject's carbohydrate eating pattern on that day. 
- **POSTLC2**: posterior probability of belonging to latent class 2 given the subject's carbohydrate eating pattern on that day. 
- **POSTLC3**: posterior probability of belonging to latent class 3 given the subject's carbohydrate eating pattern on that day.
- **BEST**: The latent class group assigned by the model (highest probability among POSTLC1, POSTLC2, and, POSTLC3).

```{r echo=FALSE}
rm(list=ls(all=TRUE))
dt <- read.csv("../Tablecsv/ndns_pp3.csv", header = T)
# names(dt) <- c("N of classes", "log-likelihood", "G^2",  "AIC", "BIC", "cAIC", "aBIC", "Entropy")
# dt[1,8] <- "---"

library(DT)
dt <- dt %>% 
  select(id, Sex, DayofWeek, Age, POSTLC1, POSTLC2, POSTLC3, BEST) %>% 
  filter(id < 40000000)
datatable(dt, options = list(
  searching = TRUE,
  pageLength = 15,
  lengthMenu = c(5, 10, 15, 20)
)) %>% 
  formatRound("POSTLC1", 3) %>% 
  formatRound("POSTLC2", 3) %>% 
  formatRound("POSTLC3", 3) 
```

## Model comparison and selection

```{r echo=FALSE}
rm(list=ls(all=TRUE))
# load("~/Documents/LSHTMproject/Rcode/day4LCA1_6.Rdata")

library(kableExtra)
dt <- read.csv("../Tablecsv/ModelsMixed.csv", header = T)
names(dt) <- c("N of classes", "log-likelihood", "G^2",  "AIC", "BIC", "cAIC", "aBIC", "Entropy")
dt[1,8] <- "---"
kable(dt, digits = 3, row.names = FALSE, align = "c",
              format = "html", caption = "Model Comparison. (All data, n = 6155)") %>%
 kable_styling(font_size = 16, 
        bootstrap_options = c("striped", "hover", "condensed"),
        position = "center", full_width = FALSE) %>% 
  row_spec(6, bold = T, color = "white", background = "#D7261E") %>% 
  footnote(general = "Abbreviation: N, number; AIC, Akaike information criterion; BIC, Bayesian information criterion; aBIC, adjusted BIC; cAIC, consistent AIC; Entropy, a pseudo-r-squared index.")
```

From comparing output of different mixed effect LCA models in the above table, **we think a 6 class model may be considered as an appropriate model among them.**

## Visualisation of the 6-class model

```{r echo=FALSE, fig.height=5, fig.width=14, message=FALSE, warning=FALSE, cache=TRUE}

library(ggthemr)
rm(list=ls(all=TRUE))

ggthemr("fresh", layout = "scientific")

dt <- read.csv("../Tablecsv/ndns_param6.csv", header = T)

dt_long <- dt[-1,-c(1,2)] %>% 
  gather(Hour, Prob, -VARIABLE, -RESPCAT) %>% 
  separate(Hour, into = c("EST" ,"Class"), sep = "T")

dt_long <- dt_long %>% 
  filter(RESPCAT != 1) %>% 
  mutate(RESPCAT = as.factor(RESPCAT)) %>% 
  mutate(Carbo = revalue(RESPCAT, c("2" = "Low_carb", "3" = "Med_carb", "4" = "High_carb"))) %>% 
  mutate(HourN = as.character(VARIABLE)) %>% 
  mutate(HourN = factor(HourN, levels = c("H0","H1" ,"H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10","H11","H12","H13","H14","H15","H16","H17","H18","H19","H20","H21","H22","H23")))

ggthemr("fresh", layout = "scientific")

ggplot(dt_long[dt_long$Class == "LC1", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
  geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 1: 12.78%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(dt_long[dt_long$Class == "LC2", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 2:  24.73%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(dt_long[dt_long$Class == "LC3", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 3:  21.24%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(dt_long[dt_long$Class == "LC4", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 4:  23.08%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(dt_long[dt_long$Class == "LC5", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 5:  4.07%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))

ggplot(dt_long[dt_long$Class == "LC6", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 6:  14.10%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))
# 
# ggplot(a_long[a_long$Class == "Class7", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
#  geom_point(size = 3) +
#   geom_line(size = 1.5) +
#   theme(axis.title = element_text(size = 18),
#         axis.text = element_text(size = 18),
#         axis.line = element_line(colour = "black"),
#         panel.border = element_blank(),
#         panel.background = element_blank(),
#         legend.text = element_text(size = 15),
#         legend.title = element_text(size = 15)
#         # legend.position = "bottom",
#         # legend.direction = "horizontal"
#         ) +
#   labs(title = "Class 7:  12.17%", x = "Hour of the day", y = "Probability",
#        color = "Carbohydrate\nintake") +
#   scale_color_discrete(labels = c("Low", "Medium", "High")) +
#   ylim(c(0,1))
# 
# ggplot(a_long[a_long$Class == "Class8", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
#  geom_point(size = 3) +
#   geom_line(size = 1.5) +
#   theme(axis.title = element_text(size = 18),
#         axis.text = element_text(size = 18),
#         axis.line = element_line(colour = "black"),
#         panel.border = element_blank(),
#         panel.background = element_blank(),
#         legend.text = element_text(size = 15),
#         legend.title = element_text(size = 15)
#         # legend.position = "bottom",
#         # legend.direction = "horizontal"
#         ) +
#   labs(title = "Class 8:  12.36%", x = "Hour of the day", y = "Probability",
#        color = "Carbohydrate\nintake") +
#   scale_color_discrete(labels = c("Low", "Medium", "High")) +
#   ylim(c(0,1))
```

## Visualisation of the classes by Carbohydrate eating types

```{r echo=FALSE, fig.height=5, fig.width=14, message=FALSE, warning=FALSE, cache=TRUE}

ggthemr("fresh", layout = "scientific")

ggplot(dt_long[dt_long$Carbo == "Low_carb", ], aes(y = Prob, x=HourN, group = Class, color = Class)) + 
  geom_point(size = 2) + 
  geom_line(size = 1.2) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Low carbohydarte food consumption by people in different classes", x = "Hour of the day", y = "Probability", color = "Latent classes") + 
  scale_color_discrete(labels = c("1", "2", "3", "4", "5", "6")) + 
  ylim(c(0,1))

ggplot(dt_long[dt_long$Carbo == "Med_carb", ], aes(y = Prob, x=HourN, group = Class, color = Class)) + 
  geom_point(size = 2) + 
  geom_line(size = 1.2) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Medium carbohydarte food consumption by people in different classes", x = "Hour of the day", y = "Probability", color = "Latent classes") + 
  scale_color_discrete(labels = c("1", "2", "3", "4", "5", "6")) + 
  ylim(c(0,1))

ggplot(dt_long[dt_long$Carbo == "High_carb", ], aes(y = Prob, x=HourN, group = Class, color = Class)) + 
  geom_point(size = 2) + 
  geom_line(size = 1.2) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "High carbohydarte food consumption by people in different classes", x = "Hour of the day", y = "Probability", color = "Latent classes") + 
  scale_color_discrete(labels = c("1", "2", "3", "4", "5", "6")) + 
  ylim(c(0,1))
```

From this results, we might consider combining high-carbohydrate with the medium-carbohydrate category since in the third figure ("High carbohydrate consumption by latent classes") above, there is hardly any difference can be recognised. So, we tried the same analysis in the same data set combining medium and high carbohydrate consumptions together:

# Mixed effect LCA with carbohydrates intake groups redefined:

1. Not eating; 
2. Eating low carbohydrate food (energy contribution less than or equal to 25%); 
3. Eating medium-or-high carbohydrate food (energy contribution higher than 26%); 
4. ~~Eating high carbohydrate food (energy contribution higher or equal to 75%).~~

## Model comparison and selection

```{r echo=FALSE}
rm(list=ls(all=TRUE))
# load("~/Documents/LSHTMproject/Rcode/day4LCA1_6.Rdata")

library(kableExtra)
dt <- read.csv("../Tablecsv/ModelsMixedLOWORMEDIUM.csv", header = T)
names(dt) <- c("N of classes", "log-likelihood", "G^2",  "AIC", "BIC", "cAIC", "aBIC", "Entropy")
dt[1,8] <- "---"
kable(dt, digits = 3, row.names = FALSE, align = "c",
              format = "html", caption = "Model Comparison. (All data, n = 6155)") %>%
 kable_styling(font_size = 16, 
        bootstrap_options = c("striped", "hover", "condensed"),
        position = "center", full_width = FALSE) %>% 
  row_spec(c(3,4,6), bold = T, color = "white", background = "#D7261E") %>% 
  footnote(general = "Abbreviation: N, number; AIC, Akaike information criterion; BIC, Bayesian information criterion; aBIC, adjusted BIC; cAIC, consistent AIC; Entropy, a pseudo-r-squared index.")
```

## Visualisation of the 3-class model

```{r echo=FALSE, fig.height=5, fig.width=14, message=FALSE, warning=FALSE, cache=TRUE}

library(ggthemr)
rm(list=ls(all=TRUE))

ggthemr("fresh", layout = "scientific")

dt <- read.csv("../Tablecsv/ndns_param3_2cat.csv", header = T)

dt_long <- dt[-1,-c(1,2)] %>% 
  gather(Hour, Prob, -VARIABLE, -RESPCAT) %>% 
  separate(Hour, into = c("EST" ,"Class"), sep = "T")

dt_long <- dt_long %>% 
  filter(RESPCAT != 1) %>% 
  mutate(RESPCAT = as.factor(RESPCAT)) %>% 
  mutate(Carbo = revalue(RESPCAT, c("2" = "Low_carb", "3" = "Med_carb", "4" = "High_carb"))) %>% 
  mutate(HourN = as.character(VARIABLE)) %>% 
  mutate(HourN = factor(HourN, levels = c("H0","H1" ,"H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10","H11","H12","H13","H14","H15","H16","H17","H18","H19","H20","H21","H22","H23")))

ggthemr("fresh", layout = "scientific")

ggplot(dt_long[dt_long$Class == "LC1", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
  geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 1: 70.85%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(dt_long[dt_long$Class == "LC2", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 2:  24.72%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(dt_long[dt_long$Class == "LC3", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 3:  4.43%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))

# 
# ggplot(dt_long[dt_long$Class == "LC4", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
#  geom_point(size = 3) + 
#   geom_line(size = 1.5) + 
#   theme(axis.title = element_text(size = 18), 
#         axis.text = element_text(size = 18), 
#         axis.line = element_line(colour = "black"), 
#         panel.border = element_blank(), 
#         panel.background = element_blank(), 
#         legend.text = element_text(size = 15), 
#         legend.title = element_text(size = 15) 
#         # legend.position = "bottom", 
#         # legend.direction = "horizontal"
#         ) + 
#   labs(title = "Class 4:  23.08%", x = "Hour of the day", y = "Probability",
#        color = "Carbohydrate\nintake") + 
#   scale_color_discrete(labels = c("Low", "Medium", "High")) + 
#   ylim(c(0,1))
# 
# 
# ggplot(dt_long[dt_long$Class == "LC5", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
#  geom_point(size = 3) +
#   geom_line(size = 1.5) +
#   theme(axis.title = element_text(size = 18),
#         axis.text = element_text(size = 18),
#         axis.line = element_line(colour = "black"),
#         panel.border = element_blank(),
#         panel.background = element_blank(),
#         legend.text = element_text(size = 15),
#         legend.title = element_text(size = 15)
#         # legend.position = "bottom",
#         # legend.direction = "horizontal"
#         ) +
#   labs(title = "Class 5:  4.07%", x = "Hour of the day", y = "Probability",
#        color = "Carbohydrate\nintake") +
#   scale_color_discrete(labels = c("Low", "Medium", "High")) +
#   ylim(c(0,1))
# 
# ggplot(dt_long[dt_long$Class == "LC6", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
#  geom_point(size = 3) +
#   geom_line(size = 1.5) +
#   theme(axis.title = element_text(size = 18),
#         axis.text = element_text(size = 18),
#         axis.line = element_line(colour = "black"),
#         panel.border = element_blank(),
#         panel.background = element_blank(),
#         legend.text = element_text(size = 15),
#         legend.title = element_text(size = 15)
#         # legend.position = "bottom",
#         # legend.direction = "horizontal"
#         ) +
#   labs(title = "Class 6:  14.10%", x = "Hour of the day", y = "Probability",
#        color = "Carbohydrate\nintake") +
#   scale_color_discrete(labels = c("Low", "Medium", "High")) +
#   ylim(c(0,1))
```

## Visualisation of the 4-class model

```{r echo=FALSE, fig.height=5, fig.width=14, message=FALSE, warning=FALSE, cache=TRUE}

library(ggthemr)
rm(list=ls(all=TRUE))

ggthemr("fresh", layout = "scientific")

dt <- read.csv("../Tablecsv/ndns_param4_2cat.csv", header = T)

dt_long <- dt[-1,-c(1,2)] %>% 
  gather(Hour, Prob, -VARIABLE, -RESPCAT) %>% 
  separate(Hour, into = c("EST" ,"Class"), sep = "T")

dt_long <- dt_long %>% 
  filter(RESPCAT != 1) %>% 
  mutate(RESPCAT = as.factor(RESPCAT)) %>% 
  mutate(Carbo = revalue(RESPCAT, c("2" = "Low_carb", "3" = "Med_carb", "4" = "High_carb"))) %>% 
  mutate(HourN = as.character(VARIABLE)) %>% 
  mutate(HourN = factor(HourN, levels = c("H0","H1" ,"H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10","H11","H12","H13","H14","H15","H16","H17","H18","H19","H20","H21","H22","H23")))

ggthemr("fresh", layout = "scientific")

ggplot(dt_long[dt_long$Class == "LC1", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
  geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 1: 43.24%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(dt_long[dt_long$Class == "LC2", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 2:  24.67%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(dt_long[dt_long$Class == "LC3", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 3:  4.59%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))

ggplot(dt_long[dt_long$Class == "LC4", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 4:  27.49%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))

# 
# ggplot(dt_long[dt_long$Class == "LC5", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
#  geom_point(size = 3) +
#   geom_line(size = 1.5) +
#   theme(axis.title = element_text(size = 18),
#         axis.text = element_text(size = 18),
#         axis.line = element_line(colour = "black"),
#         panel.border = element_blank(),
#         panel.background = element_blank(),
#         legend.text = element_text(size = 15),
#         legend.title = element_text(size = 15)
#         # legend.position = "bottom",
#         # legend.direction = "horizontal"
#         ) +
#   labs(title = "Class 5:  4.07%", x = "Hour of the day", y = "Probability",
#        color = "Carbohydrate\nintake") +
#   scale_color_discrete(labels = c("Low", "Medium", "High")) +
#   ylim(c(0,1))
# 
# ggplot(dt_long[dt_long$Class == "LC6", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
#  geom_point(size = 3) +
#   geom_line(size = 1.5) +
#   theme(axis.title = element_text(size = 18),
#         axis.text = element_text(size = 18),
#         axis.line = element_line(colour = "black"),
#         panel.border = element_blank(),
#         panel.background = element_blank(),
#         legend.text = element_text(size = 15),
#         legend.title = element_text(size = 15)
#         # legend.position = "bottom",
#         # legend.direction = "horizontal"
#         ) +
#   labs(title = "Class 6:  14.10%", x = "Hour of the day", y = "Probability",
#        color = "Carbohydrate\nintake") +
#   scale_color_discrete(labels = c("Low", "Medium", "High")) +
#   ylim(c(0,1))
```

## Visualisation of the 5-class model

```{r echo=FALSE, fig.height=5, fig.width=14, message=FALSE, warning=FALSE, cache=TRUE}

library(ggthemr)
rm(list=ls(all=TRUE))

ggthemr("fresh", layout = "scientific")

dt <- read.csv("../Tablecsv/ndns_param5_2cat.csv", header = T)

dt_long <- dt[-1,-c(1,2)] %>% 
  gather(Hour, Prob, -VARIABLE, -RESPCAT) %>% 
  separate(Hour, into = c("EST" ,"Class"), sep = "T")

dt_long <- dt_long %>% 
  filter(RESPCAT != 1) %>% 
  mutate(RESPCAT = as.factor(RESPCAT)) %>% 
  mutate(Carbo = revalue(RESPCAT, c("2" = "Low_carb", "3" = "Med_carb", "4" = "High_carb"))) %>% 
  mutate(HourN = as.character(VARIABLE)) %>% 
  mutate(HourN = factor(HourN, levels = c("H0","H1" ,"H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10","H11","H12","H13","H14","H15","H16","H17","H18","H19","H20","H21","H22","H23")))

ggthemr("fresh", layout = "scientific")

ggplot(dt_long[dt_long$Class == "LC1", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
  geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 1: 29.13%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(dt_long[dt_long$Class == "LC2", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 2:  24.67%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(dt_long[dt_long$Class == "LC3", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 3:  4.10%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))

ggplot(dt_long[dt_long$Class == "LC4", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 4:  22.53%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))

ggplot(dt_long[dt_long$Class == "LC5", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 5:  19.58%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))
# 
# ggplot(dt_long[dt_long$Class == "LC6", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
#  geom_point(size = 3) +
#   geom_line(size = 1.5) +
#   theme(axis.title = element_text(size = 18),
#         axis.text = element_text(size = 18),
#         axis.line = element_line(colour = "black"),
#         panel.border = element_blank(),
#         panel.background = element_blank(),
#         legend.text = element_text(size = 15),
#         legend.title = element_text(size = 15)
#         # legend.position = "bottom",
#         # legend.direction = "horizontal"
#         ) +
#   labs(title = "Class 6:  14.10%", x = "Hour of the day", y = "Probability",
#        color = "Carbohydrate\nintake") +
#   scale_color_discrete(labels = c("Low", "Medium", "High")) +
#   ylim(c(0,1))
```

## Visualisation of the 6-class model

```{r echo=FALSE, fig.height=5, fig.width=14, message=FALSE, warning=FALSE, cache=TRUE}

library(ggthemr)
rm(list=ls(all=TRUE))

ggthemr("fresh", layout = "scientific")

dt <- read.csv("../Tablecsv/ndns_param6_2cat.csv", header = T)

dt_long <- dt[-1,-c(1,2)] %>% 
  gather(Hour, Prob, -VARIABLE, -RESPCAT) %>% 
  separate(Hour, into = c("EST" ,"Class"), sep = "T")

dt_long <- dt_long %>% 
  filter(RESPCAT != 1) %>% 
  mutate(RESPCAT = as.factor(RESPCAT)) %>% 
  mutate(Carbo = revalue(RESPCAT, c("2" = "Low_carb", "3" = "Med_carb", "4" = "High_carb"))) %>% 
  mutate(HourN = as.character(VARIABLE)) %>% 
  mutate(HourN = factor(HourN, levels = c("H0","H1" ,"H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10","H11","H12","H13","H14","H15","H16","H17","H18","H19","H20","H21","H22","H23")))

ggthemr("fresh", layout = "scientific")

ggplot(dt_long[dt_long$Class == "LC1", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
  geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 1: 22.29%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(dt_long[dt_long$Class == "LC2", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) + 
 geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
        ) + 
  labs(title = "Class 2:  24.61%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Low", "Medium", "High")) + 
  ylim(c(0,1))

ggplot(dt_long[dt_long$Class == "LC3", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 3:  3.94%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))

ggplot(dt_long[dt_long$Class == "LC4", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 4:  22.31%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))

ggplot(dt_long[dt_long$Class == "LC5", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 5:  15.98%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))

ggplot(dt_long[dt_long$Class == "LC6", ], aes(y = Prob, x=HourN, group = Carbo, color = Carbo)) +
 geom_point(size = 3) +
  geom_line(size = 1.5) +
  theme(axis.title = element_text(size = 18),
        axis.text = element_text(size = 18),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)
        # legend.position = "bottom",
        # legend.direction = "horizontal"
        ) +
  labs(title = "Class 6:  10.86%", x = "Hour of the day", y = "Probability",
       color = "Carbohydrate\nintake") +
  scale_color_discrete(labels = c("Low", "Medium", "High")) +
  ylim(c(0,1))
```