### 2018-07-09 paper/Appendices/AppendixA.tex

% Appendix A

\chapter{R code for importing and manipulating the data} % Main appendix title

\label{AppendixA} % For referencing this appendix elsewhere, use \ref{AppendixA}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# NDNS analysis, data management ------------------------------------------}

\CommentTok{# Change the data path accordingly ----------------------------------------}
\KeywordTok{setwd}\NormalTok{(}\StringTok{"/home/wangcc-me/Downloads/UKDA-6533-stata11_se/stata11_se/"}\NormalTok{) }\CommentTok{# in Ubuntu}
\KeywordTok{library}\NormalTok{(epiDisplay)}
\KeywordTok{library}\NormalTok{(plyr)}
\KeywordTok{library}\NormalTok{(tidyverse)}

\CommentTok{# Read the data into memory -----------------------------------------------}
\KeywordTok{library}\NormalTok{(haven)}
\NormalTok{data <-}\StringTok{ }\KeywordTok{read_dta}\NormalTok{(}\StringTok{"ndns_rp_yr1-4a_foodleveldietarydata_uk.dta"}\NormalTok{)}
\NormalTok{data56 <-}\StringTok{ }\KeywordTok{read_dta}\NormalTok{(}\StringTok{"ndns_rp_yr5-6a_foodleveldietarydata.dta"}\NormalTok{)}
\NormalTok{data78 <-}\StringTok{ }\KeywordTok{read_dta}\NormalTok{(}\StringTok{"ndns_rp_yr7-8a_foodleveldietarydata.dta"}\NormalTok{)}
\KeywordTok{names}\NormalTok{(data)}
\KeywordTok{names}\NormalTok{(data56)}
\KeywordTok{names}\NormalTok{(data78)}
\KeywordTok{names}\NormalTok{(data)[}\KeywordTok{names}\NormalTok{(data) }\OperatorTok{==}\StringTok{ "seriali"}\NormalTok{] <-}\StringTok{ "id"}
\KeywordTok{names}\NormalTok{(data56)[}\KeywordTok{names}\NormalTok{(data56) }\OperatorTok{==}\StringTok{ "seriali"}\NormalTok{] <-}\StringTok{ "id"}
\KeywordTok{names}\NormalTok{(data78)[}\KeywordTok{names}\NormalTok{(data78) }\OperatorTok{==}\StringTok{ "seriali"}\NormalTok{] <-}\StringTok{ "id"}

\CommentTok{# Extract the data we needed ----------------------------------------------}
\NormalTok{df14d <-}\StringTok{ }\NormalTok{data[, }\KeywordTok{c}\NormalTok{(}\DecValTok{113}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{, }\DecValTok{3}\NormalTok{, }\DecValTok{5}\NormalTok{, }\DecValTok{6}\NormalTok{, }\DecValTok{7}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{9}\NormalTok{, }\DecValTok{21}\NormalTok{, }\DecValTok{24}\NormalTok{, }\DecValTok{55}\NormalTok{, }\DecValTok{57}\NormalTok{, }\DecValTok{58}\NormalTok{, }
    \DecValTok{59}\NormalTok{, }\DecValTok{60}\NormalTok{, }\DecValTok{61}\NormalTok{, }\DecValTok{62}\NormalTok{, }\DecValTok{63}\NormalTok{, }\DecValTok{64}\NormalTok{)]}
\NormalTok{var <-}\StringTok{ }\KeywordTok{names}\NormalTok{(df14d)}
\NormalTok{df56d <-}\StringTok{ }\NormalTok{data56 }\OperatorTok{%>%}\StringTok{ }\KeywordTok{select}\NormalTok{(var)}
\NormalTok{df78d <-}\StringTok{ }\NormalTok{data78 }\OperatorTok{%>%}\StringTok{ }\KeywordTok{select}\NormalTok{(var)}
\NormalTok{dfs1 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(df14d, df56d, df78d)}
\NormalTok{dfs2 <-}\StringTok{ }\NormalTok{dfs1[dfs1}\OperatorTok{$}\NormalTok{Age }\OperatorTok{>=}\StringTok{ }\DecValTok{19}\NormalTok{, ]}
\KeywordTok{rm}\NormalTok{(data, data56, data78)}
\NormalTok{dfs2}

\CommentTok{# Calculate the time (minute and hour) when they eat ----------------------}

\NormalTok{dfs2}\OperatorTok{$}\NormalTok{MealTime_chr <-}\StringTok{ }\KeywordTok{as.character}\NormalTok{(dfs2}\OperatorTok{$}\NormalTok{MealTime)}
\NormalTok{dfs2}\OperatorTok{$}\NormalTok{MealTime_hm <-}\StringTok{ }\KeywordTok{unlist}\NormalTok{(}\KeywordTok{strsplit}\NormalTok{(dfs2}\OperatorTok{$}\NormalTok{MealTime_chr, }\StringTok{" "}\NormalTok{))[}\KeywordTok{c}\NormalTok{(}\OtherTok{FALSE}\NormalTok{, }
    \OtherTok{TRUE}\NormalTok{)]}
\NormalTok{dfs2}\OperatorTok{$}\NormalTok{MealHourN <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{(}\KeywordTok{unlist}\NormalTok{(}\KeywordTok{strsplit}\NormalTok{(dfs2}\OperatorTok{$}\NormalTok{MealTime_hm, }\StringTok{":"}\NormalTok{))[}\KeywordTok{c}\NormalTok{(}\OtherTok{TRUE}\NormalTok{, }
    \OtherTok{FALSE}\NormalTok{, }\OtherTok{FALSE}\NormalTok{)])}
\NormalTok{dfs2}\OperatorTok{$}\NormalTok{MealMinN <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{(}\KeywordTok{unlist}\NormalTok{(}\KeywordTok{strsplit}\NormalTok{(dfs2}\OperatorTok{$}\NormalTok{MealTime_hm, }\StringTok{":"}\NormalTok{))[}\KeywordTok{c}\NormalTok{(}\OtherTok{FALSE}\NormalTok{, }
    \OtherTok{TRUE}\NormalTok{, }\OtherTok{FALSE}\NormalTok{)])}
\NormalTok{dfs2}\OperatorTok{$}\NormalTok{MealMinN0 <-}\StringTok{ }\NormalTok{(}\DecValTok{60} \OperatorTok{*}\StringTok{ }\NormalTok{dfs2}\OperatorTok{$}\NormalTok{MealHourN) }\OperatorTok{+}\StringTok{ }\NormalTok{dfs2}\OperatorTok{$}\NormalTok{MealMinN}
\NormalTok{dfs3 <-}\StringTok{ }\NormalTok{dfs2[}\KeywordTok{order}\NormalTok{(dfs2}\OperatorTok{$}\NormalTok{id, dfs2}\OperatorTok{$}\NormalTok{DayNo, dfs2}\OperatorTok{$}\NormalTok{MealMinN0), ]}
\KeywordTok{length}\NormalTok{(}\KeywordTok{unique}\NormalTok{(dfs3}\OperatorTok{$}\NormalTok{id))  ## number of participants = 6155}

\CommentTok{# Create a subset data with only the first observation of each}
\CommentTok{# participant --------}
\NormalTok{NDNS <-}\StringTok{ }\NormalTok{dfs3[}\OperatorTok{!}\KeywordTok{duplicated}\NormalTok{(dfs3}\OperatorTok{$}\NormalTok{id), ]}
\KeywordTok{with}\NormalTok{(NDNS, }\KeywordTok{tab1}\NormalTok{(SurveyYear, }\DataTypeTok{graph =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{decimal =} \DecValTok{2}\NormalTok{))}

\CommentTok{# #SurveyYear :}
            \CommentTok{# Frequency Percent Cum. percent}
\CommentTok{# NDNS Year 1       801   13.01        13.01}
\CommentTok{# NDNS Year 2       812   13.19        26.21}
\CommentTok{# NDNS Year 3       782   12.71        38.91}
\CommentTok{# NDNS Year 4      1055   17.14        56.05}
\CommentTok{# NDNS Year 5       625   10.15        66.21}
\CommentTok{# NDNS Year 6       663   10.77        76.98}
\CommentTok{# NDNS Year 7       703   11.42        88.40}
\CommentTok{# NDNS Year 8       714   11.60       100.00}
  \CommentTok{# Total          6155  100.00       100.00}

\CommentTok{# create a variable combine id and day No ---------------------------------}
\NormalTok{dfs3 <-}\StringTok{ }\NormalTok{dfs3 }\OperatorTok{%>%}
\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{id_dy =} \KeywordTok{paste}\NormalTok{(id, DayNo, }\DataTypeTok{sep =} \StringTok{"D"}\NormalTok{))}

\CommentTok{# For each subject, the total energy/carbohydrate intake for each eating }
\CommentTok{# time can be calculated --------}
\NormalTok{old <-}\StringTok{ }\KeywordTok{Sys.time}\NormalTok{()}
\NormalTok{Energy <-}\StringTok{ }\KeywordTok{ddply}\NormalTok{(dfs3, .(id_dy, id, SurveyYear, DayNo, Age, Sex, }
\NormalTok{                        DiaryDaysCompleted, MealHourN, DayofWeek), }
\NormalTok{                summarise, }
                \DataTypeTok{Tot_Energ =} \KeywordTok{sum}\NormalTok{(EnergykJ), }
                \DataTypeTok{Tot_Carb =} \KeywordTok{sum}\NormalTok{(Carbohydrateg), }
                \DataTypeTok{Tot_Sugar =} \KeywordTok{sum}\NormalTok{(Totalsugarsg), }
                \DataTypeTok{Tot_Starch =} \KeywordTok{sum}\NormalTok{(Starchg))}
\NormalTok{new <-}\StringTok{ }\KeywordTok{Sys.time}\NormalTok{() }\OperatorTok{-}\StringTok{ }\NormalTok{old}
\KeywordTok{print}\NormalTok{(new)}
\CommentTok{# Time difference of 3.876385 mins}

\KeywordTok{rm}\NormalTok{(df14d, df56d, df78d, dfs2)}

\CommentTok{# Calculate the energy from total carbohydrates ---------------------------}
\NormalTok{Energy <-}\StringTok{ }\NormalTok{Energy }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{KJcarbo =}\NormalTok{ Tot_Carb }\OperatorTok{*}\StringTok{ }\DecValTok{16}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{CarKJpercentage =}\NormalTok{ KJcarbo}\OperatorTok{/}\NormalTok{Tot_Energ) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{Carbo =} \KeywordTok{cut}\NormalTok{(CarKJpercentage, }\DataTypeTok{breaks =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\FloatTok{0.26}\NormalTok{, }\FloatTok{0.75}\NormalTok{, }\DecValTok{2}\NormalTok{), }
        \DataTypeTok{right =} \OtherTok{FALSE}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{Carbo2 =} \KeywordTok{cut}\NormalTok{(CarKJpercentage, }\DataTypeTok{breaks =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }
    \FloatTok{0.26}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{right =} \OtherTok{FALSE}\NormalTok{))}
\NormalTok{Energy0 <-}\StringTok{ }\NormalTok{Energy[}\OperatorTok{!}\NormalTok{(Energy}\OperatorTok{$}\NormalTok{Tot_Energ }\OperatorTok{==}\StringTok{ }\DecValTok{0}\NormalTok{), ]  }
          \CommentTok{# some food consumption does not contain any carbohydrates}
\NormalTok{Energy0}\OperatorTok{$}\NormalTok{Carbo <-}\StringTok{ }\KeywordTok{factor}\NormalTok{(Energy0}\OperatorTok{$}\NormalTok{Carbo, }\DataTypeTok{labels =} \KeywordTok{c}\NormalTok{(}\StringTok{"Low_carb"}\NormalTok{, }\StringTok{"Med_carb"}\NormalTok{, }
    \StringTok{"High_carb"}\NormalTok{))}
\NormalTok{Energy0}\OperatorTok{$}\NormalTok{Carbo2 <-}\StringTok{ }\KeywordTok{factor}\NormalTok{(Energy0}\OperatorTok{$}\NormalTok{Carbo2, }\DataTypeTok{labels =} \KeywordTok{c}\NormalTok{(}\StringTok{"Low_carb"}\NormalTok{, }\StringTok{"Med_or_high_carb"}\NormalTok{))}

\CommentTok{# Generate data sets for each day ----------------------------------------------------------}
\NormalTok{dta_day1 <-}\StringTok{ }\NormalTok{Energy0 }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(DayNo }\OperatorTok{==}\StringTok{ }\DecValTok{1}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"id"}\NormalTok{, }\StringTok{"Age"}\NormalTok{, }
    \StringTok{"Sex"}\NormalTok{, }\StringTok{"DayofWeek"}\NormalTok{, }\StringTok{"MealHourN"}\NormalTok{, }\StringTok{"Carbo"}\NormalTok{, }\StringTok{"Carbo2"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{DayofWeek =} \KeywordTok{factor}\NormalTok{(DayofWeek, }
    \DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{"Monday"}\NormalTok{, }\StringTok{"Tuesday"}\NormalTok{, }\StringTok{"Wednesday"}\NormalTok{, }\StringTok{"Thursday"}\NormalTok{, }\StringTok{"Friday"}\NormalTok{, }
        \StringTok{"Saturday"}\NormalTok{, }\StringTok{"Sunday"}\NormalTok{)))}

\NormalTok{dta_day2 <-}\StringTok{ }\NormalTok{Energy0 }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(DayNo }\OperatorTok{==}\StringTok{ }\DecValTok{2}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"id"}\NormalTok{, }\StringTok{"Age"}\NormalTok{, }
    \StringTok{"Sex"}\NormalTok{, }\StringTok{"DayofWeek"}\NormalTok{, }\StringTok{"MealHourN"}\NormalTok{, }\StringTok{"Carbo"}\NormalTok{, }\StringTok{"Carbo2"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{DayofWeek =} \KeywordTok{factor}\NormalTok{(DayofWeek, }
    \DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{"Monday"}\NormalTok{, }\StringTok{"Tuesday"}\NormalTok{, }\StringTok{"Wednesday"}\NormalTok{, }\StringTok{"Thursday"}\NormalTok{, }\StringTok{"Friday"}\NormalTok{, }
        \StringTok{"Saturday"}\NormalTok{, }\StringTok{"Sunday"}\NormalTok{)))}

\NormalTok{dta_day3 <-}\StringTok{ }\NormalTok{Energy0 }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(DayNo }\OperatorTok{==}\StringTok{ }\DecValTok{3}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"id"}\NormalTok{, }\StringTok{"Age"}\NormalTok{, }
    \StringTok{"Sex"}\NormalTok{, }\StringTok{"DayofWeek"}\NormalTok{, }\StringTok{"MealHourN"}\NormalTok{, }\StringTok{"Carbo"}\NormalTok{, }\StringTok{"Carbo2"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{DayofWeek =} \KeywordTok{factor}\NormalTok{(DayofWeek, }
    \DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{"Monday"}\NormalTok{, }\StringTok{"Tuesday"}\NormalTok{, }\StringTok{"Wednesday"}\NormalTok{, }\StringTok{"Thursday"}\NormalTok{, }\StringTok{"Friday"}\NormalTok{, }
        \StringTok{"Saturday"}\NormalTok{, }\StringTok{"Sunday"}\NormalTok{)))}

\NormalTok{dta_day4 <-}\StringTok{ }\NormalTok{Energy0 }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(DayNo }\OperatorTok{==}\StringTok{ }\DecValTok{4}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"id"}\NormalTok{, }\StringTok{"Age"}\NormalTok{, }
    \StringTok{"Sex"}\NormalTok{, }\StringTok{"DayofWeek"}\NormalTok{, }\StringTok{"MealHourN"}\NormalTok{, }\StringTok{"Carbo"}\NormalTok{, }\StringTok{"Carbo2"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{DayofWeek =} \KeywordTok{factor}\NormalTok{(DayofWeek, }
    \DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{"Monday"}\NormalTok{, }\StringTok{"Tuesday"}\NormalTok{, }\StringTok{"Wednesday"}\NormalTok{, }\StringTok{"Thursday"}\NormalTok{, }\StringTok{"Friday"}\NormalTok{, }
        \StringTok{"Saturday"}\NormalTok{, }\StringTok{"Sunday"}\NormalTok{)))}
\NormalTok{vecid1 <-}\StringTok{ }\KeywordTok{unique}\NormalTok{(dta_day1}\OperatorTok{$}\NormalTok{id) }\CommentTok{# n = 6153}
\NormalTok{vecid2 <-}\StringTok{ }\KeywordTok{unique}\NormalTok{(dta_day2}\OperatorTok{$}\NormalTok{id) }\CommentTok{# n = 6153}
\NormalTok{vecid3 <-}\StringTok{ }\KeywordTok{unique}\NormalTok{(dta_day3}\OperatorTok{$}\NormalTok{id) }\CommentTok{# n = 6151}
\NormalTok{vecid4 <-}\StringTok{ }\KeywordTok{unique}\NormalTok{(dta_day4}\OperatorTok{$}\NormalTok{id) }\CommentTok{# n = 6026}

\NormalTok{Noday1 <-}\StringTok{ }\KeywordTok{setdiff}\NormalTok{(vecid, vecid1)  }\CommentTok{# two subjects did not have day 1 data}
\NormalTok{Noday2 <-}\StringTok{ }\KeywordTok{setdiff}\NormalTok{(vecid, vecid2)  }\CommentTok{# two subjects did not have day 2 data}
\NormalTok{Noday3 <-}\StringTok{ }\KeywordTok{setdiff}\NormalTok{(vecid, vecid3)  }\CommentTok{# four subjects did not have day 3 data}
\NormalTok{Noday4 <-}\StringTok{ }\KeywordTok{setdiff}\NormalTok{(vecid, vecid4)  }\CommentTok{# 129 subjects did not have day 4 data}

\CommentTok{# Transform the data shape from long to wide ---------------------------------}
\NormalTok{dta_d1_wide <-}\StringTok{ }\NormalTok{dta_day1[, }\OperatorTok{-}\DecValTok{7}\NormalTok{] }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{spread}\NormalTok{(}\DataTypeTok{key =}\NormalTok{ MealHourN, }\DataTypeTok{value =}\NormalTok{ Carbo)}
\KeywordTok{names}\NormalTok{(dta_d1_wide)[}\DecValTok{5}\OperatorTok{:}\DecValTok{28}\NormalTok{] <-}\StringTok{ }\KeywordTok{paste}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\StringTok{"H"}\NormalTok{, }\DecValTok{24}\NormalTok{), }\DecValTok{0}\OperatorTok{:}\DecValTok{23}\NormalTok{, }\DataTypeTok{sep =} \StringTok{""}\NormalTok{)}

\NormalTok{dta_d2_wide <-}\StringTok{ }\NormalTok{dta_day2[, }\OperatorTok{-}\DecValTok{7}\NormalTok{] }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{spread}\NormalTok{(}\DataTypeTok{key =}\NormalTok{ MealHourN, }\DataTypeTok{value =}\NormalTok{ Carbo)}
\KeywordTok{names}\NormalTok{(dta_d2_wide)[}\DecValTok{5}\OperatorTok{:}\DecValTok{28}\NormalTok{] <-}\StringTok{ }\KeywordTok{paste}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\StringTok{"H"}\NormalTok{, }\DecValTok{24}\NormalTok{), }\DecValTok{0}\OperatorTok{:}\DecValTok{23}\NormalTok{, }\DataTypeTok{sep =} \StringTok{""}\NormalTok{)}

\NormalTok{dta_d3_wide <-}\StringTok{ }\NormalTok{dta_day3[, }\OperatorTok{-}\DecValTok{7}\NormalTok{] }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{spread}\NormalTok{(}\DataTypeTok{key =}\NormalTok{ MealHourN, }\DataTypeTok{value =}\NormalTok{ Carbo)}
\KeywordTok{names}\NormalTok{(dta_d3_wide)[}\DecValTok{5}\OperatorTok{:}\DecValTok{28}\NormalTok{] <-}\StringTok{ }\KeywordTok{paste}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\StringTok{"H"}\NormalTok{, }\DecValTok{24}\NormalTok{), }\DecValTok{0}\OperatorTok{:}\DecValTok{23}\NormalTok{, }\DataTypeTok{sep =} \StringTok{""}\NormalTok{)}

\NormalTok{dta_d4_wide <-}\StringTok{ }\NormalTok{dta_day4[, }\OperatorTok{-}\DecValTok{7}\NormalTok{] }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{spread}\NormalTok{(}\DataTypeTok{key =}\NormalTok{ MealHourN, }\DataTypeTok{value =}\NormalTok{ Carbo)}
\KeywordTok{names}\NormalTok{(dta_d4_wide)[}\DecValTok{5}\OperatorTok{:}\DecValTok{28}\NormalTok{] <-}\StringTok{ }\KeywordTok{paste}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\StringTok{"H"}\NormalTok{, }\DecValTok{24}\NormalTok{), }\DecValTok{0}\OperatorTok{:}\DecValTok{23}\NormalTok{, }\DataTypeTok{sep =} \StringTok{""}\NormalTok{)}

\CommentTok{# recode NA to not eating -------------------------------------------------}
\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{5}\OperatorTok{:}\KeywordTok{ncol}\NormalTok{(dta_d1_wide)) }
  \ControlFlowTok{if}\NormalTok{ (}\KeywordTok{is.factor}\NormalTok{(dta_d1_wide[, i])) }\KeywordTok{levels}\NormalTok{(dta_d1_wide[, }
\NormalTok{    i]) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\KeywordTok{levels}\NormalTok{(dta_d1_wide[, i]), }\StringTok{"Not_eating"}\NormalTok{)}

\NormalTok{dta_d1_wide[}\KeywordTok{is.na}\NormalTok{(dta_d1_wide)] <-}\StringTok{ "Not_eating"}

\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{5}\OperatorTok{:}\KeywordTok{ncol}\NormalTok{(dta_d2_wide)) }
  \ControlFlowTok{if}\NormalTok{ (}\KeywordTok{is.factor}\NormalTok{(dta_d2_wide[, i])) }\KeywordTok{levels}\NormalTok{(dta_d2_wide[, }
\NormalTok{    i]) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\KeywordTok{levels}\NormalTok{(dta_d2_wide[, i]), }\StringTok{"Not_eating"}\NormalTok{)}

\NormalTok{dta_d2_wide[}\KeywordTok{is.na}\NormalTok{(dta_d2_wide)] <-}\StringTok{ "Not_eating"}

\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{5}\OperatorTok{:}\KeywordTok{ncol}\NormalTok{(dta_d3_wide)) }
  \ControlFlowTok{if}\NormalTok{ (}\KeywordTok{is.factor}\NormalTok{(dta_d3_wide[, i])) }\KeywordTok{levels}\NormalTok{(dta_d3_wide[, }
\NormalTok{    i]) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\KeywordTok{levels}\NormalTok{(dta_d3_wide[, i]), }\StringTok{"Not_eating"}\NormalTok{)}

\NormalTok{dta_d3_wide[}\KeywordTok{is.na}\NormalTok{(dta_d3_wide)] <-}\StringTok{ "Not_eating"}

\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{5}\OperatorTok{:}\KeywordTok{ncol}\NormalTok{(dta_d4_wide)) }
  \ControlFlowTok{if}\NormalTok{ (}\KeywordTok{is.factor}\NormalTok{(dta_d4_wide[, i])) }\KeywordTok{levels}\NormalTok{(dta_d4_wide[, }
\NormalTok{    i]) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\KeywordTok{levels}\NormalTok{(dta_d4_wide[, i]), }\StringTok{"Not_eating"}\NormalTok{)}

\NormalTok{dta_d4_wide[}\KeywordTok{is.na}\NormalTok{(dta_d4_wide)] <-}\StringTok{ "Not_eating"}
\end{Highlighting}
\end{Shaded}

%\section{How do I change the colors of links?}
%
%The color of links can be changed to your liking using:
%
%{\small\verb!\hypersetup{urlcolor=red}!}, or
%
%{\small\verb!\hypersetup{citecolor=green}!}, or
%
%{\small\verb!\hypersetup{allcolor=blue}!}.
%
%\noindent If you want to completely hide the links, you can use:
%
%{\small\verb!\hypersetup{allcolors=.}!}, or even better:
%
%{\small\verb!\hypersetup{hidelinks}!}.
%
%\noindent If you want to have obvious links in the PDF but not %the printed text, use:
%
%{\small\verb!\hypersetup{colorlinks=false}!}.