### 2018-07-28 results/Summary of survey analysis.Rmd
table 1

---
title: "Summary of Survey Analysis"
author: "Chaochen Wang"
date: "27 July 2018"
output:
  html_document:
    nature:
      highlightLines: yes
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(RColorBrewer)
library(plyr)
library(tidyverse)
library(dplyr)
library(epiDisplay)
library(scales)
library(readr)
library(kableExtra)
```

# Multilevel classes results (2by2)

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
dt <- read.csv("/home/wangcc-me/Documents/LSHTMproject/Tablecsv/CW2CB2results.csv", header = T, row.names = NULL, colClasses = "character")
names(dt) <- c("Variables", "Latent class = 1", "Latent class = 2", "P value [note]")

dt[28, 1] <- c("Hypertension (%) [note]")
dt[32, 1] <- c("Carbohydrate percent (%) [note]")
dt[35, 1] <- c("DM [note]")

kable(dt,  align = "l", escape = F, row.names = FALSE,
              format = "html", caption = "Means, percentages, and 95% CIs of the characteristics by latent class memberships in the UK adults (NDNS RP 2008/09-2015/16, sample size = 6155).") %>%
 kable_styling(font_size = 16, 
        bootstrap_options = c("striped", "hover", "condensed"),
        position = "center", full_width = FALSE) %>% 
  add_indent(c(3:6, 9, 10, 12:19, 23:25, 27, 29)) %>% 
  footnote(general = c("Abbreviations: CI, confidence intervals; NDNS RP, national dietary and nutrition survey rolling programme; BMI body mass index; WC, waist circumference; A1C, haemoglobin A1c; DM, diabetes mellitus.", 
                       "Variables from the blood tests (glucose and A1C) are weighted by blood sample weights, the others are weighted by individual weights.", 
                       "Glucose and A1C levels are estimated in subgroups of people without diabetes.")) %>% 
  add_footnote(c("For continuous variables, the F test was used to determine differences between latent classes. For categorical variables, differences between latent classes were assessed using the adjusted Pearson Chi-2 test for survey data.",
                 "Hypertension was defined as either systolic blood pressure >= 140 mmHg or diastolic blood pressure >= 90 mmHg, or under treatment for hypertension.", 
                 "Carbohydrate percent indicates the percentage of energy from carbohydrate in total energy intake",
                 "DM was defined by A1C > 6.5%."), 
               notation = "symbol")
```

# Multilevel classes results (3by2)

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
dt <- read.csv("/home/wangcc-me/Documents/LSHTMproject/Tablecsv/CW3CB2results.csv", header = T, row.names = NULL, colClasses = "character")
names(dt) <- c("Variables", "Latent class = 1", "Latent class = 2", "P value [note]")

dt[28, 1] <- c("Hypertension (%) [note]")
dt[32, 1] <- c("Carbohydrate percent (%) [note]")
dt[35, 1] <- c("DM [note]")

kable(dt,  align = "l", escape = F, row.names = FALSE,
              format = "html", caption = "Means, percentages, and 95% CIs of the characteristics by latent class memberships in the UK adults (NDNS RP 2008/09-2015/16, sample size = 6155).") %>%
 kable_styling(font_size = 16, 
        bootstrap_options = c("striped", "hover", "condensed"),
        position = "center", full_width = FALSE) %>% 
  add_indent(c(3:6, 9, 10, 12:19, 23:25, 27, 29)) %>% 
  footnote(general = c("Abbreviations: CI, confidence intervals; NDNS RP, national dietary and nutrition survey rolling programme; BMI body mass index; WC, waist circumference; A1C, haemoglobin A1c; DM, diabetes mellitus.", 
                       "Variables from the blood tests (glucose and A1C) are weighted by blood sample weights, the others are weighted by individual weights.", 
                       "Glucose and A1C levels are estimated in subgroups of people without diabetes.")) %>% 
  add_footnote(c("For continuous variables, the F test was used to determine differences between latent classes. For categorical variables, differences between latent classes were assessed using the adjusted Pearson Chi-2 test for survey data.",
                 "Hypertension was defined as either systolic blood pressure >= 140 mmHg or diastolic blood pressure >= 90 mmHg, or under treatment for hypertension.", 
                 "Carbohydrate percent indicates the percentage of energy from carbohydrate in total energy intake",
                 "DM was defined by A1C > 6.5%."), 
               notation = "symbol")
```

# LCGA 2 classes results

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
dt <- read.csv("/home/wangcc-me/Documents/LSHTMproject/Tablecsv/LCGA2Cresults.csv", header = T, row.names = NULL, colClasses = "character")
names(dt) <- c("Variables", "Latent class = 1", "Latent class = 2", "P value [note]")

dt[28, 1] <- c("Hypertension (%) [note]")
dt[32, 1] <- c("Carbohydrate percent (%) [note]")
dt[35, 1] <- c("DM [note]")

kable(dt,  align = "l", escape = F, row.names = FALSE,
              format = "html", caption = "Means, percentages, and 95% CIs of the characteristics by latent class memberships in the UK adults (NDNS RP 2008/09-2015/16, sample size = 6155).") %>%
 kable_styling(font_size = 16, 
        bootstrap_options = c("striped", "hover", "condensed"),
        position = "center", full_width = FALSE) %>% 
  add_indent(c(3:6, 9, 10, 12:19, 23:25, 27, 29)) %>% 
  footnote(general = c("Abbreviations: CI, confidence intervals; NDNS RP, national dietary and nutrition survey rolling programme; BMI body mass index; WC, waist circumference; A1C, haemoglobin A1c; DM, diabetes mellitus.", 
                       "Variables from the blood tests (glucose and A1C) are weighted by blood sample weights, the others are weighted by individual weights.", 
                       "Glucose and A1C levels are estimated in subgroups of people without diabetes.")) %>% 
  add_footnote(c("For continuous variables, the F test was used to determine differences between latent classes. For categorical variables, differences between latent classes were assessed using the adjusted Pearson Chi-2 test for survey data.",
                 "Hypertension was defined as either systolic blood pressure >= 140 mmHg or diastolic blood pressure >= 90 mmHg, or under treatment for hypertension.", 
                 "Carbohydrate percent indicates the percentage of energy from carbohydrate in total energy intake",
                 "DM was defined by A1C > 6.5%."), 
               notation = "symbol")
```

# LCGA 3 classes results

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
dt <- read.csv("/home/wangcc-me/Documents/LSHTMproject/Tablecsv/LCGA3Cresults.csv", header = T, row.names = NULL, colClasses = "character")
names(dt) <- c("Variables", "Latent class = 1", "Latent class = 2", "Latent class = 3", "P value [note]")

dt[28, 1] <- c("Hypertension (%) [note]")
dt[32, 1] <- c("Carbohydrate percent (%) [note]")
dt[35, 1] <- c("DM [note]")

kable(dt,  align = "l", escape = F, row.names = FALSE,
              format = "html", caption = "Means, percentages, and 95% CIs of the characteristics by latent class memberships in the UK adults (NDNS RP 2008/09-2015/16, sample size = 6155).") %>%
 kable_styling(font_size = 16, 
        bootstrap_options = c("striped", "hover", "condensed"),
        position = "center", full_width = FALSE) %>% 
  add_indent(c(3:6, 9, 10, 12:19, 23:25, 27, 29)) %>% 
  footnote(general = c("Abbreviations: CI, confidence intervals; NDNS RP, national dietary and nutrition survey rolling programme; BMI body mass index; WC, waist circumference; A1C, haemoglobin A1c; DM, diabetes mellitus.", 
                       "Variables from the blood tests (glucose and A1C) are weighted by blood sample weights, the others are weighted by individual weights.", 
                       "Glucose and A1C levels are estimated in subgroups of people without diabetes.")) %>% 
  add_footnote(c("For continuous variables, the F test was used to determine differences between latent classes with Bonferroni correction to account for multiple testing across >2 classes. For categorical variables, differences between latent classes were assessed using the adjusted Pearson Chi-2 test for survey data.",
                 "Hypertension was defined as either systolic blood pressure >= 140 mmHg or diastolic blood pressure >= 90 mmHg, or under treatment for hypertension.", 
                 "Carbohydrate percent indicates the percentage of energy from carbohydrate in total energy intake",
                 "DM was defined by A1C > 6.5%."), 
               notation = "symbol")
```