### 2018-07-28 results/posterior analysis.Rmd

---
title: "Svyset analysis"
author: "Chaochen Wang"
date: "25 July 2018"
output:
  html_document:
    nature:
      highlightLines: yes
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, message=FALSE}
statapath <- "/usr/local/stata15/stata" # <- Ubuntu path to stata
#statapath <- "D:/software/STATA/Stata-64.exe" # <- Windows STATA path modify if needed
knitr::opts_chunk$set(engine.path = list(
  stata = statapath
))
library(naniar)
library(epiDisplay)
library(readr)
library(plyr)
library(dplyr)
library(tidyverse)
library(haven)
library(survey)
 # install packages if needed
```

# Read the datasets and prepare datasets

## CW2CB2 data set (2by2 multilevel classes)

```{r cache=TRUE, message=FALSE}
CW2CB2 <- read_table2("~/Documents/LSHTMproject/results/50NDNS_CW2CB2.txt",  
                              col_names = FALSE)# change the path to your own path

names(CW2CB2) <- c("H0", "H1", "H2", "H3", "H4", "H5", "H6", "H7",
                   "H8", "H9", "H10", "H11", "H12", "H13", "H14",
                   "H15", "H16", "H17", "H18", "H19", "H20", "H21",
                   "H22", "H23", "ID_DAY", "AGE", "SEX", "CPROB1",
                   "CPROB2", "CPROB3", "CPROB4",
                   # "CPROB5",
                   # "CPROB6",
                   "CB", "CW", "MLCJOINT", "ID")
CW2CB2_reg <- CW2CB2[!duplicated(CW2CB2$ID), ]

CW2CB2_reg <- CW2CB2_reg %>% 
  select(ID, AGE, SEX, CB) # extract only the CB variable (Between individual classes == 1 or 2)

tab1(CW2CB2_reg$CB, graph = FALSE)
```

## CW3CB2 data set (3by2 multilevel classes)

```{r message=FALSE, cache=TRUE}
CW3CB2 <- read_table2("~/Documents/LSHTMproject/results/50NDNS_CW3CB2.txt",  
                      col_names = FALSE)# change the path to your own path

names(CW3CB2) <- c("H0", "H1", "H2", "H3", "H4", "H5", "H6", "H7",
                   "H8", "H9", "H10", "H11", "H12", "H13", "H14",
                   "H15", "H16", "H17", "H18", "H19", "H20", "H21",
                   "H22", "H23", "ID_DAY", "AGE", "SEX", "CPROB1",
                   "CPROB2", "CPROB3", "CPROB4", "CPROB5", "CPROB6",
                   "CB", "CW", "MLCJOINT", "ID")
CW3CB2_reg <- CW3CB2[!duplicated(CW3CB2$ID), ]

CW3CB2_reg <- CW3CB2_reg %>% 
  select(ID, AGE, SEX, CB) # extract only the CB variable (Between individual classes == 1 or 2)

tab1(CW3CB2_reg$CB, graph = FALSE)
```

## LCGA data set (2 classes)

```{r message=FALSE, cache=TRUE, warning=FALSE}
CARB_50_LGCA_2CLASS <- read_table2("~/Documents/LSHTMproject/results/LCGA/CARB_50_LGCA_2CLASS.txt", 
    col_names = FALSE)# change the path to your own path

names(CARB_50_LGCA_2CLASS) <- c("H0_X",  "H1_X", "H2_X",  "H3_X", "H4_X",  "H5_X", "H6_X",  "H7_X",
  "H8_X",  "H9_X",  "H10_X",  "H11_X", "H12_X",  "H13_X", "H14_X",  "H15_X", "H16_X",  "H17_X",
  "H18_X",  "H19_X", "H20_X",  "H21_X", "H22_X",  "H23_X", "H0_Y",  "H1_Y", "H2_Y",  "H3_Y",
  "H4_Y",  "H5_Y", "H6_Y",  "H7_Y", "H8_Y",  "H9_Y", "H10_Y",  "H11_Y", "H12_Y",  "H13_Y",
  "H14_Y",  "H15_Y", "H16_Y",  "H17_Y", "H18_Y",  "H19_Y", "H20_Y",  "H21_Y",
  "H22_Y",  "H23_Y", "H0_X_X",  "H1_X_X", "H2_X_X",  "H3_X_X", "H4_X_X",  "H5_X_X",
  "H6_X_X",  "H7_X_X", "H8_X_X",  "H9_X_X", "H10_X_X",  "H11_X_X", "H12_X_X",  "H13_X_X",
  "H14_X_X",  "H15_X_X", "H16_X_X",  "H17_X_X", "H18_X_X",  "H19_X_X", "H20_X_X",  "H21_X_X",
  "H22_X_X",  "H23_X_X", "H0_Y_Y",  "H1_Y_Y", "H2_Y_Y",  "H3_Y_Y", "H4_Y_Y",  "H5_Y_Y",
  "H6_Y_Y",  "H7_Y_Y", "H8_Y_Y",  "H9_Y_Y", "H10_Y_Y",  "H11_Y_Y", "H12_Y_Y",  "H13_Y_Y",
  "H14_Y_Y",  "H15_Y_Y", "H16_Y_Y",  "H17_Y_Y", "H18_Y_Y",  "H19_Y_Y", "H20_Y_Y",  "H21_Y_Y",
  "H22_Y_Y",  "H23_Y_Y", "ID", "CPROB1",  "CPROB2",
  #"CPROB3",  #"CPROB4", 
  "C")

CARB_50_LGCA_2CLASS[CARB_50_LGCA_2CLASS == "*"] <- NA

LCGA_2class <- CARB_50_LGCA_2CLASS %>% 
  select(ID, C) # extract only the C variable (classes == 1 or 2)

tab1(LCGA_2class$C, graph = FALSE)
```

## LCGA data set (3 classes)

```{r message=FALSE, cache=TRUE, warning=FALSE}
CARB_50_LGCA_3CLASS <- read_table2("~/Documents/LSHTMproject/results/LCGA/CARB_50_LGCA_3CLASS.DAT", 
    col_names = FALSE)# change the path to your own path

names(CARB_50_LGCA_3CLASS) <- c("H0_X",  "H1_X", "H2_X",  "H3_X", "H4_X",  "H5_X", "H6_X",  "H7_X",
  "H8_X",  "H9_X",  "H10_X",  "H11_X", "H12_X",  "H13_X", "H14_X",  "H15_X", "H16_X",  "H17_X",
  "H18_X",  "H19_X", "H20_X",  "H21_X", "H22_X",  "H23_X", "H0_Y",  "H1_Y", "H2_Y",  "H3_Y",
  "H4_Y",  "H5_Y", "H6_Y",  "H7_Y", "H8_Y",  "H9_Y", "H10_Y",  "H11_Y", "H12_Y",  "H13_Y",
  "H14_Y",  "H15_Y", "H16_Y",  "H17_Y", "H18_Y",  "H19_Y", "H20_Y",  "H21_Y",
  "H22_Y",  "H23_Y", "H0_X_X",  "H1_X_X", "H2_X_X",  "H3_X_X", "H4_X_X",  "H5_X_X",
  "H6_X_X",  "H7_X_X", "H8_X_X",  "H9_X_X", "H10_X_X",  "H11_X_X", "H12_X_X",  "H13_X_X",
  "H14_X_X",  "H15_X_X", "H16_X_X",  "H17_X_X", "H18_X_X",  "H19_X_X", "H20_X_X",  "H21_X_X",
  "H22_X_X",  "H23_X_X", "H0_Y_Y",  "H1_Y_Y", "H2_Y_Y",  "H3_Y_Y", "H4_Y_Y",  "H5_Y_Y",
  "H6_Y_Y",  "H7_Y_Y", "H8_Y_Y",  "H9_Y_Y", "H10_Y_Y",  "H11_Y_Y", "H12_Y_Y",  "H13_Y_Y",
  "H14_Y_Y",  "H15_Y_Y", "H16_Y_Y",  "H17_Y_Y", "H18_Y_Y",  "H19_Y_Y", "H20_Y_Y",  "H21_Y_Y",
  "H22_Y_Y",  "H23_Y_Y", "ID", "CPROB1",  "CPROB2",
  "CPROB3",   "C")

CARB_50_LGCA_3CLASS[CARB_50_LGCA_3CLASS == "*"] <- NA

LCGA_3class <- CARB_50_LGCA_3CLASS %>% 
  select(ID, C) # extract only the C variable (classes == 1 or 2)

tab1(LCGA_3class$C, graph = FALSE)
```

## Individual data (BMI, WC, blood pressure, A1C etc.)

```{r message=FALSE, cache=TRUE, warning=FALSE}
# change the following path according to your own data folders

blood78 <- read_dta("~/Downloads/UKDA-6533-stata11_se/stata11_se/ndns_rp_yr7-8a_indiv.dta")
blood56 <- read_dta("~/Downloads/UKDA-6533-stata11_se/stata11_se/ndns_rp_yr5-6a_indiv.dta")
blood14 <- read_dta("~/Downloads/UKDA-6533-stata11_se/stata11_se/ndns_rp_yr1-4a_indiv_uk.dta")

food14 <- read_dta("~/Downloads/UKDA-6533-stata11_se/stata11_se/ndns_rp_yr1-4a_personleveldietarydata_uk.dta")

food56 <- read_dta("~/Downloads/UKDA-6533-stata11_se/stata11_se/ndns_rp_yr5-6a_personleveldietarydata.dta")

food78 <- read_dta("~/Downloads/UKDA-6533-stata11_se/stata11_se/ndns_rp_yr7-8a_personleveldietarydata.dta")

names(blood78)[names(blood78)=="seriali"] <- "ID"
names(blood56)[names(blood56)=="seriali"] <- "ID"
names(blood14)[names(blood14)=="seriali"] <- "ID"
names(food78)[names(food78)=="seriali"] <- "ID"
names(food56)[names(food56)=="seriali"] <- "ID"
names(food14)[names(food14)=="seriali"] <- "ID"

BMI78 <- blood78 %>% 
  select(ID, Sex, age, bmival, wstval, Diabetes, bpmedc2, bpmedd2, hyper140_2, hibp140_2, MVPAtime, 
         Glucose, A1C, cigsta3, dnoft3, dnnow, wti_Y78, wtn_Y78, wtb_Y78, cluster1, cluster2, cluster3, 
         cluster4, cluster5, area, gor) %>% 
  rename(wti = wti_Y78, wtn = wtn_Y78, wtb = wtb_Y78, drink = dnoft3) %>% 
  mutate(Years = "7-8") %>% 
  replace_with_na(replace = list(bmival = -1, 
                                 wstval = -1, 
                                 bpmedd2 = -1, 
                                 bpmedc2 = -1,
                                 hyper140_2 = -7, 
                                 # hyper140_2 = -1, 
                                 hibp140_2 = -7, 
                                 # hibp140_2 = -1, 
                                 Glucose = -1, 
                                 A1C  = -1, 
                                 dnnow = -1,
                                 drink = -1,
                                 cigsta3 = -1, 
                                 MVPAtime = -4)) %>% 
  replace_with_na(replace = list(hyper140_2 = -1, hibp140_2 = -1, MVPAtime = -1,
                                 drink = -8)) %>% 
    replace_with_na(replace = list(drink = -9,
                                 cigsta3 = -8))

BMI56 <- blood56 %>% 
  select(ID, Sex, age, area, bmival, wstval, Diabetes, bpmedc2, bpmedd2, hyper140_2, hibp140_2, MVPAtime,
         Glucose, A1C, cigsta3, dnoft3, dnnow, wti_Y56, wtn_Y56, wtb_Y56, cluster1, cluster2, cluster3, 
         cluster4, cluster5, area, gor) %>% 
  mutate(Years = "5-6") %>% 
  rename(wti = wti_Y56, wtn = wtn_Y56, wtb = wtb_Y56, drink = dnoft3) %>% 
  replace_with_na(replace = list(bmival = -1, 
                                 wstval = -1, 
                                 bpmedd2 = -1, 
                                 bpmedc2 = -1,
                                 hyper140_2 = -7, 
                                 # hyper140_2 = -1, 
                                 hibp140_2 = -7, 
                                 # hibp140_2 = -1, 
                                 Glucose = -1, 
                                 A1C  = -1, 
                                 dnnow = -1,
                                 drink = -1,
                                 cigsta3 = -1,
                                 MVPAtime = -1)) %>% 
  replace_with_na(replace = list(hyper140_2 = -1, hibp140_2 = -1, 
                                 drink = -8)) %>% 
    replace_with_na(replace = list(drink = -9,
                                 cigsta3 = -8))

BMI14 <- blood14 %>% 
  select(ID, Sex, age, bmival, wstval, Diabetes, bpmedc, bpmedd, hyper140, hibp140, MVPAtime,
         Glucose, A1C, cigsta3, dnoft3, dnnow, wti_CY1234, wtn_CY1234, wtb_CY1234, cluster, area, gor) %>%
  rename(hyper140_2 = hyper140, hibp140_2 = hibp140, bpmedd2 = bpmedd, 
         bpmedc2 = bpmedc, cluster1 = cluster, 
         wti = wti_CY1234, wtn = wtn_CY1234, wtb =  wtb_CY1234, drink = dnoft3) %>% 
  mutate(cluster2 = NA, cluster3 = NA, cluster4 = NA, cluster5 = NA, Years = "1-4") %>% 
  replace_with_na(replace = list(bmival = -1, 
                                 wstval = -1, 
                                 bpmedd2 = -1, 
                                 bpmedc2 = -1,
                                 hyper140_2 = -7, 
                                 # hyper140_2 = -1, 
                                 hibp140_2 = -7, 
                                 # hibp140_2 = -1, 
                                 Glucose = -1, 
                                 A1C  = -1, 
                                 dnnow = -1,
                                 drink = -1,
                                 cigsta3 = -1,
                                 MVPAtime = -4)) %>% 
  replace_with_na(replace = list(hyper140_2 = -1, hibp140_2 = -1, MVPAtime = -1,
                                 drink = -8)) %>% 
  replace_with_na(replace = list(drink = -9,
                                 cigsta3 = -8))

BMI <- bind_rows(BMI14, BMI56, BMI78)

Energy14 <- food14 %>% 
  select(ID, Country, SurveyYear, EnergykJ, Carbohydrateg, CHOpctotE)
Energy14$Country[Energy14$Country == "Northern Ireland"] <- "NI"

Energy56 <- food56 %>% 
  select(ID, Country, Surveyyear, EnergykJ, Carbohydrateg, CHOpctotE) %>% 
  rename(SurveyYear = Surveyyear)

Energy78 <- food78 %>% 
  select(ID, Country, SurveyYear, EnergykJ, Carbohydrateg, CHOpctotE)

Energy <- bind_rows(Energy14, Energy56, Energy78)

CW2CB2_regss <- CW2CB2_reg %>% 
  left_join(BMI, by = "ID") %>% 
  left_join(Energy, by = "ID")## dataset for 2by2 multilevel latent classes

CW3CB2_regss <- CW3CB2_reg %>% 
  left_join(BMI, by = "ID") %>% 
  left_join(Energy, by = "ID")## dataset for 3by2 multilevel latent classes

LCGA_2class <- LCGA_2class %>% 
  left_join(BMI, by = "ID") %>% 
  left_join(Energy, by = "ID")## dataset for 2 classes LCGA

LCGA_3class <- LCGA_3class %>% 
  left_join(BMI, by = "ID") %>% 
  left_join(Energy, by = "ID")## dataset for 3 classes LCGA

rm(blood14, blood56, blood78, BMI14, BMI56, BMI78, BMI, food14, food56, food78, Energy14,    Energy56, Energy78, Energy, CW2CB2, CW2CB2_reg, CW3CB2, CW3CB2_reg, CARB_50_LGCA_2CLASS, CARB_50_LGCA_3CLASS)
```

# Rescale the weighting values

```{r message=FALSE, cache=TRUE, warning=FALSE}
# individual weighting
a <- sum(CW2CB2_regss[CW2CB2_regss$Years == "1-4",]$wti)
b <- sum(CW2CB2_regss[CW2CB2_regss$Years == "5-6",]$wti)
c <- sum(CW2CB2_regss[CW2CB2_regss$Years == "7-8",]$wti)

CW2CB2_regss$wti1to8 <- CW2CB2_regss$wti

CW2CB2_regss[CW2CB2_regss$Years == "1-4",]$wti1to8 <- CW2CB2_regss[CW2CB2_regss$Years == "1-4",]$wti*(a+b+c)*(1/2)/a
CW2CB2_regss[CW2CB2_regss$Years == "5-6",]$wti1to8 <- CW2CB2_regss[CW2CB2_regss$Years == "5-6",]$wti*(a+b+c)*(1/4)/b
CW2CB2_regss[CW2CB2_regss$Years == "7-8",]$wti1to8 <- CW2CB2_regss[CW2CB2_regss$Years == "7-8",]$wti*(a+b+c)*(1/4)/c
mean(CW2CB2_regss$wti1to8)

CW2CB2_regss$wti1to8 <- CW2CB2_regss$wti1to8/1.209816814
summ(CW2CB2_regss$wti1to8, graph = FALSE)
sum(CW2CB2_regss$wti1to8, graph = FALSE) #Check if the weighting sum up to the sample size we have

# Nurse weights

a <- sum(CW2CB2_regss[CW2CB2_regss$Years == "1-4",]$wtn)
b <- sum(CW2CB2_regss[CW2CB2_regss$Years == "5-6",]$wtn)
c <- sum(CW2CB2_regss[CW2CB2_regss$Years == "7-8",]$wtn)

CW2CB2_regss$wtn1to8 <- CW2CB2_regss$wtn

CW2CB2_regss[CW2CB2_regss$Years == "1-4",]$wtn1to8 <- CW2CB2_regss[CW2CB2_regss$Years == "1-4",]$wtn*(a+b+c)*(1/2)/a
CW2CB2_regss[CW2CB2_regss$Years == "5-6",]$wtn1to8 <- CW2CB2_regss[CW2CB2_regss$Years == "5-6",]$wtn*(a+b+c)*(1/4)/b
CW2CB2_regss[CW2CB2_regss$Years == "7-8",]$wtn1to8 <- CW2CB2_regss[CW2CB2_regss$Years == "7-8",]$wtn*(a+b+c)*(1/4)/c
mean(CW2CB2_regss$wtn1to8)

CW2CB2_regss$wtn1to8 <- CW2CB2_regss$wtn1to8/0.907003577
summ(CW2CB2_regss$wtn1to8, graph = FALSE)
sum(CW2CB2_regss$wtn1to8, graph = FALSE) #Check if the weighting sum up to the sample size we have

# Blood weights
a <- sum(CW2CB2_regss[CW2CB2_regss$Years == "1-4",]$wtb)
b <- sum(CW2CB2_regss[CW2CB2_regss$Years == "5-6",]$wtb)
c <- sum(CW2CB2_regss[CW2CB2_regss$Years == "7-8",]$wtb)

CW2CB2_regss$wtb1to8 <- CW2CB2_regss$wtb

CW2CB2_regss[CW2CB2_regss$Years == "1-4",]$wtb1to8 <- CW2CB2_regss[CW2CB2_regss$Years == "1-4",]$wtb*(a+b+c)*(1/2)/a
CW2CB2_regss[CW2CB2_regss$Years == "5-6",]$wtb1to8 <- CW2CB2_regss[CW2CB2_regss$Years == "5-6",]$wtb*(a+b+c)*(1/4)/b
CW2CB2_regss[CW2CB2_regss$Years == "7-8",]$wtb1to8 <- CW2CB2_regss[CW2CB2_regss$Years == "7-8",]$wtb*(a+b+c)*(1/4)/c
mean(CW2CB2_regss$wtb1to8)

CW2CB2_regss$wtb1to8 <- CW2CB2_regss$wtb1to8/0.4817444505
summ(CW2CB2_regss$wtb1to8, graph = FALSE)
sum(CW2CB2_regss$wtb1to8, graph = FALSE) #Check if the weighting sum up to the sample size we have

weightings <- CW2CB2_regss %>% select(ID, wti1to8, wtn1to8, wtb1to8)

# add the weightings to the other datasets
CW3CB2_regss <- CW3CB2_regss %>% 
  left_join(weightings, by = "ID")

LCGA_2class <- LCGA_2class %>% 
  left_join(weightings, by = "ID")

LCGA_3class <- LCGA_3class %>% 
  left_join(weightings, by = "ID")
```

# (BMI) survey designed analysis

## 2by2 multilevel LCA dataset (BMI)
```{r message=FALSE, cache=TRUE, warning=FALSE, eval=FALSE}
 # specifying a survey design

CW2CB2_regss$dnnow <- as.factor(CW2CB2_regss$dnnow)
CW2CB2_regss$cigsta3 <- as.factor(CW2CB2_regss$cigsta3)

cw2cb2 <- svydesign(id = ~area, strat = ~gor, weights=~wti1to8, data = CW2CB2_regss, nest = TRUE)

summary(svyglm(bmival ~ CB, design = cw2cb2))
summary(svyglm(bmival ~ CB + AGE + SEX + cigsta3 + dnnow, design = cw2cb2))
```

```
## Call:
## svyglm(formula = bmival ~ CB, design = cw2cb2)
## 
## Survey design:
## svydesign(id = ~area, strat = ~gor, weights = ~wti1to8, data = CW2CB2_regss, 
##     nest = TRUE)
## 
## Coefficients:
##             Estimate Std. Error t value Pr(>|t|)    
## (Intercept)  26.6815     0.2917  91.468  < 2e-16 ***
## CB            0.5474     0.2026   2.702  0.00699 ** 
## ---
## Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
## 
## (Dispersion parameter for gaussian family taken to be 28.88983)
## 
## Number of Fisher Scoring iterations: 2
## 
## 
## Call:
## svyglm(formula = bmival ~ CB + AGE + SEX + cigsta3 + dnnow, design = cw2cb2)
## 
## Survey design:
## svydesign(id = ~area, strat = ~gor, weights = ~wti1to8, data = CW2CB2_regss, 
##     nest = TRUE)
## 
## Coefficients:
##             Estimate Std. Error t value Pr(>|t|)    
## (Intercept) 24.53607    0.49745  49.324  < 2e-16 ***
## CB           0.44258    0.20174   2.194   0.0285 *  
## AGE          0.04177    0.00582   7.176 1.33e-12 ***
## SEX         -0.21736    0.18620  -1.167   0.2433    
## cigsta32     1.18384    0.29947   3.953 8.22e-05 ***
## cigsta33     0.48760    0.24885   1.959   0.0503 .  
## dnnow2       0.42970    0.24531   1.752   0.0801 .  
## ---
## Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
## 
## (Dispersion parameter for gaussian family taken to be 27.97746)
## 
## Number of Fisher Scoring iterations: 2
```

After adjusting for age, sex, smoking, and drinking, subjects in latent class 2 were averagely with 0.748599 kg/m^2^ higher BMI compared with subjects in latent class 1.

### Compare with results in Stata (they are very similar, and we use Stata subsequently):

```{r engine='stata', echo=FALSE}
use "/home/wangcc-me/Downloads/UKDA-6533-stata11_se/stata11_se/CW2CB2_regss.dta", clear

label define smoking 1 "current" 2 "ex-smoker" 3 "Never"
label values cigsta3 smoking
label define drinking 1 "no" 2 "yes"
label values dnnow drinking
label define gender 1 "Men" 2 "Women"
label values Sex gender

// weighting with individual weights, area is primary sampling unit, gor is the cluster variable

svyset area [pweight = wti1to8], strata(gor)

svydescribe wti
svy: mean bmival

// two-way table

svy: tabulate Sex CB, row se ci format(%7.3f)
svy: tabulate Sex CB, col se ci format(%7.3f)

// comparing means
svy: mean bmi, over(CB)

svy: regress bmival i.CB
svy: regress bmival i.CB age i.Sex i.cigsta3 i.dnnow

svy: regress wstval i.CB#i.Sex age  i.cigsta3 i.dnnow

svy: regress wst i.CB age i.cigsta3 i.dnnow if Sex == 1
svy: regress wst i.CB age i.cigsta3 i.dnnow if Sex == 2
```

## 3by2 multilevel LCA dataset (BMI)

```{r engine='stata', echo=FALSE}
use "/home/wangcc-me/Downloads/UKDA-6533-stata11_se/stata11_se/CW3CB2_regss.dta", clear

label define smoking 1 "current" 2 "ex-smoker" 3 "Never"
label values cigsta3 smoking
label define drinking 1 "no" 2 "yes"
label values dnnow drinking
label define gender 1 "Men" 2 "Women"
label values Sex gender

svyset area [pweight = wti1to8], strata(gor)
svydescribe wti
svy: mean bmival
svy: mean bmival if CB == 1
svy: mean bmival if CB == 2

svy: regress bmival i.CB
svy: regress bmival i.CB age i.Sex i.cigsta3 i.dnnow
```

## 2 classes LCGA dataset (BMI)

```{r engine='stata', echo=FALSE}

use "/home/wangcc-me/Downloads/UKDA-6533-stata11_se/stata11_se/LCGA_2class.dta", clear

label define smoking 1 "current" 2 "ex-smoker" 3 "Never"
label values cigsta3 smoking
label define drinking 1 "no" 2 "yes"
label values dnnow drinking
label define gender 1 "Men" 2 "Women"
label values Sex gender

svyset area [pweight = wti1to8], strata(gor)
svydescribe wti
svy: mean bmival
svy: mean bmival if C == 1
svy: mean bmival if C == 2

svy: regress bmival i.C
svy: regress bmival i.C age i.Sex i.cigsta3 i.dnnow
```

## 3 classes LCGA dataset (BMI)

```{r engine='stata', echo=FALSE}
use "/home/wangcc-me/Downloads/UKDA-6533-stata11_se/stata11_se/LCGA_3class.dta", clear

label define smoking 1 "current" 2 "ex-smoker" 3 "Never"
label values cigsta3 smoking
label define drinking 1 "no" 2 "yes"
label values dnnow drinking
label define gender 1 "Men" 2 "Women"
label values Sex gender

svyset area [pweight = wti1to8], strata(gor)
svy: mean bmival
svy: mean bmival if C == 1
svy: mean bmival if C == 2
svy: mean bmival if C == 3
svy: regress bmival i.C
svy: regress bmival i.C age i.Sex i.cigsta3 i.dnnow
```

# (Waist Circumference) survey designed analysis

## 2by2 multilevel LCA dataset (WC)

```{r engine='stata', echo=FALSE}
use "/home/wangcc-me/Downloads/UKDA-6533-stata11_se/stata11_se/CW2CB2_regss.dta", clear

label define smoking 1 "current" 2 "ex-smoker" 3 "Never"
label values cigsta3 smoking
label define drinking 1 "no" 2 "yes"
label values dnnow drinking
label define gender 1 "Men" 2 "Women"
label values Sex gender

svyset area [pweight = wti1to8], strata(gor)

//svydescribe wti
svy: mean wstval
svy: mean wstval if CB == 1
svy: mean wstval if CB == 2

svy: regress wst i.CB
svy: regress wst i.CB age i.Sex i.cigsta3 i.dnnow
svy: regress wstval i.CB#i.Sex age  i.cigsta3 i.dnnow

svy: regress wst i.CB age i.cigsta3 i.dnnow if Sex == 1
svy: regress wst i.CB age i.cigsta3 i.dnnow if Sex == 2
```

## 3by2 multilevel LCA dataset (WC)

```{r engine='stata', echo=FALSE}
use "/home/wangcc-me/Downloads/UKDA-6533-stata11_se/stata11_se/CW3CB2_regss.dta", clear

label define smoking 1 "current" 2 "ex-smoker" 3 "Never"
label values cigsta3 smoking
label define drinking 1 "no" 2 "yes"
label values dnnow drinking
label define gender 1 "Men" 2 "Women"
label values Sex gender

svyset area [pweight = wti1to8], strata(gor)
//svydescribe wti
svy: mean wstval
svy: mean wstval if CB == 1
svy: mean wstval if CB == 2

svy: regress wstval i.CB
svy: regress wstval i.CB age i.Sex i.cigsta3 i.dnnow
svy: regress wstval i.CB#i.Sex age  i.cigsta3 i.dnnow

svy: regress wst i.CB age i.cigsta3 i.dnnow if Sex == 1
svy: regress wst i.CB age i.cigsta3 i.dnnow if Sex == 2
```