### 2018-07-30 results/Timeslots/multilevel.Rmd

---
title: "NDNS multilevel carb_50, 7 time slots"
author: "Chaochen Wang"
date: "29 July 2018 created, `r Sys.Date()` updated"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    nature:
      highlightLines: true
editor_options: 
  chunk_output_type: console
---

# Level 1 classes selection

- In this analysis, within each **time slot** of the day carbohydrates intake was defined as: 
    - not eating; 
    - eating and carbohydrates contributed less than 50% of energy;
    - eating and carbohydrates contributed higher or equal to 50% of energy.

- Time slots are defined as: 
    - **Breakfast:** 6 am to 9 am; 
    - **Morning snack:** 9 am to 12 noon; 
    - **Lunch:** 12 noon to 2 pm; 
    - **Afternoon snack:** 2 pm to 5 pm; 
    - **Dinner:** 5 pm to 8 pm; 
    - **Before bedtime snack:** 8 pm to 10 pm; 
    - **Midnight food:** 10pm to 6 am;

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(RColorBrewer)
library(plyr)
library(tidyverse)
library(dplyr)
library(epiDisplay)
library(scales)
library(readr)
rm(list=ls(all=TRUE))
# load("~/Documents/LSHTMproject/Rcode/day4LCA1_6.Rdata")

library(kableExtra)
dt <- read.csv("/home/wangcc-me/Documents/LSHTMproject/Tablecsv/Level1_7slots.csv", header = T, colClasses = "character")
names(dt) <- c("N of classes", "N of free parameters","log-likelihood",  "AIC", "BIC", "aBIC", "Entropy", "Lo-Mendel-Rubin LRT")
# dt[1,8] <- "--"
dt[2,8] <- "< 0.0001"
dt[4,8] <- "< 0.0001"
dt[5,8] <- "< 0.0001"
kable(dt, digits = 3, row.names = FALSE, align = "c",
              format = "html", caption = "Level 1 class selection. (All data, n = 6155, 24483 data points)") %>%
 kable_styling(font_size = 16, 
        bootstrap_options = c("striped", "hover", "condensed"),
        position = "center", full_width = FALSE) %>% 
  # row_spec(c(2,3), bold = T, color = "white", background = "#D7261E") %>% 
  footnote(general = "Abbreviation: N, number; AIC, Akaike information criterion; BIC, Bayesian information criterion; aBIC, adjusted BIC; Entropy, a pseudo-r-squared index; Lo-Mendel-Rubin LRT, likelihood ratio test comparing q classes models with q-1 classes models.")
```

## Fit information for each model (combination of level 1 and level 2 classes)

```{r echo=FALSE}
rm(list=ls(all=TRUE))
# load("~/Documents/LSHTMproject/Rcode/day4LCA1_6.Rdata")

library(kableExtra)
dt <- read.csv("/home/wangcc-me/Documents/LSHTMproject/Tablecsv/Mixed_7slots.csv", header = T)
names(dt) <- c("Model", "1 class","2 classes",  "3 classes", "4 classes", "5 classes", "6 classes")
dt[is.na(dt)] <- ""
# library(formattable)
dt %>% 
  # mutate(
  #   `2 classes` = cell_spec(`2 classes`, color = ifelse())) %>% 
  kable(escape = F, 
        digits = 3, row.names = FALSE, align = "l",
        format = "html", caption = "Models Specification") %>% 
  add_header_above(c(" ", "Number of level 1 classes" = 6)) %>% 
  group_rows("Fixed effects model", 1, 5) %>%
  group_rows("Random effects model", 6, 20) %>%
  add_indent(c(2:5,7:10,12:15,17:20)) %>% 
  kable_styling(font_size = 16, 
        bootstrap_options = c("striped", "hover", "condensed"),
        position = "center", full_width = FALSE) %>% 
  #row_spec(c(2,3), bold = T, color = "white", background = "#D7261E") %>% 
  footnote(general = "Abbreviation: No, number; BIC, Bayesian information criterion; Entropy, a pseudo-r-squared index; Lo-Mendel-Rubin LRT, likelihood ratio test comparing q classes models with q-1 classes models.")
```

# (CW = 2) Level 1 latent classes

## Visualisation of level 1 latent classes

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

CW2CB2 <- read_table2("/home/wangcc-me/Documents/LSHTMproject/results/Timeslots/NDNSslot_CW2CB2.txt",
                         col_names = FALSE)

names(CW2CB2) <- c("Breakfast", 
                   "Morning.snack",
                   "Lunch",
                   "Afternoon.snack",
                   "Dinner",
                   "Before.bedtime.snack",
                   "Midnight.food",
                      "ID_DAY",
                      "AGE",
                      "SEX",
                      "CPROB1",
                      "CPROB2",
                      "CPROB3",
                      "CPROB4",
                      # "CPROB5",
                      # "CPROB6",
                      "CB",
                      "CW",
                      "MLCJOINT",
                      "ID")

pp <- CW2CB2 %>% 
  group_by(CW) %>% 
  summarise(Breakfast_0 = sum(Breakfast == 0)/length(Breakfast), 
            Breakfast_1 = sum(Breakfast == 1)/length(Breakfast), 
            Breakfast_2 = sum(Breakfast == 2)/length(Breakfast),
            Morning.snack_0 = sum(Morning.snack == 0)/length(Morning.snack),
            Morning.snack_1 = sum(Morning.snack == 1)/length(Morning.snack), 
            Morning.snack_2 = sum(Morning.snack == 2)/length(Morning.snack),
            Lunch_0 = sum(Lunch == 0)/length(Lunch),
            Lunch_1 = sum(Lunch == 1)/length(Lunch), 
            Lunch_2 = sum(Lunch == 2)/length(Lunch),
            Afternoon.snack_0 = sum(Afternoon.snack == 0)/length(Afternoon.snack),
            Afternoon.snack_1 = sum(Afternoon.snack == 1)/length(Afternoon.snack), 
            Afternoon.snack_2 = sum(Afternoon.snack == 2)/length(Afternoon.snack),
            Dinner_0 = sum(Dinner == 0)/length(Dinner),
            Dinner_1 = sum(Dinner == 1)/length(Dinner), 
            Dinner_2 = sum(Dinner == 2)/length(Dinner),
            Before.bedtime.snack_0 = sum(Before.bedtime.snack == 0)/length(Before.bedtime.snack),
            Before.bedtime.snack_1 = sum(Before.bedtime.snack == 1)/length(Before.bedtime.snack), 
            Before.bedtime.snack_2 = sum(Before.bedtime.snack == 2)/length(Before.bedtime.snack),
            Midnight.food_0 = sum(Midnight.food == 0)/length(Midnight.food),
            Midnight.food_1 = sum(Midnight.food == 1)/length(Midnight.food), 
            Midnight.food_2 = sum(Midnight.food == 2)/length(Midnight.food))

pp_long <- pp %>% 
  gather(Hour, Prob, -CW) %>% 
  separate(Hour, into = c("Slots", "Carbo"), sep = "_")

pp_long$Slots <- factor(pp_long$Slots, levels = c("Breakfast", 
                   "Morning.snack",
                   "Lunch",
                   "Afternoon.snack",
                   "Dinner",
                   "Before.bedtime.snack",
                   "Midnight.food"))
```

```{r echo=FALSE, fig.height=6, fig.width=14, message=FALSE, warning=FALSE, cache=TRUE}
library(ggthemr)
ggthemr("fresh", layout = "scientific")
library(ggplot2)

pp_long$Time <- as.character(pp_long$Slots)
pp_long$Time[pp_long$Slots == "Breakfast"] <- 7.5
pp_long$Time[pp_long$Slots == "Morning.snack"] <- 10.5
pp_long$Time[pp_long$Slots == "Lunch"] <- 13
pp_long$Time[pp_long$Slots == "Afternoon.snack"] <- 15.5
pp_long$Time[pp_long$Slots == "Dinner"] <- 18.5
pp_long$Time[pp_long$Slots == "Before.bedtime.snack"] <- 21
pp_long$Time[pp_long$Slots == "Midnight.food"] <- 26
pp_long$Time <- as.numeric(pp_long$Time)

ggplot(pp_long[pp_long$CW == 1, ], aes(y = Prob, x=Time, group = Carbo, color = Carbo))  + 
  scale_x_continuous(limits = c(6, 29),
  breaks = seq(6, 29, by = 1)) +
  annotate("rect", xmin = 6, xmax = 9, ymin = 0, ymax = 1.00, 
            alpha = 0.5, fill = "grey") + 
  annotate("rect", xmin = 12, xmax = 14, ymin = 0, ymax = 1.00, 
             alpha = 0.5, fill = "grey") + 
  annotate("rect", xmin = 17, xmax = 20, ymin = 0, ymax = 1.00, 
            alpha = 0.5, fill = "grey") + 
  annotate("rect", xmin = 22, xmax = 29, ymin = 0, ymax = 1.00, 
             alpha = 0.5, fill = "grey") +
  geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
  ) + 
    # theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
  labs(title = "Class 1 days", x = "Hours of the day (next day from hour 24)", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Not eating", "< 50%", ">= 50%")) + 
  ylim(c(0,1)) #+ 
  # scale_x_discrete(labels = c("6","7","8","9","10","11","12","13","14",
  #                             "15","16","17","18","19", "20","21","22",
  #                             "23","0","1","2","3","4","5"))

ggplot(pp_long[pp_long$CW == 2, ], aes(y = Prob, x=Time, group = Carbo, color = Carbo)) + 
  scale_x_continuous(limits = c(6, 29),
  breaks = seq(6, 29, by = 1)) +
  annotate("rect", xmin = 6, xmax = 9, ymin = 0, ymax = 1.00, 
            alpha = 0.5, fill = "grey") + 
  annotate("rect", xmin = 12, xmax = 14, ymin = 0, ymax = 1.00, 
             alpha = 0.5, fill = "grey") + 
  annotate("rect", xmin = 17, xmax = 20, ymin = 0, ymax = 1.00, 
            alpha = 0.5, fill = "grey") + 
  annotate("rect", xmin = 22, xmax = 29, ymin = 0, ymax = 1.00, 
             alpha = 0.5, fill = "grey") +
  geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
  ) + 
    # theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
  labs(title = "Class 2 days", x = "Hours of the day (next day from hour 24)", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Not eating", "< 50%", ">= 50%")) + 
  ylim(c(0,1))

```

## Visualisation of level 2 latent classes (CB = 2)

```{r echo=FALSE, fig.height=6, fig.width=7, message=FALSE, warning=FALSE, cache=TRUE}

# ndnsNNN <- read_csv("results/ndnsNNN.dat")
# dt <- ndnsNNN[, c(2,5)]
# names(dt) <- c("ID_DAY", "DayofWeek")
# 
# 
# CW2CB2 <- CW2CB2 %>% 
#   left_join(dt, by = "ID_DAY")
#

chart.data <- CW2CB2 %>% 
  group_by(CB, CW) %>% 
  tally %>% 
  group_by(CB) %>% 
  mutate(pct = n/sum(n))

chart.data <- ddply(chart.data, .(CB),
                     transform, pos = cumsum(pct) - (0.5 * pct))

chart.data$CW <- factor(chart.data$CW, levels = c("2", "1"), 
                        labels = c("Class 2 days", "Class 1 days"))
chart.data$CB <- factor(chart.data$CB, levels = c("1", "2"), 
                        labels = c("Individual Class 1\n(56.3%)",  "Individual class 2\n(43.7%)"))

library(ggthemr)
ggthemr("dust", layout = "scientific")
ggplot() + 
  geom_bar(aes(y = pct, x = CB, fill = CW), data = chart.data, width = 0.6,
           stat="identity") + 
  geom_text(data=chart.data, aes(x = CB, y = pos, label = paste0(sprintf("%1.1f", pct*100),"%")),
            size=4, colour="white", family="Atlas Grotesk Medium") + 
  theme(legend.position="bottom", legend.direction="horizontal",
        legend.title = element_blank(), 
        axis.title = element_text(size = 18), 
        axis.text = element_text(size = 15), 
        axis.line = element_line(colour = "black"), 
        plot.title=element_text(family="Atlas Grotesk Medium"),
        text=element_text(family="Atlas Grotesk Light")) + 
  labs(title = "Multilevel latent class solution", x = "Between Individual classes", y = "Percentage") +
  scale_y_continuous(labels=percent)
```

## Visualisation of level 2 latent classes (CB = 3)

```{r echo=FALSE, fig.height=6, fig.width=7, message=FALSE, warning=FALSE, cache=TRUE}

CW2CB3 <- read_table2("/home/wangcc-me/Documents/LSHTMproject/results/Timeslots/NDNSslot_CW2CB3.txt",
                         col_names = FALSE)

names(CW2CB3) <- c("Breakfast", 
                   "Morning.snack",
                   "Lunch",
                   "Afternoon.snack",
                   "Dinner",
                   "Before.bedtime.snack",
                   "Midnight.food",
                      "ID_DAY",
                      "AGE",
                      "SEX",
                      "CPROB1",
                      "CPROB2",
                      "CPROB3",
                      "CPROB4",
                      "CPROB5",
                      "CPROB6",
                      "CB",
                      "CW",
                      "MLCJOINT",
                      "ID")

chart.data <- CW2CB3 %>% 
  group_by(CB, CW) %>% 
  tally %>% 
  group_by(CB) %>% 
  mutate(pct = n/sum(n))

# chart.data$CW[chart.data$CW == 1] <- 0 
# chart.data$CW[chart.data$CW == 3] <- 1 
# chart.data$CW[chart.data$CW == 0] <- 3

chart.data <- chart.data[order(chart.data$CB, chart.data$CW),]
chart.data <- ddply(chart.data, .(CB), 
                    transform, pos = cumsum(pct) - (0.5 * pct))

chart.data$CW <- factor(chart.data$CW, levels = c("2", "1"), 
                        labels = c("Class 2 days", "Class 1 days"))
chart.data$CB <- factor(chart.data$CB, levels = c("1", "2", "3"), 
                        labels = c("Individual\nclass 1\n(42.4%)",  "Individual\nclass 2\n(29.5%)", 
                                   "Individual\nclass 3\n(28.1%)"))

library(ggthemr)
ggthemr("dust", layout = "scientific")
ggplot() + 
  geom_bar(aes(y = pct, x = CB, fill = CW), data = chart.data, width = 0.6,
           stat="identity") + 
  geom_text(data=chart.data, aes(x = CB, y = pos, label = paste0(sprintf("%1.1f", pct*100),"%")),
            size=4, colour="white", family="Atlas Grotesk Medium") + 
  theme(legend.position="bottom", legend.direction="horizontal",
        legend.title = element_blank(), 
        axis.title = element_text(size = 18), 
        axis.text = element_text(size = 15), 
        axis.line = element_line(colour = "black"), 
        plot.title=element_text(family="Atlas Grotesk Medium"),
        text=element_text(family="Atlas Grotesk Light")) + 
  labs(title = "Multilevel latent class solution", x = "Between Individual classes", y = "Percentage") +
  scale_y_continuous(labels=percent)
```

## Visualisation of level 2 latent classes (CB = 4)

```{r echo=FALSE, fig.height=6, fig.width=7, message=FALSE, warning=FALSE, cache=TRUE}

CW2CB4 <- read_table2("/home/wangcc-me/Documents/LSHTMproject/results/Timeslots/NDNSslot_CW2CB4.txt",
                         col_names = FALSE)

names(CW2CB4) <- c("Breakfast", 
                   "Morning.snack",
                   "Lunch",
                   "Afternoon.snack",
                   "Dinner",
                   "Before.bedtime.snack",
                   "Midnight.food",
                      "ID_DAY",
                      "AGE",
                      "SEX",
                      "CPROB1",
                      "CPROB2",
                      "CPROB3",
                      "CPROB4",
                      "CPROB5",
                      "CPROB6",
                      "CPROB7",
                      "CPROB8",
                      "CB",
                      "CW",
                      "MLCJOINT",
                      "ID")

chart.data <- CW2CB4 %>% 
  group_by(CB, CW) %>% 
  tally %>% 
  group_by(CB) %>% 
  mutate(pct = n/sum(n))

# chart.data$CW[chart.data$CW == 1] <- 0 
# chart.data$CW[chart.data$CW == 3] <- 1 
# chart.data$CW[chart.data$CW == 0] <- 3

chart.data <- chart.data[order(chart.data$CB, chart.data$CW),]
chart.data <- ddply(chart.data, .(CB), 
                    transform, pos = cumsum(pct) - (0.5 * pct))

chart.data$CW <- factor(chart.data$CW, levels = c("2", "1"), 
                        labels = c("Class 2 days", "Class 1 days"))
chart.data$CB <- factor(chart.data$CB, levels = c("1", "2", "3", "4"), 
                        labels = c("Individual\nclass 1\n(23.4%)",  "Individual\nclass 2\n(23.7%)", 
                                   "Individual\nclass 3\n(28.9%)", "Individual\nclass 4\n(23.9%)"))

library(ggthemr)
ggthemr("dust", layout = "scientific")
ggplot() + 
  geom_bar(aes(y = pct, x = CB, fill = CW), data = chart.data, width = 0.6,
           stat="identity") + 
  geom_text(data=chart.data, aes(x = CB, y = pos, label = paste0(sprintf("%1.1f", pct*100),"%")),
            size=4, colour="white", family="Atlas Grotesk Medium") + 
  theme(legend.position="bottom", legend.direction="horizontal",
        legend.title = element_blank(), 
        axis.title = element_text(size = 18), 
        axis.text = element_text(size = 15), 
        axis.line = element_line(colour = "black"), 
        plot.title=element_text(family="Atlas Grotesk Medium"),
        text=element_text(family="Atlas Grotesk Light")) + 
  labs(title = "Multilevel latent class solution", x = "Between Individual classes", y = "Percentage") +
  scale_y_continuous(labels=percent)
```

# (CW = 3) Level 1 latent classes

## Visualisation of level 1 latent classes

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

CW3CB2 <- read_table2("/home/wangcc-me/Documents/LSHTMproject/results/Timeslots/NDNSslot_CW3CB2.txt",
                         col_names = FALSE)

names(CW3CB2) <- c("Breakfast", 
                   "Morning.snack",
                   "Lunch",
                   "Afternoon.snack",
                   "Dinner",
                   "Before.bedtime.snack",
                   "Midnight.food",
                      "ID_DAY",
                      "AGE",
                      "SEX",
                      "CPROB1",
                      "CPROB2",
                      "CPROB3",
                      "CPROB4",
                      "CPROB5",
                      "CPROB6",
                      "CB",
                      "CW",
                      "MLCJOINT",
                      "ID")

pp <- CW3CB2 %>% 
  group_by(CW) %>% 
  summarise(Breakfast_0 = sum(Breakfast == 0)/length(Breakfast), 
            Breakfast_1 = sum(Breakfast == 1)/length(Breakfast), 
            Breakfast_2 = sum(Breakfast == 2)/length(Breakfast),
            Morning.snack_0 = sum(Morning.snack == 0)/length(Morning.snack),
            Morning.snack_1 = sum(Morning.snack == 1)/length(Morning.snack), 
            Morning.snack_2 = sum(Morning.snack == 2)/length(Morning.snack),
            Lunch_0 = sum(Lunch == 0)/length(Lunch),
            Lunch_1 = sum(Lunch == 1)/length(Lunch), 
            Lunch_2 = sum(Lunch == 2)/length(Lunch),
            Afternoon.snack_0 = sum(Afternoon.snack == 0)/length(Afternoon.snack),
            Afternoon.snack_1 = sum(Afternoon.snack == 1)/length(Afternoon.snack), 
            Afternoon.snack_2 = sum(Afternoon.snack == 2)/length(Afternoon.snack),
            Dinner_0 = sum(Dinner == 0)/length(Dinner),
            Dinner_1 = sum(Dinner == 1)/length(Dinner), 
            Dinner_2 = sum(Dinner == 2)/length(Dinner),
            Before.bedtime.snack_0 = sum(Before.bedtime.snack == 0)/length(Before.bedtime.snack),
            Before.bedtime.snack_1 = sum(Before.bedtime.snack == 1)/length(Before.bedtime.snack), 
            Before.bedtime.snack_2 = sum(Before.bedtime.snack == 2)/length(Before.bedtime.snack),
            Midnight.food_0 = sum(Midnight.food == 0)/length(Midnight.food),
            Midnight.food_1 = sum(Midnight.food == 1)/length(Midnight.food), 
            Midnight.food_2 = sum(Midnight.food == 2)/length(Midnight.food))

pp_long <- pp %>% 
  gather(Hour, Prob, -CW) %>% 
  separate(Hour, into = c("Slots", "Carbo"), sep = "_")

pp_long$Slots <- factor(pp_long$Slots, levels = c("Breakfast", 
                   "Morning.snack",
                   "Lunch",
                   "Afternoon.snack",
                   "Dinner",
                   "Before.bedtime.snack",
                   "Midnight.food"))
```

```{r echo=FALSE,fig.height=6, fig.width=14, message=FALSE, warning=FALSE, cache=TRUE}
library(ggthemr)
ggthemr("fresh", layout = "scientific")
library(ggplot2)

pp_long$Time <- as.character(pp_long$Slots)
pp_long$Time[pp_long$Slots == "Breakfast"] <- 7.5
pp_long$Time[pp_long$Slots == "Morning.snack"] <- 10.5
pp_long$Time[pp_long$Slots == "Lunch"] <- 13
pp_long$Time[pp_long$Slots == "Afternoon.snack"] <- 15.5
pp_long$Time[pp_long$Slots == "Dinner"] <- 18.5
pp_long$Time[pp_long$Slots == "Before.bedtime.snack"] <- 21
pp_long$Time[pp_long$Slots == "Midnight.food"] <- 26
pp_long$Time <- as.numeric(pp_long$Time)

ggplot(pp_long[pp_long$CW == 1, ], aes(y = Prob, x=Time, group = Carbo, color = Carbo))  + 
  scale_x_continuous(limits = c(6, 29),
  breaks = seq(6, 29, by = 1)) +
  annotate("rect", xmin = 6, xmax = 9, ymin = 0, ymax = 1.00, 
            alpha = 0.5, fill = "grey") + 
  annotate("rect", xmin = 12, xmax = 14, ymin = 0, ymax = 1.00, 
             alpha = 0.5, fill = "grey") + 
  annotate("rect", xmin = 17, xmax = 20, ymin = 0, ymax = 1.00, 
            alpha = 0.5, fill = "grey") + 
  annotate("rect", xmin = 22, xmax = 29, ymin = 0, ymax = 1.00, 
             alpha = 0.5, fill = "grey") +
  geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
  ) + 
    # theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
  labs(title = "Class 1 days", x = "Hours of the day (next day from hour 24)", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Not eating", "< 50%", ">= 50%")) + 
  ylim(c(0,1)) #+ 
  # scale_x_discrete(labels = c("6","7","8","9","10","11","12","13","14",
  #                             "15","16","17","18","19", "20","21","22",
  #                             "23","0","1","2","3","4","5"))

ggplot(pp_long[pp_long$CW == 2, ], aes(y = Prob, x=Time, group = Carbo, color = Carbo)) + 
  scale_x_continuous(limits = c(6, 29),
  breaks = seq(6, 29, by = 1)) +
  annotate("rect", xmin = 6, xmax = 9, ymin = 0, ymax = 1.00, 
            alpha = 0.5, fill = "grey") + 
  annotate("rect", xmin = 12, xmax = 14, ymin = 0, ymax = 1.00, 
             alpha = 0.5, fill = "grey") + 
  annotate("rect", xmin = 17, xmax = 20, ymin = 0, ymax = 1.00, 
            alpha = 0.5, fill = "grey") + 
  annotate("rect", xmin = 22, xmax = 29, ymin = 0, ymax = 1.00, 
             alpha = 0.5, fill = "grey") +
  geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
  ) + 
    # theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
  labs(title = "Class 2 days", x = "Hours of the day (next day from hour 24)", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Not eating", "< 50%", ">= 50%")) + 
  ylim(c(0,1))

ggplot(pp_long[pp_long$CW == 3, ], aes(y = Prob, x=Time, group = Carbo, color = Carbo)) + 
  scale_x_continuous(limits = c(6, 29),
  breaks = seq(6, 29, by = 1)) +
  annotate("rect", xmin = 6, xmax = 9, ymin = 0, ymax = 1.00, 
            alpha = 0.5, fill = "grey") + 
  annotate("rect", xmin = 12, xmax = 14, ymin = 0, ymax = 1.00, 
             alpha = 0.5, fill = "grey") + 
  annotate("rect", xmin = 17, xmax = 20, ymin = 0, ymax = 1.00, 
            alpha = 0.5, fill = "grey") + 
  annotate("rect", xmin = 22, xmax = 29, ymin = 0, ymax = 1.00, 
             alpha = 0.5, fill = "grey") +
  geom_point(size = 3) + 
  geom_line(size = 1.5) + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 18), 
        axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank(), 
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15) 
        # legend.position = "bottom", 
        # legend.direction = "horizontal"
  ) + 
    # theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
  labs(title = "Class 3 days", x = "Hours of the day (next day from hour 24)", y = "Probability",
       color = "Carbohydrate\nintake") + 
  scale_color_discrete(labels = c("Not eating", "< 50%", ">= 50%")) + 
  ylim(c(0,1))

```

## Visualisation of level 2 latent classes (CB = 2)

```{r echo=FALSE, fig.height=6, fig.width=7, message=FALSE, warning=FALSE, cache=TRUE}

# ndnsNNN <- read_csv("results/ndnsNNN.dat")
# dt <- ndnsNNN[, c(2,5)]
# names(dt) <- c("ID_DAY", "DayofWeek")
# 
# 
# CW2CB2 <- CW2CB2 %>% 
#   left_join(dt, by = "ID_DAY")
#

chart.data <- CW3CB2 %>% 
  group_by(CB, CW) %>% 
  tally %>% 
  group_by(CB) %>% 
  mutate(pct = n/sum(n))

chart.data <- ddply(chart.data, .(CB),
                     transform, pos = cumsum(pct) - (0.5 * pct))

chart.data$CW <- factor(chart.data$CW, levels = c("3", "2", "1"), 
                        labels = c("Class 3 days", "Class 2 days", "Class 1 days"))
chart.data$CB <- factor(chart.data$CB, levels = c("1", "2"), 
                        labels = c("Individual Class 1\n(35.3%)",  "Individual class 2\n(64.7%)"))

library(ggthemr)
ggthemr("dust", layout = "scientific")
ggplot() + 
  geom_bar(aes(y = pct, x = CB, fill = CW), data = chart.data, width = 0.6,
           stat="identity") + 
  geom_text(data=chart.data, aes(x = CB, y = pos, label = paste0(sprintf("%1.1f", pct*100),"%")),
            size=4, colour="white", family="Atlas Grotesk Medium") + 
  theme(legend.position="bottom", legend.direction="horizontal",
        legend.title = element_blank(), 
        axis.title = element_text(size = 18), 
        axis.text = element_text(size = 15), 
        axis.line = element_line(colour = "black"), 
        plot.title=element_text(family="Atlas Grotesk Medium"),
        text=element_text(family="Atlas Grotesk Light")) + 
  labs(title = "Multilevel latent class solution", x = "Between Individual classes", y = "Percentage") +
  scale_y_continuous(labels=percent)
```

## Visualisation of level 2 latent classes (CB = 3)

```{r echo=FALSE, fig.height=6, fig.width=7, message=FALSE, warning=FALSE, cache=TRUE}

CW3CB3 <- read_table2("/home/wangcc-me/Documents/LSHTMproject/results/Timeslots/NDNSslot_CW3CB3.txt",
                         col_names = FALSE)

names(CW3CB3) <- c("Breakfast", 
                   "Morning.snack",
                   "Lunch",
                   "Afternoon.snack",
                   "Dinner",
                   "Before.bedtime.snack",
                   "Midnight.food",
                      "ID_DAY",
                      "AGE",
                      "SEX",
                      "CPROB1",
                      "CPROB2",
                      "CPROB3",
                      "CPROB4",
                      "CPROB5",
                      "CPROB6",
                      "CPROB7",
                      "CPROB8",
                      "CPROB9",
                      "CB",
                      "CW",
                      "MLCJOINT",
                      "ID")

chart.data <- CW3CB3 %>% 
  group_by(CB, CW) %>% 
  tally %>% 
  group_by(CB) %>% 
  mutate(pct = n/sum(n))

# chart.data$CW[chart.data$CW == 1] <- 0 
# chart.data$CW[chart.data$CW == 3] <- 1 
# chart.data$CW[chart.data$CW == 0] <- 3

chart.data <- chart.data[order(chart.data$CB, chart.data$CW),]
chart.data <- ddply(chart.data, .(CB), 
                    transform, pos = cumsum(pct) - (0.5 * pct))

chart.data$CW <- factor(chart.data$CW, levels = c("3", "2", "1"), 
                        labels = c("Class 3 days", "Class 2 days", "Class 1 days"))
chart.data$CB <- factor(chart.data$CB, levels = c("1", "2", "3"), 
                        labels = c("Individual\nclass 1\n(28.2%)",  "Individual\nclass 2\n(28.9%)", 
                                   "Individual\nclass 3\n(42.8%)"))

library(ggthemr)
ggthemr("dust", layout = "scientific")
ggplot() + 
  geom_bar(aes(y = pct, x = CB, fill = CW), data = chart.data, width = 0.6,
           stat="identity") + 
  geom_text(data=chart.data, aes(x = CB, y = pos, label = paste0(sprintf("%1.1f", pct*100),"%")),
            size=4, colour="white", family="Atlas Grotesk Medium") + 
  theme(legend.position="bottom", legend.direction="horizontal",
        legend.title = element_blank(), 
        axis.title = element_text(size = 18), 
        axis.text = element_text(size = 15), 
        axis.line = element_line(colour = "black"), 
        plot.title=element_text(family="Atlas Grotesk Medium"),
        text=element_text(family="Atlas Grotesk Light")) + 
  labs(title = "Multilevel latent class solution", x = "Between Individual classes", y = "Percentage") +
  scale_y_continuous(labels=percent)
```

## Visualisation of level 2 latent classes (CB = 4)

```{r echo=FALSE, fig.height=6, fig.width=7, message=FALSE, warning=FALSE, cache=TRUE}

CW3CB4 <- read_table2("/home/wangcc-me/Documents/LSHTMproject/results/Timeslots/NDNSslot_CW3CB4.txt",
                         col_names = FALSE)

names(CW3CB4) <- c("Breakfast", 
                   "Morning.snack",
                   "Lunch",
                   "Afternoon.snack",
                   "Dinner",
                   "Before.bedtime.snack",
                   "Midnight.food",
                      "ID_DAY",
                      "AGE",
                      "SEX",
                      "CPROB1",
                      "CPROB2",
                      "CPROB3",
                      "CPROB4",
                      "CPROB5",
                      "CPROB6",
                      "CPROB7",
                      "CPROB8",
                      "CPROB9",
                      "CPROB10",
                      "CPROB11",
                      "CPROB12",
                      "CB",
                      "CW",
                      "MLCJOINT",
                      "ID")

chart.data <- CW3CB4 %>% 
  group_by(CB, CW) %>% 
  tally %>% 
  group_by(CB) %>% 
  mutate(pct = n/sum(n))

# chart.data$CW[chart.data$CW == 1] <- 0 
# chart.data$CW[chart.data$CW == 3] <- 1 
# chart.data$CW[chart.data$CW == 0] <- 3

chart.data <- chart.data[order(chart.data$CB, chart.data$CW),]
chart.data <- ddply(chart.data, .(CB), 
                    transform, pos = cumsum(pct) - (0.5 * pct))

chart.data$CW <- factor(chart.data$CW, levels = c("3", "2", "1"), 
                        labels = c("Class 3 days", "Class 2 days", "Class 1 days"))
chart.data$CB <- factor(chart.data$CB, levels = c("1", "2", "3", "4"), 
                        labels = c("Individual\nclass 1\n(21.8%)",  "Individual\nclass 2\n(21.1%)", 
                                   "Individual\nclass 3\n(33.8%)", "Individual\nclass 4\n(23.3%)"))

library(ggthemr)
ggthemr("dust", layout = "scientific")
ggplot() + 
  geom_bar(aes(y = pct, x = CB, fill = CW), data = chart.data, width = 0.6,
           stat="identity") + 
  geom_text(data=chart.data, aes(x = CB, y = pos, label = paste0(sprintf("%1.1f", pct*100),"%")),
            size=4, colour="white", family="Atlas Grotesk Medium") + 
  theme(legend.position="bottom", legend.direction="horizontal",
        legend.title = element_blank(), 
        axis.title = element_text(size = 18), 
        axis.text = element_text(size = 15), 
        axis.line = element_line(colour = "black"), 
        plot.title=element_text(family="Atlas Grotesk Medium"),
        text=element_text(family="Atlas Grotesk Light")) + 
  labs(title = "Multilevel latent class solution", x = "Between Individual classes", y = "Percentage") +
  scale_y_continuous(labels=percent)
```