### 2018-08-01 paper/Tables/2by2Table1.tex
tables replaced with time slots versions

\documentclass[11pt,a4paper]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
    \setmainfont[]{Arial}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=0.8in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Tables in the project},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Tables in the project}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}

\begin{document}
\maketitle

\rowcolors{2}{gray!6}{white}

\begin{table}

\caption{\label{tab:unnamed-chunk-1}Means, percentages, and 95 percent CIs of the characteristics by latent class memberships in the UK adults (NDNS RP 2008/09-2015/16, sample size = 6155).}
\centering
\fontsize{9}{11}\selectfont
\begin{tabular}[t]{llll}
\hiderowcolors
\toprule
Variables & Latent class = 1 & Latent class = 2 & P value \textsuperscript{*}\\
\midrule
\showrowcolors
Total (\%) & 66.4  (64.8 , 68.1) & 33.6  (31.9 , 35.2) & \\
Countries (\%) &  &  & \\
\hspace{1em}England & 83.1  (80.9 , 85.1) & 85.4  (83.0 , 87.5) & 0.203\\
\hspace{1em}Northern Ireland & 3.0  (2.3 , 3.9) & 2.3  (1.7 , 3.1) & \\
\hspace{1em}Scotland & 8.9  (7.2 , 11.0) & 7.7  (5.9 , 9.9) & \\
\hspace{1em}Wales & 5.0  (4.1 , 6.0) & 4.6  (3.6 , 6.0) & \\
Age (years) & 47.3 (46.4, 48.2) & 50.1 (49.1, 51.1) & < 0.001\\
Sex (\%) &  &  & \\
\hspace{1em}Men & 47.3  (45.2 , 49.5) & 51.0  (48.1 , 53.9) & 0.048\\
\hspace{1em}Women & 52.7  (50.5 , 54.8) & 49.0  (46.1 , 51.9) & \\
Survey years (\%) &  &  & \\
\hspace{1em}1 & 13.1 (10.8, 15.8) & 15.0 (12.1, 18.5) & 0.365\\
\hspace{1em}2 & 12.3 (10.1, 14.9) & 11.9 (9.4, 14.2) & \\
\hspace{1em}3 & 12.0 (9.8, 14.8) & 10.4 (8.2, 13.1) & \\
\hspace{1em}4 & 13.2 (10.9, 16.0) & 11.4 (9.0, 14.2) & \\
\hspace{1em}5 & 13.0 (10.8, 15.7) & 13.9 (11.1, 17.3) & \\
\hspace{1em}6 & 12.0 (9.9, 14.4) & 11.1 (8.8, 13.9) & \\
\hspace{1em}7 & 12.3 (10.1,14.9) & 13.5 (10.9, 16.7) & \\
\hspace{1em}8 & 12.0 (9.9, 14.5) & 12.8 (10.3, 15.8) & \\
BMI (kg/m\textasciicircum{}2) & 27.2 (26.9, 27.5) & 27.7 (27.5, 28.1) & 0.007\\
WC (cm) & 92.6 (91.9, 93.3) & 94.2 (93.1, 95.2) & 0.013\\
Smoking status (\%) &  &  & \\
\hspace{1em}Current & 19.8  (18.2 , 21.4) & 23.8 (21.3 , 26.5) & < 0.001\\
\hspace{1em}Ex-smoker & 22.5  (20.8 , 24.2) & 28.0 (25.4 , 30.7) & \\
\hspace{1em}Never & 57.8  (55.7 , 59.8) & 48.2 (45.3 , 51.1) & \\
Current drinking status (\%) &  &  & \\
\hspace{1em}Yes & 23.1  (21.3 , 25.1) & 11.1  (9.5 , 13.0) & < 0.001\\
Hypertension (\%) \textsuperscript{\dag} &  &  & \\
\hspace{1em}Yes & 27.4  (25.0 , 29.9) & 32.6  (29.3 , 36.1) & 0.012\\
Total energy intake (KJ) & 7425.7 (7323.7, 7527.8) & 8149.9 (7997.3, 8302.7) & < 0.001\\
Carbohydrate intake (g) & 226.0 (222.8, 229.3) & 210 (206.1, 213.9) & < 0.001\\
Carbohydrate
percent (\%) \textsuperscript{\ddag} & 48.3 (48.0, 48.6) & 40.9 (40.6, 41.3) & < 0.001\\
Glucose (mmol/l) & 5.13 (5.09, 5.17) & 5.17 (5.11, 5.22) & 0.292\\
A1C (\%) & 5.49 (5.47, 5.52) & 5.47 (5.44,5.51) & 0.264\\
DM \textsuperscript{\S} & 4.1  (3.1 , 5.3 ) & 5.9  (4.3 , 8.0 ) & 0.061\\
Physical
activity (hours/day) \textsuperscript{\P} & 1.51 (1.39, 1.63) & 1.64 (1.45, 1.82) & 0.244\\
\bottomrule
\multicolumn{4}{l}{\textit{Note: }}\\
\multicolumn{4}{l}{Abbreviations: CI, confidence intervals; NDNS RP, national dietary and nutrition survey}\\
\multicolumn{4}{l}{rolling programme; BMI body mass index; WC, waist circumference; A1C, haemoglobin A1c;}\\
\multicolumn{4}{l}{DM, diabetes mellitus.}\\
\multicolumn{4}{l}{Variables from the blood tests (glucose and A1C) are weighted by blood sample weights,}\\
\multicolumn{4}{l}{the others are weighted by individual weights.}\\
\multicolumn{4}{l}{Glucose and A1C levels are estimated in subgroups of people without diabetes.}\\
\multicolumn{4}{l}{\textsuperscript{*} For continuous variables, the F test was used to determine differences between latent classes.}\\
\multicolumn{4}{l}{For categorical variables, differences between latent classes were assessed using the adjusted}\\
\multicolumn{4}{l}{Pearson Chi-2 test for survey data.}\\
\multicolumn{4}{l}{\textsuperscript{\dag} Hypertension was defined as either systolic blood pressure >= 140 mmHg or diastolic blood}\\
\multicolumn{4}{l}{pressure >= 90 mmHg, or under treatment for hypertension.}\\
\multicolumn{4}{l}{\textsuperscript{\ddag} Carbohydrate percent indicates the percentage of energy from carbohydrate in total energy intake}\\
\multicolumn{4}{l}{\textsuperscript{\S} DM was defined by A1C > 6.5\%.}\\
\multicolumn{4}{l}{\textsuperscript{\P} Physical activity was calculated as mean time spent at moderate or vigorous physical activity}\\
\multicolumn{4}{l}{during the survey.}\\
\end{tabular}
\end{table}

\rowcolors{2}{white}{white}

\rowcolors{2}{gray!6}{white}

\begin{table}

\caption{\label{tab:unnamed-chunk-2}Level 1 class selection. (All data, n = 6155, 24483 data points)}
\centering
\fontsize{9}{11}\selectfont
\begin{tabular}[t]{cccccccc}
\hiderowcolors
\toprule
N of classes & N of free parameters & log-likelihood & AIC & BIC & aBIC & Entropy & Lo-Mendel-Rubin LRT\\
\midrule
\showrowcolors
1 & 48 & -372017.3 & 744130.6 & 744519.7 & 744367.1 & -- & --\\
2 & 97 & -368913.7 & 738021.4 & 738807.7 & 738499.4 & 0.777 & < 0.0001\\
3 & 146 & -366665.0 & 733621.9 & 734805.4 & 734341.4 & 0.666 & < 0.0001\\
4 & 195 & -365528.6 & 731447.1 & 733027.7 & 732408.0 & 0.658 & 0.8478\\
5 & 244 & -364901.2 & 730290.3 & 732268.1 & 731492.7 & 0.648 & 0.7602\\
6 & 293 & -363641.8 & 727869.5 & 730244.5 & 729313.4 & 0.701 & 0.7632\\
7 & 342 & -362789.9 & 726263.7 & 729035.9 & 727949.0 & 0.729 & 0.7702\\
8 & 391 & -362047.9 & 724877.9 & 728047.2 & 726804.6 & 0.737 & 0.8261\\
\bottomrule
\multicolumn{8}{l}{\textit{Note: }}\\
\multicolumn{8}{l}{Abbreviation: N, number; AIC, Akaike information criterion; BIC, Bayesian information criterion; aBIC, adjusted BIC;}\\
\multicolumn{8}{l}{Entropy, a pseudo-r-squared index; Lo-Mendel-Rubin LRT, likelihood ratio test comparing q classes models with}\\
\multicolumn{8}{l}{q-1 classes models.}\\
\end{tabular}
\end{table}

\rowcolors{2}{white}{white}

\rowcolors{2}{gray!6}{white}

\begin{table}

\caption{\label{tab:unnamed-chunk-3}Means, percentages, and 95 percent CIs of the characteristics by latent class memberships in the UK adults (NDNS RP 2008/09-2015/16, sample size = 6155).}
\centering
\fontsize{9}{11}\selectfont
\begin{tabular}[t]{lllll}
\hiderowcolors
\toprule
Variables & Latent class = 1 & Latent class = 2 & Latent class = 3 & P value \textsuperscript{*}\\
\midrule
\showrowcolors
Total (\%) & 28.4 (26.8, 30.1) & 7.0 (6.2, 7.9) & 64.6 (62.9, 66.2) & \\
Countries (\%) &  &  &  & \\
\hspace{1em}England & 81.4 (78.5, 84.0) & 87.5 (82.9, 91.0) & 84.6 (82.5, 86.4) & 0.004\\
\hspace{1em}Northern Ireland & 3.9 (2.9, 5.1) & 0.6 (0.3, 1.2) & 2.5 (2.0, 3.2) & \\
\hspace{1em}Scotland & 9.5 (7.4, 12.3) & 6.2 (3.5, 10.6) & 8.3 (6.7, 10.3) & \\
\hspace{1em}Wales & 5.2 (4.1, 6.6) & 5.7 (3.8, 8.5) & 4.6 (3.8, 5.6) & \\
Age (years) & 43.8 (42.4, 45.1) & 49.1 (47.2, 50.9) & 50.1 (49.3, 50.9) & < 0.001\\
Sex (\%) &  &  &  & \\
\hspace{1em}Men & 50.6 (47.3, 53.9) & 49.6 (43.7, 55.4) & 47.6 (45.4, 49.7) & 0.273\\
\hspace{1em}Women & 49.4 (46.1, 52.7) & 50.4 (44.6, 56.3) & 52.4 (50.3, 54.6) & \\
Survey years (\%) &  &  &  & \\
\hspace{1em}1 & 11.4 (8.8, 14.7) & 17.1 (12.4, 23.3) & 14.4 (11.9, 17.4) & 0.002\\
\hspace{1em}2 & 10.1 (7.8, 13.1) & 18.3 (13.3, 24.7) & 12.4 (10.2, 15.0) & \\
\hspace{1em}3 & 13.9 (10.8, 17.7) & 9.1 (5.7, 14.1) & 10.7 (8.6, 13.1) & \\
\hspace{1em}4 & 10.9 (8.5, 13.9) & 13.8 (9.7, 19.4) & 13.2 (10.9, 16.0) & \\
\hspace{1em}5 & 13.5 (10.6, 17.0) & 12.8 (8.4, 19.1) & 13.3 (11.0, 16.1) & \\
\hspace{1em}6 & 12.8 (10.1, 16.1) & 8.7 (5.7, 12.9) & 11.5 (9.5, 13.9) & \\
\hspace{1em}7 & 14.3 (11.5, 17.6) & 9.5 (6.5, 13.8) & 12.4 (10.2, 15.0) & \\
\hspace{1em}8 & 13.2 (10.5, 16.4) & 10.5 (7.4, 14.8) & 12.1 (9.9, 14.6) & \\
BMI (kg/m\textasciicircum{}2) & 27.5 (27.1, 27.9) & 27.0 (26.4, 27.6) & 27.4 (27.2, 27.6) & 0.433\\
WC (cm) & 93.3 (92.1, 94.5) & 92.9 (90.9, 95.0) & 93.1 (92.3, 93.8) & 0.928\\
Smoking status (\%) &  &  &  & \\
\hspace{1em}Current & 24.1 (21.5, 27.0) & 30.0 (24.8, 35.8) & 18.8 (17.2, 20.6) & < 0.001\\
\hspace{1em}Ex-smoker & 20.0 (17.6, 22.6) & 27.5 (22.4, 33.2) & 25.9 (24.1, 27.7) & \\
\hspace{1em}Never & 55.9 (52.7, 59.0) & 42.5 (36.6, 48.7) & 55.3 (53.2, 57.4) & \\
Current drinking status (\%) &  &  &  & \\
\hspace{1em}Yes & 24.6 (21.7, 27.7) & 18.3 (14.0, 23.6) & 16.8 (15.3, 18.4) & < 0.001\\
Hypertension (\%) \textsuperscript{\dag} &  &  &  & \\
\hspace{1em}Yes & 25.9 (22.3, 29.9) & 31.8 (25.3, 39.1) & 30.4 (27.9 32.8) & 0.111\\
Total energy intake (KJ) & 6713.8 (6575.7, 6851.8) & 9256.0 (8850.8, 9661.2) & 7916.9 (7814.0, 8019.9) & < 0.001\\
Carbohydrate intake (g) & 192.9 (188.5, 197.3) & 275.6 (263.4, 287.8) & 226.9 (223.9, 229.9) & < 0.001\\
Carbohydrate
percent (\%) \textsuperscript{\ddag} & 45.8 (45.3, 46.4) & 47.4 (46.5, 48.3) & 45.6 (45.3, 45.9) & 0.001\\
Glucose (mmol/l) & 5.16 (5.08, 5.23) & 5.09 (5.00, 5.18) & 5.14 (5.10, 5.19) & 0.537\\
A1C (\%) & 5.47 (5.42, 5.51) & 5.48 (5.42, 5.54) & 5.49 (5.47. 5.52) & 0.403\\
DM \textsuperscript{\S} & 5.9 (4.2, 8.2) & 1.1 (0.2, 5.2) & 4.7 (3.6, 6.0) & 0.053\\
Physical
activity (hours/day) \textsuperscript{\P} & 1.31 (1.14, 1.49) & 1.82 (1.44, 2.19) & 1.62 (1.49, 1.76) & 0.018\\
\bottomrule
\multicolumn{5}{l}{\textit{Note: }}\\
\multicolumn{5}{l}{Abbreviations: CI, confidence intervals; NDNS RP, national dietary and nutrition survey}\\
\multicolumn{5}{l}{rolling programme; BMI body mass index; WC, waist circumference; A1C, haemoglobin A1c;}\\
\multicolumn{5}{l}{DM, diabetes mellitus.}\\
\multicolumn{5}{l}{Variables from the blood tests (glucose and A1C) are weighted by blood sample weights,}\\
\multicolumn{5}{l}{the others are weighted by individual weights.}\\
\multicolumn{5}{l}{Glucose and A1C levels are estimated in subgroups of people without diabetes.}\\
\multicolumn{5}{l}{\textsuperscript{*} For continuous variables, the F test was used to determine differences between latent classes with Bonferroni correction to}\\
\multicolumn{5}{l}{account for multiple testing across >2 classes. For categorical variables, differences between latent classes were assessed}\\
\multicolumn{5}{l}{using the adjusted Pearson Chi-2 test for survey data.}\\
\multicolumn{5}{l}{\textsuperscript{\dag} Hypertension was defined as either systolic blood pressure >= 140 mmHg or diastolic blood pressure >= 90 mmHg, or under}\\
\multicolumn{5}{l}{treatment for hypertension.}\\
\multicolumn{5}{l}{\textsuperscript{\ddag} Carbohydrate percent indicates the percentage of energy from carbohydrate in total energy intake}\\
\multicolumn{5}{l}{\textsuperscript{\S} DM was defined by A1C > 6.5\%.}\\
\multicolumn{5}{l}{\textsuperscript{\P} Physical activity was calculated as mean time spent at moderate or vigorous physical activity during the survey.}\\
\end{tabular}
\end{table}

\rowcolors{2}{white}{white}

\rowcolors{2}{gray!6}{white}

\begin{table}

\caption{\label{tab:unnamed-chunk-4}Fit Criteria for Each Model Specification}
\centering
\fontsize{9}{11}\selectfont
\begin{tabular}[t]{cccccc}
\hiderowcolors
\toprule
\multicolumn{1}{c}{ } & \multicolumn{5}{c}{Number of level 1 classes} \\
\cmidrule(l{2pt}r{2pt}){2-6}
Model & 1 class & 2 classes & 3 classes & 4 classes & 5 classes\\
\midrule
\showrowcolors
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Fixed effects model}}\\
\hspace{1em}No. of free parameters & 14 & 29 & 44 & 59 & 74\\
\hspace{1em}\hspace{1em}Log-likelihood & -173793.306 & -172669.771 & -172039.204 & -171633.941 & -171377.292\\
\hspace{1em}\hspace{1em}BIC & 347728.092 & 345632.608 & 344523.06 & 343864.121 & 343502.409\\
\hspace{1em}\hspace{1em}Lo-Mendell-Rubun LRT & -- & < 0.0001 & 1e-04 & < 0.0001 & < 0.0001\\
\hspace{1em}\hspace{1em}Entropy & 1 & 0.31 & 0.392 & 0.51 & 0.481\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Random effects model}}\\
\hspace{1em}2 between classes &  &  &  &  & \\
\hspace{1em}\hspace{1em}No. of free parameters &  & 59 & 89 & 119 & \\
\hspace{1em}\hspace{1em}Log-likelihood &  & -169331.132 & -168700.96 & -168366.193 & \\
\hspace{1em}\hspace{1em}BIC &  & 339258.502 & 338301.338 & 337934.968 & \\
\hspace{1em}\hspace{1em}Entropy &  & 0.581 & 0.569 & 0.555 & \\
\hspace{1em}3 between classes &  &  &  &  & \\
\hspace{1em}\hspace{1em}No. of free parameters &  & 89 & 134 & 179 & \\
\hspace{1em}\hspace{1em}Log-likelihood &  & -166936.279 & -166348.815 & -166062.761 & \\
\hspace{1em}\hspace{1em}BIC &  & 334771.968 & 334051.799 & 333934.448 & \\
\hspace{1em}\hspace{1em}Entropy &  & 0.677 & 0.63 & 0.644 & \\
\hspace{1em}4 between classes &  &  &  &  & \\
\hspace{1em}\hspace{1em}No. of free parameters &  & 119 & 179 &  & \\
\hspace{1em}\hspace{1em}Log-likelihood &  & -165441.731 & -164845.696 &  & \\
\hspace{1em}\hspace{1em}BIC &  & 332086.045 & 331500.318 &  & \\
\hspace{1em}\hspace{1em}Entropy &  & 0.729 & 0.659 &  & \\
\bottomrule
\multicolumn{6}{l}{\textit{Note: }}\\
\multicolumn{6}{l}{Abbreviation: No, number; BIC, Bayesian information criterion; Entropy, a pseudo-r-squared index; Lo-Mendel-Rubin LRT, likelihood ratio test comparing q classes models with q-1 classes models.}\\
\end{tabular}
\end{table}

\rowcolors{2}{white}{white}

\rowcolors{2}{gray!6}{white}

\begin{table}

\caption{\label{tab:unnamed-chunk-5}Day Level Latent Class Solution for Three-Class Model (No Individual level Model)}
\centering
\fontsize{9}{11}\selectfont
\begin{tabular}[t]{lllllll}
\hiderowcolors
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{1}{c}{High carbohydrate day} & \multicolumn{1}{c}{Lower carbohydrate day} & \multicolumn{1}{c}{Regular day} \\
\cmidrule(l{2pt}r{2pt}){5-5} \cmidrule(l{2pt}r{2pt}){6-6} \cmidrule(l{2pt}r{2pt}){7-7}
Time slots of the day & Responses of carbohydrate intake & Frequency & Precentages (\%) & Class 1 (30.8\%) & Class 2 (39.8\%) & Class 3 (29.5\%)\\
\midrule
\showrowcolors
6 am – 9 am &  &  &  &  &  & \\
 & Not eating & 7655 & 31.2 & 0.129 & 0.45 & 0.32\\
 & Eating, Carbohydrate < 50\% & 4500 & 18.4 & 0.13 & 0.267 & 0.128\\
 & Eating, Carbohydrate >= 50\% & 12328 & 50.4 & 0.741 & 0.283 & 0.552\\
9 am – 12 am &  &  &  &  &  & \\
 & Not eating & 5447 & 22.2 & 0.237 & 0.079 & 0.401\\
 & Eating, Carbohydrate < 50\% & 7227 & 29.5 & 0.158 & 0.492 & 0.173\\
 & Eating, Carbohydrate >= 50\% & 11809 & 48.2 & 0.605 & 0.429 & 0.426\\
12 noon – 2 pm &  &  &  &  &  & \\
 & Not eating & 4783 & 19.5 & 0.156 & 0.356 & 0.019\\
 & Eating, Carbohydrate < 50\% & 11112 & 45.4 & 0.405 & 0.413 & 0.56\\
 & Eating, Carbohydrate >= 50\% & 8588 & 35.1 & 0.439 & 0.231 & 0.421\\
2 pm – 5 pm &  &  &  &  &  & \\
 & Not eating & 6926 & 28.3 & 0.13 & 0.123 & 0.659\\
 & Eating, Carbohydrate < 50\% & 8277 & 33.8 & 0.249 & 0.602 & 0.076\\
 & Eating, Carbohydrate >= 50\% & 9280 & 37.9 & 0.621 & 0.276 & 0.266\\
5 pm – 8 pm &  &  &  &  &  & \\
 & Not eating & 3043 & 12.4 & 0.114 & 0.199 & 0.034\\
 & Eating, Carbohydrate < 50\% & 14240 & 58.2 & 0.516 & 0.59 & 0.639\\
 & Eating, Carbohydrate >= 50\% & 7200 & 29.4 & 0.37 & 0.211 & 0.328\\
8 pm – 10 pm &  &  &  &  &  & \\
 & Not eating & 8722 & 35.6 & 0.322 & 0.291 & 0.48\\
 & Eating, Carbohydrate < 50\% & 8898 & 36.3 & 0.266 & 0.551 & 0.212\\
 & Eating, Carbohydrate >= 50\% & 6863 & 28.0 & 0.412 & 0.158 & 0.308\\
10 pm – 6 am &  &  &  &  &  & \\
 & Not eating & 16295 & 66.6 & 0.68 & 0.59 & 0.751\\
 & Eating, Carbohydrate < 50\% & 4144 & 16.9 & 0.074 & 0.294 & 0.101\\
 & Eating, Carbohydrate >= 50\% & 4044 & 16.5 & 0.246 & 0.115 & 0.148\\
\bottomrule
\end{tabular}
\end{table}

\rowcolors{2}{white}{white}

\end{document}