### 2018-08-01 results/table1slots.Rmd

---
title: "Summary of Survey Analysis NDNS carb50, 7 time slots"
author: "Chaochen Wang"
date: "30 July 2018 created, `r Sys.Date()` updated"
output:
  html_document:
    nature:
      highlightLines: yes
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(RColorBrewer)
library(plyr)
library(epiDisplay)
library(tidyverse)
library(dplyr)
library(scales)
library(readr)
library(kableExtra)
library(haven)
library(naniar)
```

# Multilevel classes results (3by3)

```{r bigdataset, echo=FALSE, eval=FALSE, warning=FALSE}
# read the big data sets
blood78 <- read_dta("~/Downloads/UKDA-6533-stata11_se/stata11_se/ndns_rp_yr7-8a_indiv.dta")
# blood78 <- read_dta("~/Documents/UKDA-6533-stata11_se/stata11_se/ndns_rp_yr7-8a_indiv.dta")

blood56 <- read_dta("~/Downloads/UKDA-6533-stata11_se/stata11_se/ndns_rp_yr5-6a_indiv.dta")
blood14 <- read_dta("~/Downloads/UKDA-6533-stata11_se/stata11_se/ndns_rp_yr1-4a_indiv_uk.dta")

food14 <- read_dta("~/Downloads/UKDA-6533-stata11_se/stata11_se/ndns_rp_yr1-4a_personleveldietarydata_uk.dta")

food56 <- read_dta("~/Downloads/UKDA-6533-stata11_se/stata11_se/ndns_rp_yr5-6a_personleveldietarydata.dta")

food78 <- read_dta("~/Downloads/UKDA-6533-stata11_se/stata11_se/ndns_rp_yr7-8a_personleveldietarydata.dta")

names(blood78)[names(blood78)=="seriali"] <- "ID"
names(blood56)[names(blood56)=="seriali"] <- "ID"
names(blood14)[names(blood14)=="seriali"] <- "ID"
names(food78)[names(food78)=="seriali"] <- "ID"
names(food56)[names(food56)=="seriali"] <- "ID"
names(food14)[names(food14)=="seriali"] <- "ID"

BMI78 <- blood78 %>% 
  select(ID, Sex, age, bmival, wstval, Diabetes, bpmedc2, bpmedd2, hyper140_2, hibp140_2,  
         Glucose, A1C, cigsta3, dnoft3, dnnow, wti_Y78, wtn_Y78, wtb_Y78, ethgrp5, ethgrp2,
         cluster1, cluster2, cluster3, nssec8, paidemployment, qual7, eqvinc, MarSt2,
         cluster4, cluster5, area, gor) %>% 
  rename(wti = wti_Y78, wtn = wtn_Y78, wtb = wtb_Y78, drink = dnoft3) %>% 
  mutate(Years = "7-8", MVPAtime = NA, MarStat = NA) %>% 
  replace_with_na(replace = list(bmival = -1, qual7 = -8,
                                 wstval = -1, eqvinc = -1,
                                 bpmedd2 = -1, MarSt2 = -1,
                                 bpmedc2 = -1,
                                 hyper140_2 = -7, 
                                 # hyper140_2 = -1, 
                                 hibp140_2 = -7, 
                                 # hibp140_2 = -1, 
                                 Glucose = -1, 
                                 A1C  = -1, 
                                 dnnow = -1,
                                 drink = -1,
                                 ethgrp5 = -4,
                                 ethgrp2 = -4,
                                 cigsta3 = -1, 
                                 nssec8 = -9, 
                                 paidemployment = -1)) %>% 
  replace_with_na(replace = list(hyper140_2 = -1, hibp140_2 = -1, nssec8 = 99, qual7 = -1,
                                 drink = -8, ethgrp2 = -9, ethgrp5 = -9)) %>% 
    replace_with_na(replace = list(drink = -9, ethgrp2 = -8, ethgrp5 = -8,
                                 cigsta3 = -8)) %>% 
  replace_with_na(replace = list(ethgrp2 = -1, ethgrp5 = -1))

BMI56 <- blood56 %>% 
  select(ID, Sex, age, area, bmival, wstval, Diabetes, bpmedc2, bpmedd2, hyper140_2, hibp140_2, MVPAtime,
         Glucose, A1C, cigsta3, dnoft3, dnnow, wti_Y56, wtn_Y56, wtb_Y56, ethgrp5, ethgrp2, qual7,
         cluster1, cluster2, cluster3, nssec8, paidemployment, eqvinc, MarSt2,
         cluster4, cluster5, area, gor) %>% 
  mutate(Years = "5-6", MarStat = NA) %>% 
  rename(wti = wti_Y56, wtn = wtn_Y56, wtb = wtb_Y56, drink = dnoft3) %>% 
  replace_with_na(replace = list(bmival = -1, paidemployment = -9,
                                 wstval = -1, qual7 = -8,
                                 bpmedd2 = -1, eqvinc = -1,
                                 bpmedc2 = -1, MarSt2 = -1,
                                 hyper140_2 = -7, 
                                 # hyper140_2 = -1, 
                                 hibp140_2 = -7, 
                                 # hibp140_2 = -1, 
                                 Glucose = -1, 
                                 A1C  = -1, 
                                 dnnow = -1,
                                 drink = -1,
                                 ethgrp5 = -4,
                                 ethgrp2 = -4,
                                 cigsta3 = -1,
                                 MVPAtime = -1, 
                                 nssec8 = -9)) %>% 
  replace_with_na(replace = list(hyper140_2 = -1, hibp140_2 = -1, paidemployment = -8,
                                 drink = -8, nssec8 = 99)) %>% 
    replace_with_na(replace = list(drink = -9, paidemployment = -1, qual7 = -1,
                                 cigsta3 = -8))

BMI14 <- blood14 %>% 
  select(ID, Sex, age, bmival, wstval, Diabetes, bpmedc, bpmedd, hyper140, hibp140, MVPAtime,
         Glucose, A1C, cigsta3, dnoft3, dnnow, wti_CY1234, wtn_CY1234, wtb_CY1234, ethgr5, ethgr2,
         cluster, area, gor, nssec8, paidemployment, qual7, eqvinc, MarSt2, MarStat) %>%
  rename(hyper140_2 = hyper140, hibp140_2 = hibp140, bpmedd2 = bpmedd, 
         bpmedc2 = bpmedc, cluster1 = cluster, ethgrp5 = ethgr5, ethgrp2 = ethgr2, 
         wti = wti_CY1234, wtn = wtn_CY1234, wtb =  wtb_CY1234, drink = dnoft3) %>% 
  mutate(cluster2 = NA, cluster3 = NA, cluster4 = NA, cluster5 = NA, Years = "1-4") %>% 
  replace_with_na(replace = list(bmival = -1, paidemployment = -9, 
                                 wstval = -1, qual7 = -8,
                                 bpmedd2 = -1, eqvinc = -1,
                                 bpmedc2 = -1, MarSt2 = -4, 
                                 hyper140_2 = -7, MarStat = -4,
                                 # hyper140_2 = -1, 
                                 hibp140_2 = -7, 
                                 # hibp140_2 = -1, 
                                 Glucose = -1, 
                                 A1C  = -1, 
                                 dnnow = -1,
                                 drink = -1,
                                 ethgrp5 = -4,
                                 ethgrp2 = -4,
                                 cigsta3 = -1,
                                 MVPAtime = -4, 
                                 nssec8 = -8)) %>% 
  replace_with_na(replace = list(hyper140_2 = -1, hibp140_2 = -1, MVPAtime = -1, paidemployment = -8,
                                 drink = -8, nssec8 = -1, qual7 = -1, MarSt2 = -1)) %>% 
  replace_with_na(replace = list(drink = -9, paidemployment = -4,
                                 cigsta3 = -8, nssec8 = 99)) %>% 
  replace_with_na(replace = list(paidemployment = -1))

BMI <- bind_rows(BMI14, BMI56, BMI78)

Energy14 <- food14 %>% 
  select(ID, Country, SurveyYear, EnergykJ, Carbohydrateg, CHOpctotE, Proteing, ProteinpctotE, 
         Fatg, FatpctotE)
Energy14$Country[Energy14$Country == "Northern Ireland"] <- "NI"

Energy56 <- food56 %>% 
  select(ID, Country, Surveyyear, EnergykJ, Carbohydrateg, CHOpctotE, Proteing, ProteinpctotE, 
         Fatg, FatpctotE) %>% 
  rename(SurveyYear = Surveyyear)

Energy78 <- food78 %>% 
  select(ID, Country, SurveyYear, EnergykJ, Carbohydrateg, CHOpctotE, Proteing, ProteinpctotE, 
         Fatg, FatpctotE)

Energy <- bind_rows(Energy14, Energy56, Energy78)

```

```{r combinedata, echo=FALSE, eval=FALSE, warning=FALSE}
# ready to combine with the classification data

CW3CB3 <- read_table2("~/Documents/LSHTMproject/results/Timeslots/NDNSslot_CW3CB2.txt",  
                              col_names = FALSE)# change the path to your own path

names(CW3CB3) <- c("Breakfast", "Morning.snack", "Lunch",
                   "Afternoon.snack", "Dinner", "Before.bedtime.snack",
                   "Midnight.food", "ID_DAY", "AGE", "SEX", "CPROB1",
                      "CPROB2", "CPROB3", "CPROB4", "CPROB5", "CPROB6", 
                      "CB", "CW", "MLCJOINT", "ID")

dta_NDNS <- read_csv("/home/wangcc-me/Documents/LSHTMproject/Rcode/dta_NDNS.csv") # extract the day of week data

CW3idday <- CW3CB3 %>% 
  select(ID, ID_DAY, CW, CB, MLCJOINT, AGE, SEX)

CW3idday$DayNo <- ave(CW3idday$ID_DAY, CW3idday$ID, FUN = seq_along) # adding the day no

dta_DayofWeek <- dta_NDNS %>% 
  select(id, id_dy, DayofWeek)

dta_DayofWeek$DayNo <- unlist(strsplit(dta_DayofWeek$id_dy,"D"))[c(FALSE, TRUE)] # creating a day no
names(dta_DayofWeek)[1] <- "ID"
dta_DayofWeek$DayNo <- as.numeric(dta_DayofWeek$DayNo)

CW3idday <-   CW3idday %>%  
  left_join(dta_DayofWeek, by= c("ID", "DayNo"))

CW3idday$DayofWeek[CW3idday$ID_DAY == 40714261000] <- "Saturday"
CW3idday$DayofWeek[CW3idday$ID_DAY == 112050710000] <- "Tuesday"
CW3idday$DayofWeek[CW3idday$ID_DAY == 310122510000] <- "Wednesday"
CW3idday$DayofWeek[CW3idday$ID_DAY == 5050616100] <- "Monday"
CW3idday$DayofWeek[CW3idday$ID_DAY == 50506161000] <- "Tuesday"
CW3idday$DayofWeek[CW3idday$ID_DAY == 505061610000] <- "Wednesday"
CW3idday$DayofWeek[CW3idday$ID_DAY == 7090824100] <- "Monday"
CW3idday$DayofWeek[CW3idday$ID_DAY == 70908241000] <- "Tuesday"
CW3idday$DayofWeek[CW3idday$ID_DAY == 709082410000] <- "Wednesday"
CW3idday$DayofWeek[CW3idday$ID_DAY == 807021910000] <- "Monday"
CW3idday$DayofWeek[CW3idday$ID_DAY == 812101310000] <- "Sunday"

with(CW3idday, tabpct(DayofWeek, CW))
with(CW3idday, chisq.test(DayofWeek, CW))

CW3idday$Weekend <- ifelse((CW3idday$DayofWeek == "Sunday" | CW3idday$DayofWeek == "Saturday"), TRUE, FALSE)

with(CW3idday, tabpct(Weekend, CW))
with(CW3idday, chisq.test(Weekend, CW))

CW3CB3_reg <- CW3CB3[!duplicated(CW3CB3$ID), ]

CW3CB3_reg <- CW3CB3_reg %>% 
  select(ID, AGE, SEX, CB) # extract only the CB variable (Between individual classes == 1 or 2)

tab1(CW3CB3_reg$CB, graph = FALSE)

CW3CB3_regss <- CW3CB3_reg %>% 
  left_join(BMI, by = "ID") %>% 
  left_join(Energy, by = "ID")## dataset for 3by3 multilevel latent classes

Slots_LGCA_3CLASS <- read_table2("/home/wangcc-me/Documents/LSHTMproject/results/Timeslots/LGCA/NDNSslot_LGCA_3CLASS.txt",  col_names = FALSE)

Slots_LGCA_3CLASS[, 1:32] <- sapply(Slots_LGCA_3CLASS[, 1:32], as.numeric)
names(Slots_LGCA_3CLASS) <- c(
  "H6_9_X",  "H9_12_X",
  "H12_14_X",  "H14_17_X",
  "H17_20_X",  "H20_22_X",
  "H22_6_X",  
  "H6_9_Y",  "H9_12_Y",
  "H12_14_Y",  "H14_17_Y",
  "H17_20_Y",  "H20_22_Y",
  "H22_6_Y",  
  "H6_9_X_X",  "H9_12_X_X",
  "H12_14_X_X",  "H14_17_X_X",
  "H17_20_X_X",  "H20_22_X_X",
  "H22_6_X_X",  
  "H6_9_Y_Y",  "H9_12_Y_Y",
  "H12_14_Y_Y",  "H14_17_Y_Y",
  "H17_20_Y_Y",  "H20_22_Y_Y",
  "H22_6_Y_Y",  
  "ID",
  "CPROB1",  "CPROB2",
  "CPROB3",  #"CPROB4", 
  "C")

Slots_LGCA_3CLASS[Slots_LGCA_3CLASS == "*"] <- NA

LCGA_3class <- Slots_LGCA_3CLASS %>% 
  select(ID, C) # extract only the C variable (classes == 1 or 2)

tab1(LCGA_3class$C, graph = FALSE)

LCGA_3class <- LCGA_3class %>% 
  left_join(BMI, by = "ID") %>% 
  left_join(Energy, by = "ID")## dataset for 3 classes latent classes growth analysis

```

```{r rescale the weighting,eval=FALSE,message=FALSE, cache=TRUE, warning=FALSE}
# Rescale the weighting values
# individual weighting
a <- sum(CW3CB3_regss[CW3CB3_regss$Years == "1-4",]$wti)
b <- sum(CW3CB3_regss[CW3CB3_regss$Years == "5-6",]$wti)
c <- sum(CW3CB3_regss[CW3CB3_regss$Years == "7-8",]$wti)

CW3CB3_regss$wti1to8 <- CW3CB3_regss$wti

CW3CB3_regss[CW3CB3_regss$Years == "1-4",]$wti1to8 <- CW3CB3_regss[CW3CB3_regss$Years == "1-4",]$wti*(a+b+c)*(1/2)/a
CW3CB3_regss[CW3CB3_regss$Years == "5-6",]$wti1to8 <- CW3CB3_regss[CW3CB3_regss$Years == "5-6",]$wti*(a+b+c)*(1/4)/b
CW3CB3_regss[CW3CB3_regss$Years == "7-8",]$wti1to8 <- CW3CB3_regss[CW3CB3_regss$Years == "7-8",]$wti*(a+b+c)*(1/4)/c
mean(CW3CB3_regss$wti1to8)

CW3CB3_regss$wti1to8 <- CW3CB3_regss$wti1to8/1.209816814
summ(CW3CB3_regss$wti1to8, graph = FALSE)
sum(CW3CB3_regss$wti1to8, graph = FALSE) #Check if the weighting sum up to the sample size we have

# Nurse weights

a <- sum(CW3CB3_regss[CW3CB3_regss$Years == "1-4",]$wtn)
b <- sum(CW3CB3_regss[CW3CB3_regss$Years == "5-6",]$wtn)
c <- sum(CW3CB3_regss[CW3CB3_regss$Years == "7-8",]$wtn)

CW3CB3_regss$wtn1to8 <- CW3CB3_regss$wtn

CW3CB3_regss[CW3CB3_regss$Years == "1-4",]$wtn1to8 <- CW3CB3_regss[CW3CB3_regss$Years == "1-4",]$wtn*(a+b+c)*(1/2)/a
CW3CB3_regss[CW3CB3_regss$Years == "5-6",]$wtn1to8 <- CW3CB3_regss[CW3CB3_regss$Years == "5-6",]$wtn*(a+b+c)*(1/4)/b
CW3CB3_regss[CW3CB3_regss$Years == "7-8",]$wtn1to8 <- CW3CB3_regss[CW3CB3_regss$Years == "7-8",]$wtn*(a+b+c)*(1/4)/c
mean(CW3CB3_regss$wtn1to8)

CW3CB3_regss$wtn1to8 <- CW3CB3_regss$wtn1to8/0.907003577
summ(CW3CB3_regss$wtn1to8, graph = FALSE)
sum(CW3CB3_regss$wtn1to8, graph = FALSE) #Check if the weighting sum up to the sample size we have

# Blood weights
a <- sum(CW3CB3_regss[CW3CB3_regss$Years == "1-4",]$wtb)
b <- sum(CW3CB3_regss[CW3CB3_regss$Years == "5-6",]$wtb)
c <- sum(CW3CB3_regss[CW3CB3_regss$Years == "7-8",]$wtb)

CW3CB3_regss$wtb1to8 <- CW3CB3_regss$wtb

CW3CB3_regss[CW3CB3_regss$Years == "1-4",]$wtb1to8 <- CW3CB3_regss[CW3CB3_regss$Years == "1-4",]$wtb*(a+b+c)*(1/2)/a
CW3CB3_regss[CW3CB3_regss$Years == "5-6",]$wtb1to8 <- CW3CB3_regss[CW3CB3_regss$Years == "5-6",]$wtb*(a+b+c)*(1/4)/b
CW3CB3_regss[CW3CB3_regss$Years == "7-8",]$wtb1to8 <- CW3CB3_regss[CW3CB3_regss$Years == "7-8",]$wtb*(a+b+c)*(1/4)/c
mean(CW3CB3_regss$wtb1to8)

CW3CB3_regss$wtb1to8 <- CW3CB3_regss$wtb1to8/0.4817444505
summ(CW3CB3_regss$wtb1to8, graph = FALSE)
sum(CW3CB3_regss$wtb1to8, graph = FALSE) #Check if the weighting sum up to the sample size we have

weightings <- CW3CB3_regss %>% select(ID, wti1to8, wtn1to8, wtb1to8)

# add the weightings to the other datasets
# CW3CB3_regss <- CW3CB3_regss %>% 
#   left_join(weightings, by = "ID")

LCGA_3class <- LCGA_3class %>%
  left_join(weightings, by = "ID")

write_dta(LCGA_3class, "/home/wangcc-me/Downloads/UKDA-6533-stata11_se/stata11_se/LCGA_7slots3class.dta")
# write_dta(LCGA_2class, "/home/wangcc-me/Downloads/UKDA-6533-stata11_se/stata11_se/LCGA_2class.dta")
write_dta(CW3CB3_regss, "/home/wangcc-me/Downloads/UKDA-6533-stata11_se/stata11_se/CW3CB3_7regss.dta")
# write_dta(CW2CB2_regss, "/home/wangcc-me/Downloads/UKDA-6533-stata11_se/stata11_se/CW2CB2_regss.dta")
```

## Association between day-level classes and weekdays

```{r dayweek, echo=FALSE, message=FALSE}

CW3CB3 <- read_table2("~/Documents/LSHTMproject/results/Timeslots/NDNSslot_CW3CB2.txt",  
                              col_names = FALSE)# change the path to your own path

names(CW3CB3) <- c("Breakfast", "Morning.snack", "Lunch",
                   "Afternoon.snack", "Dinner", "Before.bedtime.snack",
                   "Midnight.food", "ID_DAY", "AGE", "SEX", "CPROB1",
                      "CPROB2", "CPROB3", "CPROB4", "CPROB5", "CPROB6", 
                      "CB", "CW", "MLCJOINT", "ID")

dta_NDNS <- read_csv("/home/wangcc-me/Documents/LSHTMproject/Rcode/dta_NDNS.csv") # extract the day of week data

CW3idday <- CW3CB3 %>% 
  select(ID, ID_DAY, CW, CB, MLCJOINT, AGE, SEX)

CW3idday$DayNo <- ave(CW3idday$ID_DAY, CW3idday$ID, FUN = seq_along) # adding the day no

dta_DayofWeek <- dta_NDNS %>% 
  select(id, id_dy, DayofWeek)

dta_DayofWeek$DayNo <- unlist(strsplit(dta_DayofWeek$id_dy,"D"))[c(FALSE, TRUE)] # creating a day no
names(dta_DayofWeek)[1] <- "ID"
dta_DayofWeek$DayNo <- as.numeric(dta_DayofWeek$DayNo)

CW3idday <-   CW3idday %>%  
  left_join(dta_DayofWeek, by= c("ID", "DayNo"))

CW3idday$DayofWeek[CW3idday$ID_DAY == 40714261000] <- "Saturday"
CW3idday$DayofWeek[CW3idday$ID_DAY == 112050710000] <- "Tuesday"
CW3idday$DayofWeek[CW3idday$ID_DAY == 310122510000] <- "Wednesday"
CW3idday$DayofWeek[CW3idday$ID_DAY == 5050616100] <- "Monday"
CW3idday$DayofWeek[CW3idday$ID_DAY == 50506161000] <- "Tuesday"
CW3idday$DayofWeek[CW3idday$ID_DAY == 505061610000] <- "Wednesday"
CW3idday$DayofWeek[CW3idday$ID_DAY == 7090824100] <- "Monday"
CW3idday$DayofWeek[CW3idday$ID_DAY == 70908241000] <- "Tuesday"
CW3idday$DayofWeek[CW3idday$ID_DAY == 709082410000] <- "Wednesday"
CW3idday$DayofWeek[CW3idday$ID_DAY == 807021910000] <- "Monday"
CW3idday$DayofWeek[CW3idday$ID_DAY == 812101310000] <- "Sunday"

with(CW3idday, tabpct(DayofWeek, CW))
with(CW3idday, chisq.test(DayofWeek, CW))

CW3idday$Weekend <- ifelse((CW3idday$DayofWeek == "Sunday" | CW3idday$DayofWeek == "Saturday"), TRUE, FALSE)

with(CW3idday, tabpct(Weekend, CW))
with(CW3idday, chisq.test(Weekend, CW))

CW3CB3_reg <- CW3CB3[!duplicated(CW3CB3$ID), ]

CW3CB3_reg <- CW3CB3_reg %>% 
  select(ID, AGE, SEX, CB) # extract only the CB variable (Between individual classes == 1 or 2)
```

## Social economic

```{r tab1, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
dt <- read.csv("/home/wangcc-me/Documents/LSHTMproject/Tablecsv/Timeslots_CW3CB3results.csv", header = T, row.names = NULL, colClasses = "character")
names(dt) <- c("Variables", "Latent class = 1", "Latent class = 2", "Latent class = 3", "P value [note]")

# 
# dt[28, 1] <- c("Hypertension (%) [note]")
# dt[32, 1] <- c("Carbohydrate\npercent (%) [note]")
# dt[35, 1] <- c("DM [note]")
# dt[36, 1] <- c("Physical\nactivity (hours/day) [note]")
dt[20, 1] <- c("Paid employment[note] (%)")
dt[23, 1] <- c("Live with partner[note] (%)")

kable(dt,  align = "l", escape = F, row.names = FALSE,
              format = "html", caption = "Means, row percentages, and 95% CIs of the characteristics by latent class memberships in the UK adults (NDNS RP 2008/09-2015/16, sample size = 6155).") %>%
 kable_styling(font_size = 16, 
        bootstrap_options = c("striped", "hover", "condensed"),
        position = "center", full_width = FALSE) %>% 
  add_indent(c(3:6, 9, 10, 12:19, 21, 22, 24, 25, 28, 29, 31, 32)) %>% 
  footnote(general = c("Abbreviations: CI, confidence intervals; NDNS RP, national dietary and nutrition survey rolling programme;"
                       # " BMI body mass index; WC, waist circumference; A1C, haemoglobin A1c; DM, diabetes mellitus.", 
                       # "Variables from the blood tests (glucose and A1C) are weighted by blood sample weights, the others are weighted by individual weights.", 
                       # "Glucose and A1C levels are estimated in subgroups of people without diabetes."
                       )) %>% 
  add_footnote(c("For continuous variables, the F test was used to determine differences between latent classes with Bonferroni correction to account for multiple testing across >2 classes. For categorical variables, differences between latent classes were assessed using the adjusted Pearson Chi-2 test for survey data.",
                 "Paid employment was defined as being in paid employment during the last 4 weeks prior to the survey." , 
                 "Live with partner was defined as either living with a married husband/wife or a legally recognised civil partnership."
                 # "Hypertension was defined as either systolic blood pressure >= 140 mmHg or diastolic blood pressure >= 90 mmHg, or under treatment for hypertension.", 
                 # "Carbohydrate percent indicates the percentage of energy from carbohydrate in total energy intake",
                 # "DM was defined by A1C > 6.5%.", 
                 # "Physical activity was calculated as mean time spent at moderate or vigorous physical activity during the survey."
                 ), 
               notation = "symbol")
```

## Anthropometric measurements and biochemical analysis

```{r tab2, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
dt <- read.csv("/home/wangcc-me/Documents/LSHTMproject/Tablecsv/Timeslots_CW3CB3table2_bmiDM.csv", header = T, row.names = NULL, colClasses = "character")
names(dt) <- c("Variables", "Latent class = 1", "Latent class = 2", "Latent class = 3", "P value [note]")

# 
dt[9, 1] <- c("Hypertension[note] (%)")
dt[12, 1] <- c("Carbohydrate percent[note] (%)")

dt[21, 1] <- c("Physical\nactivity (hours/day) [note]")
dt[20, 1] <- c("DM [note]")
# dt[20, 1] <- c("Paid employment[note] (%)")
# dt[23, 1] <- c("Live with partner[note] (%)")

kable(dt,  align = "l", escape = F, row.names = FALSE,
              format = "html", caption = "Means, column percentages, and 95% CIs of the characteristics by latent class memberships in the UK adults (NDNS RP 2008/09-2015/16, sample size = 6155).") %>%
 kable_styling(font_size = 16, 
        bootstrap_options = c("striped", "hover", "condensed"),
        position = "center", full_width = FALSE) %>% 
  add_indent(c(4:6,8,10)) %>% 
  footnote(general = c("Abbreviations: CI, confidence intervals; NDNS RP, national dietary and nutrition survey rolling programme;",
                        " BMI body mass index; WC, waist circumference; A1C, haemoglobin A1c; DM, diabetes mellitus.", 
                        "Variables from the blood tests (glucose and A1C) are weighted by blood sample weights, the others are weighted by individual weights.", 
                        "Glucose and A1C levels are estimated in subgroups of people without diabetes.")) %>% 
  add_footnote(c("For continuous variables, the F test was used to determine differences between latent classes with Bonferroni correction to account for multiple testing across >2 classes. For categorical variables, differences between latent classes were assessed using the adjusted Pearson Chi-2 test for survey data.",
                 # "Paid employment was defined as being in paid employment during the last 4 weeks prior to the survey." , 
                 # "Live with partner was defined as either living with a married husband/wife or a legally recognised civil partnership."
                 "Hypertension was defined as either systolic blood pressure >= 140 mmHg or diastolic blood pressure >= 90 mmHg, or under treatment for hypertension.",
                 "Carbohydrate percent indicates the percentage of energy from carbohydrate in total energy intake",
                 "DM was defined by A1C > 6.5%.",
                 "Physical activity was calculated as mean time spent at moderate or vigorous physical activity during the survey."
                 ), 
               notation = "symbol")
```

<!-- # LCGA results (3 classes) -->

<!-- ```{r tab1LCGA, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE} -->
<!-- dt <- read.csv("/home/wangcc-me/Documents/LSHTMproject/Tablecsv/Timeslots_LCGA2classesresults.csv", header = T, row.names = NULL, colClasses = "character") -->
<!-- names(dt) <- c("Variables", "Latent class = 1", "Latent class = 2", "Latent class = 3", "P value [note]") -->

<!-- dt[28, 1] <- c("Hypertension (%) [note]") -->
<!-- dt[32, 1] <- c("Carbohydrate\npercent (%) [note]") -->
<!-- dt[35, 1] <- c("DM [note]") -->
<!-- dt[36, 1] <- c("Physical\nactivity (hours/day) [note]") -->

<!-- kable(dt,  align = "l", escape = F, row.names = FALSE, -->
<!--               format = "html", caption = "Means, percentages, and 95% CIs of the characteristics by latent class memberships in the UK adults (NDNS RP 2008/09-2015/16, sample size = 6155).") %>% -->
<!--  kable_styling(font_size = 16,  -->
<!--         bootstrap_options = c("striped", "hover", "condensed"), -->
<!--         position = "center", full_width = FALSE) %>%  -->
<!--   add_indent(c(3:6, 9, 10, 12:19, 23:25, 27, 29)) %>%  -->
<!--   footnote(general = c("Abbreviations: CI, confidence intervals; NDNS RP, national dietary and nutrition survey rolling programme; BMI body mass index; WC, waist circumference; A1C, haemoglobin A1c; DM, diabetes mellitus.",  -->
<!--                        "Variables from the blood tests (glucose and A1C) are weighted by blood sample weights, the others are weighted by individual weights.",  -->
<!--                        "Glucose and A1C levels are estimated in subgroups of people without diabetes.")) %>%  -->
<!--   add_footnote(c("For continuous variables, the F test was used to determine differences between latent classes with Bonferroni correction to account for multiple testing across >2 classes. For categorical variables, differences between latent classes were assessed using the adjusted Pearson Chi-2 test for survey data.", -->
<!--                  "Hypertension was defined as either systolic blood pressure >= 140 mmHg or diastolic blood pressure >= 90 mmHg, or under treatment for hypertension.",  -->
<!--                  "Carbohydrate percent indicates the percentage of energy from carbohydrate in total energy intake", -->
<!--                  "DM was defined by A1C > 6.5%.",  -->
<!--                  "Physical activity was calculated as mean time spent at moderate or vigorous physical activity during the survey."),  -->
<!--                notation = "symbol") -->
<!-- ``` -->