### 2018-08-03 paper/Tables/2by2Table1.tex

\documentclass[11pt,a4paper]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
    \setmainfont[]{Arial}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=0.8in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Tables in the project},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Tables in the project}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}

\begin{document}
\maketitle

\section{Multilevel 2by2 table 1}\label{multilevel-2by2-table-1}

\rowcolors{2}{gray!6}{white}

\begin{table}

\caption{\label{tab:unnamed-chunk-1}Means, percentages, and 95 percent CIs of the characteristics by latent class memberships in the UK adults (NDNS RP 2008/09-2015/16, sample size = 6155).}
\centering
\fontsize{9}{11}\selectfont
\begin{tabular}[t]{llll}
\hiderowcolors
\toprule
Variables & Latent class = 1 & Latent class = 2 & P value \textsuperscript{*}\\
\midrule
\showrowcolors
Total (\%) & 66.4  (64.8 , 68.1) & 33.6  (31.9 , 35.2) & \\
Countries (\%) &  &  & \\
\hspace{1em}England & 83.1  (80.9 , 85.1) & 85.4  (83.0 , 87.5) & 0.203\\
\hspace{1em}Northern Ireland & 3.0  (2.3 , 3.9) & 2.3  (1.7 , 3.1) & \\
\hspace{1em}Scotland & 8.9  (7.2 , 11.0) & 7.7  (5.9 , 9.9) & \\
\hspace{1em}Wales & 5.0  (4.1 , 6.0) & 4.6  (3.6 , 6.0) & \\
Age (years) & 47.3 (46.4, 48.2) & 50.1 (49.1, 51.1) & < 0.001\\
Sex (\%) &  &  & \\
\hspace{1em}Men & 47.3  (45.2 , 49.5) & 51.0  (48.1 , 53.9) & 0.048\\
\hspace{1em}Women & 52.7  (50.5 , 54.8) & 49.0  (46.1 , 51.9) & \\
Survey years (\%) &  &  & \\
\hspace{1em}1 & 13.1 (10.8, 15.8) & 15.0 (12.1, 18.5) & 0.365\\
\hspace{1em}2 & 12.3 (10.1, 14.9) & 11.9 (9.4, 14.2) & \\
\hspace{1em}3 & 12.0 (9.8, 14.8) & 10.4 (8.2, 13.1) & \\
\hspace{1em}4 & 13.2 (10.9, 16.0) & 11.4 (9.0, 14.2) & \\
\hspace{1em}5 & 13.0 (10.8, 15.7) & 13.9 (11.1, 17.3) & \\
\hspace{1em}6 & 12.0 (9.9, 14.4) & 11.1 (8.8, 13.9) & \\
\hspace{1em}7 & 12.3 (10.1,14.9) & 13.5 (10.9, 16.7) & \\
\hspace{1em}8 & 12.0 (9.9, 14.5) & 12.8 (10.3, 15.8) & \\
BMI (kg/m\textasciicircum{}2) & 27.2 (26.9, 27.5) & 27.7 (27.5, 28.1) & 0.007\\
WC (cm) & 92.6 (91.9, 93.3) & 94.2 (93.1, 95.2) & 0.013\\
Smoking status (\%) &  &  & \\
\hspace{1em}Current & 19.8  (18.2 , 21.4) & 23.8 (21.3 , 26.5) & < 0.001\\
\hspace{1em}Ex-smoker & 22.5  (20.8 , 24.2) & 28.0 (25.4 , 30.7) & \\
\hspace{1em}Never & 57.8  (55.7 , 59.8) & 48.2 (45.3 , 51.1) & \\
Current drinking status (\%) &  &  & \\
\hspace{1em}Yes & 23.1  (21.3 , 25.1) & 11.1  (9.5 , 13.0) & < 0.001\\
Hypertension (\%) \textsuperscript{\dag} &  &  & \\
\hspace{1em}Yes & 27.4  (25.0 , 29.9) & 32.6  (29.3 , 36.1) & 0.012\\
Total energy intake (KJ) & 7425.7 (7323.7, 7527.8) & 8149.9 (7997.3, 8302.7) & < 0.001\\
Carbohydrate intake (g) & 226.0 (222.8, 229.3) & 210 (206.1, 213.9) & < 0.001\\
Carbohydrate
percent (\%) \textsuperscript{\ddag} & 48.3 (48.0, 48.6) & 40.9 (40.6, 41.3) & < 0.001\\
Glucose (mmol/l) & 5.13 (5.09, 5.17) & 5.17 (5.11, 5.22) & 0.292\\
A1C (\%) & 5.49 (5.47, 5.52) & 5.47 (5.44,5.51) & 0.264\\
DM \textsuperscript{\S} & 4.1  (3.1 , 5.3 ) & 5.9  (4.3 , 8.0 ) & 0.061\\
Physical
activity (hours/day) \textsuperscript{\P} & 1.51 (1.39, 1.63) & 1.64 (1.45, 1.82) & 0.244\\
\bottomrule
\multicolumn{4}{l}{\textit{Note: }}\\
\multicolumn{4}{l}{Abbreviations: CI, confidence intervals; NDNS RP, national dietary and nutrition survey}\\
\multicolumn{4}{l}{rolling programme; BMI body mass index; WC, waist circumference; A1C, haemoglobin A1c;}\\
\multicolumn{4}{l}{DM, diabetes mellitus.}\\
\multicolumn{4}{l}{Variables from the blood tests (glucose and A1C) are weighted by blood sample weights,}\\
\multicolumn{4}{l}{the others are weighted by individual weights.}\\
\multicolumn{4}{l}{Glucose and A1C levels are estimated in subgroups of people without diabetes.}\\
\multicolumn{4}{l}{\textsuperscript{*} For continuous variables, the F test was used to determine differences between latent classes.}\\
\multicolumn{4}{l}{For categorical variables, differences between latent classes were assessed using the adjusted}\\
\multicolumn{4}{l}{Pearson Chi-2 test for survey data.}\\
\multicolumn{4}{l}{\textsuperscript{\dag} Hypertension was defined as either systolic blood pressure >= 140 mmHg or diastolic blood}\\
\multicolumn{4}{l}{pressure >= 90 mmHg, or under treatment for hypertension.}\\
\multicolumn{4}{l}{\textsuperscript{\ddag} Carbohydrate percent indicates the percentage of energy from carbohydrate in total energy intake}\\
\multicolumn{4}{l}{\textsuperscript{\S} DM was defined by A1C > 6.5\%.}\\
\multicolumn{4}{l}{\textsuperscript{\P} Physical activity was calculated as mean time spent at moderate or vigorous physical activity}\\
\multicolumn{4}{l}{during the survey.}\\
\end{tabular}
\end{table}

\rowcolors{2}{white}{white}

\section{Day level only models}\label{day-level-only-models}

\rowcolors{2}{gray!6}{white}

\begin{table}

\caption{\label{tab:unnamed-chunk-2}Level 1 class selection. (All data, n = 6155, 24483 data points)}
\centering
\fontsize{9}{11}\selectfont
\begin{tabular}[t]{cccccccc}
\hiderowcolors
\toprule
N of classes & N of free parameters & log-likelihood & AIC & BIC & aBIC & Entropy & Lo-Mendel-Rubin LRT\\
\midrule
\showrowcolors
1 & 48 & -372017.3 & 744130.6 & 744519.7 & 744367.1 & -- & --\\
2 & 97 & -368913.7 & 738021.4 & 738807.7 & 738499.4 & 0.777 & < 0.0001\\
3 & 146 & -366665.0 & 733621.9 & 734805.4 & 734341.4 & 0.666 & < 0.0001\\
4 & 195 & -365528.6 & 731447.1 & 733027.7 & 732408.0 & 0.658 & 0.8478\\
5 & 244 & -364901.2 & 730290.3 & 732268.1 & 731492.7 & 0.648 & 0.7602\\
6 & 293 & -363641.8 & 727869.5 & 730244.5 & 729313.4 & 0.701 & 0.7632\\
7 & 342 & -362789.9 & 726263.7 & 729035.9 & 727949.0 & 0.729 & 0.7702\\
8 & 391 & -362047.9 & 724877.9 & 728047.2 & 726804.6 & 0.737 & 0.8261\\
\bottomrule
\multicolumn{8}{l}{\textit{Note: }}\\
\multicolumn{8}{l}{Abbreviation: N, number; AIC, Akaike information criterion; BIC, Bayesian information criterion; aBIC, adjusted BIC;}\\
\multicolumn{8}{l}{Entropy, a pseudo-r-squared index; Lo-Mendel-Rubin LRT, likelihood ratio test comparing q classes models with}\\
\multicolumn{8}{l}{q-1 classes models.}\\
\end{tabular}
\end{table}

\rowcolors{2}{white}{white}

\section{LCGA models table 1}\label{lcga-models-table-1}

\rowcolors{2}{gray!6}{white}

\begin{table}

\caption{\label{tab:unnamed-chunk-3}Means, percentages, and 95 percent CIs of the characteristics by latent class memberships in the UK adults (NDNS RP 2008/09-2015/16, sample size = 6155).}
\centering
\fontsize{9}{11}\selectfont
\begin{tabular}[t]{lllll}
\hiderowcolors
\toprule
Variables & Latent class = 1 & Latent class = 2 & Latent class = 3 & P value \textsuperscript{*}\\
\midrule
\showrowcolors
Total (\%) & 28.4 (26.8, 30.1) & 7.0 (6.2, 7.9) & 64.6 (62.9, 66.2) & \\
Countries (\%) &  &  &  & \\
\hspace{1em}England & 81.4 (78.5, 84.0) & 87.5 (82.9, 91.0) & 84.6 (82.5, 86.4) & 0.004\\
\hspace{1em}Northern Ireland & 3.9 (2.9, 5.1) & 0.6 (0.3, 1.2) & 2.5 (2.0, 3.2) & \\
\hspace{1em}Scotland & 9.5 (7.4, 12.3) & 6.2 (3.5, 10.6) & 8.3 (6.7, 10.3) & \\
\hspace{1em}Wales & 5.2 (4.1, 6.6) & 5.7 (3.8, 8.5) & 4.6 (3.8, 5.6) & \\
Age (years) & 43.8 (42.4, 45.1) & 49.1 (47.2, 50.9) & 50.1 (49.3, 50.9) & < 0.001\\
Sex (\%) &  &  &  & \\
\hspace{1em}Men & 50.6 (47.3, 53.9) & 49.6 (43.7, 55.4) & 47.6 (45.4, 49.7) & 0.273\\
\hspace{1em}Women & 49.4 (46.1, 52.7) & 50.4 (44.6, 56.3) & 52.4 (50.3, 54.6) & \\
Survey years (\%) &  &  &  & \\
\hspace{1em}1 & 11.4 (8.8, 14.7) & 17.1 (12.4, 23.3) & 14.4 (11.9, 17.4) & 0.002\\
\hspace{1em}2 & 10.1 (7.8, 13.1) & 18.3 (13.3, 24.7) & 12.4 (10.2, 15.0) & \\
\hspace{1em}3 & 13.9 (10.8, 17.7) & 9.1 (5.7, 14.1) & 10.7 (8.6, 13.1) & \\
\hspace{1em}4 & 10.9 (8.5, 13.9) & 13.8 (9.7, 19.4) & 13.2 (10.9, 16.0) & \\
\hspace{1em}5 & 13.5 (10.6, 17.0) & 12.8 (8.4, 19.1) & 13.3 (11.0, 16.1) & \\
\hspace{1em}6 & 12.8 (10.1, 16.1) & 8.7 (5.7, 12.9) & 11.5 (9.5, 13.9) & \\
\hspace{1em}7 & 14.3 (11.5, 17.6) & 9.5 (6.5, 13.8) & 12.4 (10.2, 15.0) & \\
\hspace{1em}8 & 13.2 (10.5, 16.4) & 10.5 (7.4, 14.8) & 12.1 (9.9, 14.6) & \\
BMI (kg/m\textasciicircum{}2) & 27.5 (27.1, 27.9) & 27.0 (26.4, 27.6) & 27.4 (27.2, 27.6) & 0.433\\
WC (cm) & 93.3 (92.1, 94.5) & 92.9 (90.9, 95.0) & 93.1 (92.3, 93.8) & 0.928\\
Smoking status (\%) &  &  &  & \\
\hspace{1em}Current & 24.1 (21.5, 27.0) & 30.0 (24.8, 35.8) & 18.8 (17.2, 20.6) & < 0.001\\
\hspace{1em}Ex-smoker & 20.0 (17.6, 22.6) & 27.5 (22.4, 33.2) & 25.9 (24.1, 27.7) & \\
\hspace{1em}Never & 55.9 (52.7, 59.0) & 42.5 (36.6, 48.7) & 55.3 (53.2, 57.4) & \\
Current drinking status (\%) &  &  &  & \\
\hspace{1em}Yes & 24.6 (21.7, 27.7) & 18.3 (14.0, 23.6) & 16.8 (15.3, 18.4) & < 0.001\\
Hypertension (\%) \textsuperscript{\dag} &  &  &  & \\
\hspace{1em}Yes & 25.9 (22.3, 29.9) & 31.8 (25.3, 39.1) & 30.4 (27.9 32.8) & 0.111\\
Total energy intake (KJ) & 6713.8 (6575.7, 6851.8) & 9256.0 (8850.8, 9661.2) & 7916.9 (7814.0, 8019.9) & < 0.001\\
Carbohydrate intake (g) & 192.9 (188.5, 197.3) & 275.6 (263.4, 287.8) & 226.9 (223.9, 229.9) & < 0.001\\
Carbohydrate
percent (\%) \textsuperscript{\ddag} & 45.8 (45.3, 46.4) & 47.4 (46.5, 48.3) & 45.6 (45.3, 45.9) & 0.001\\
Glucose (mmol/l) & 5.16 (5.08, 5.23) & 5.09 (5.00, 5.18) & 5.14 (5.10, 5.19) & 0.537\\
A1C (\%) & 5.47 (5.42, 5.51) & 5.48 (5.42, 5.54) & 5.49 (5.47. 5.52) & 0.403\\
DM \textsuperscript{\S} & 5.9 (4.2, 8.2) & 1.1 (0.2, 5.2) & 4.7 (3.6, 6.0) & 0.053\\
Physical
activity (hours/day) \textsuperscript{\P} & 1.31 (1.14, 1.49) & 1.82 (1.44, 2.19) & 1.62 (1.49, 1.76) & 0.018\\
\bottomrule
\multicolumn{5}{l}{\textit{Note: }}\\
\multicolumn{5}{l}{Abbreviations: CI, confidence intervals; NDNS RP, national dietary and nutrition survey}\\
\multicolumn{5}{l}{rolling programme; BMI body mass index; WC, waist circumference; A1C, haemoglobin A1c;}\\
\multicolumn{5}{l}{DM, diabetes mellitus.}\\
\multicolumn{5}{l}{Variables from the blood tests (glucose and A1C) are weighted by blood sample weights,}\\
\multicolumn{5}{l}{the others are weighted by individual weights.}\\
\multicolumn{5}{l}{Glucose and A1C levels are estimated in subgroups of people without diabetes.}\\
\multicolumn{5}{l}{\textsuperscript{*} For continuous variables, the F test was used to determine differences between latent classes with Bonferroni correction to}\\
\multicolumn{5}{l}{account for multiple testing across >2 classes. For categorical variables, differences between latent classes were assessed}\\
\multicolumn{5}{l}{using the adjusted Pearson Chi-2 test for survey data.}\\
\multicolumn{5}{l}{\textsuperscript{\dag} Hypertension was defined as either systolic blood pressure >= 140 mmHg or diastolic blood pressure >= 90 mmHg, or under}\\
\multicolumn{5}{l}{treatment for hypertension.}\\
\multicolumn{5}{l}{\textsuperscript{\ddag} Carbohydrate percent indicates the percentage of energy from carbohydrate in total energy intake}\\
\multicolumn{5}{l}{\textsuperscript{\S} DM was defined by A1C > 6.5\%.}\\
\multicolumn{5}{l}{\textsuperscript{\P} Physical activity was calculated as mean time spent at moderate or vigorous physical activity during the survey.}\\
\end{tabular}
\end{table}

\rowcolors{2}{white}{white}

\section{Model fit criteria}\label{model-fit-criteria}

\rowcolors{2}{gray!6}{white}

\begin{table}

\caption{\label{tab:unnamed-chunk-4}Fit Criteria for Each Model Specification}
\centering
\fontsize{9}{11}\selectfont
\begin{tabular}[t]{cccccc}
\hiderowcolors
\toprule
\multicolumn{1}{c}{ } & \multicolumn{5}{c}{Number of level 1 classes} \\
\cmidrule(l{2pt}r{2pt}){2-6}
Model & 1 class & 2 classes & 3 classes & 4 classes & 5 classes\\
\midrule
\showrowcolors
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Fixed effects model}}\\
\hspace{1em}No. of free parameters & 14 & 29 & 44 & 59 & 74\\
\hspace{1em}\hspace{1em}Log-likelihood & -173793.306 & -172669.771 & -172039.204 & -171633.941 & -171377.292\\
\hspace{1em}\hspace{1em}BIC & 347728.092 & 345632.608 & 344523.06 & 343864.121 & 343502.409\\
\hspace{1em}\hspace{1em}Lo-Mendell-Rubun LRT & -- & < 0.0001 & 1e-04 & < 0.0001 & < 0.0001\\
\hspace{1em}\hspace{1em}Entropy & 1 & 0.31 & 0.392 & 0.51 & 0.481\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Random effects model}}\\
\hspace{1em}2 between classes &  &  &  &  & \\
\hspace{1em}\hspace{1em}No. of free parameters &  & 59 & 89 & 119 & \\
\hspace{1em}\hspace{1em}Log-likelihood &  & -169331.132 & -168700.96 & -168366.193 & \\
\hspace{1em}\hspace{1em}BIC &  & 339258.502 & 338301.338 & 337934.968 & \\
\hspace{1em}\hspace{1em}Entropy &  & 0.581 & 0.569 & 0.555 & \\
\hspace{1em}3 between classes &  &  &  &  & \\
\hspace{1em}\hspace{1em}No. of free parameters &  & 89 & 134 & 179 & \\
\hspace{1em}\hspace{1em}Log-likelihood &  & -166936.279 & -166348.815 & -166062.761 & \\
\hspace{1em}\hspace{1em}BIC &  & 334771.968 & 334051.799 & 333934.448 & \\
\hspace{1em}\hspace{1em}Entropy &  & 0.677 & 0.63 & 0.644 & \\
\hspace{1em}4 between classes &  &  &  &  & \\
\hspace{1em}\hspace{1em}No. of free parameters &  & 119 & 179 &  & \\
\hspace{1em}\hspace{1em}Log-likelihood &  & -165441.731 & -164845.696 &  & \\
\hspace{1em}\hspace{1em}BIC &  & 332086.045 & 331500.318 &  & \\
\hspace{1em}\hspace{1em}Entropy &  & 0.729 & 0.659 &  & \\
\bottomrule
\multicolumn{6}{l}{\textit{Note: }}\\
\multicolumn{6}{l}{Abbreviation: No, number; BIC, Bayesian information criterion; Entropy, a pseudo-r-squared index; Lo-Mendel-Rubin LRT, likelihood ratio test comparing q classes models with q-1 classes models.}\\
\end{tabular}
\end{table}

\rowcolors{2}{white}{white}

\section{Day level solution 1}\label{day-level-solution-1}

\rowcolors{2}{gray!6}{white}

\begin{table}

\caption{\label{tab:unnamed-chunk-5}Day Level Latent Class Solution for Three-Class Model (No Individual level Model)}
\centering
\fontsize{9}{11}\selectfont
\begin{tabular}[t]{lllllll}
\hiderowcolors
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{1}{c}{High carbohydrate day} & \multicolumn{1}{c}{Lower carbohydrate day} & \multicolumn{1}{c}{Regular day} \\
\cmidrule(l{2pt}r{2pt}){5-5} \cmidrule(l{2pt}r{2pt}){6-6} \cmidrule(l{2pt}r{2pt}){7-7}
Time slots of the day & Responses of carbohydrate intake & Frequency & Precentages (\%) & Class 1 (30.8\%) & Class 2 (39.8\%) & Class 3 (29.5\%)\\
\midrule
\showrowcolors
6 am – 9 am &  &  &  &  &  & \\
 & Not eating & 7655 & 31.2 & 0.129 & 0.45 & 0.32\\
 & Eating, Carbohydrate < 50\% & 4500 & 18.4 & 0.13 & 0.267 & 0.128\\
 & Eating, Carbohydrate >= 50\% & 12328 & 50.4 & 0.741 & 0.283 & 0.552\\
9 am – 12 am &  &  &  &  &  & \\
 & Not eating & 5447 & 22.2 & 0.237 & 0.079 & 0.401\\
 & Eating, Carbohydrate < 50\% & 7227 & 29.5 & 0.158 & 0.492 & 0.173\\
 & Eating, Carbohydrate >= 50\% & 11809 & 48.2 & 0.605 & 0.429 & 0.426\\
12 noon – 2 pm &  &  &  &  &  & \\
 & Not eating & 4783 & 19.5 & 0.156 & 0.356 & 0.019\\
 & Eating, Carbohydrate < 50\% & 11112 & 45.4 & 0.405 & 0.413 & 0.56\\
 & Eating, Carbohydrate >= 50\% & 8588 & 35.1 & 0.439 & 0.231 & 0.421\\
2 pm – 5 pm &  &  &  &  &  & \\
 & Not eating & 6926 & 28.3 & 0.13 & 0.123 & 0.659\\
 & Eating, Carbohydrate < 50\% & 8277 & 33.8 & 0.249 & 0.602 & 0.076\\
 & Eating, Carbohydrate >= 50\% & 9280 & 37.9 & 0.621 & 0.276 & 0.266\\
5 pm – 8 pm &  &  &  &  &  & \\
 & Not eating & 3043 & 12.4 & 0.114 & 0.199 & 0.034\\
 & Eating, Carbohydrate < 50\% & 14240 & 58.2 & 0.516 & 0.59 & 0.639\\
 & Eating, Carbohydrate >= 50\% & 7200 & 29.4 & 0.37 & 0.211 & 0.328\\
8 pm – 10 pm &  &  &  &  &  & \\
 & Not eating & 8722 & 35.6 & 0.322 & 0.291 & 0.48\\
 & Eating, Carbohydrate < 50\% & 8898 & 36.3 & 0.266 & 0.551 & 0.212\\
 & Eating, Carbohydrate >= 50\% & 6863 & 28.0 & 0.412 & 0.158 & 0.308\\
10 pm – 6 am &  &  &  &  &  & \\
 & Not eating & 16295 & 66.6 & 0.68 & 0.59 & 0.751\\
 & Eating, Carbohydrate < 50\% & 4144 & 16.9 & 0.074 & 0.294 & 0.101\\
 & Eating, Carbohydrate >= 50\% & 4044 & 16.5 & 0.246 & 0.115 & 0.148\\
\bottomrule
\end{tabular}
\end{table}

\rowcolors{2}{white}{white}

\section{Day level solution
characteristics}\label{day-level-solution-characteristics}

\rowcolors{2}{gray!6}{white}

\begin{table}

\caption{\label{tab:unnamed-chunk-6}Means (sd), and counts  of the characteristics of different types of days according to carbohydrate intake.}
\centering
\fontsize{9}{11}\selectfont
\begin{tabular}[t]{lllll}
\hiderowcolors
\toprule
  & High carbohydrate day & Low carbohydrate day & Regular meals day & P value\textsuperscript{*}\\
\midrule
\showrowcolors
Counts & 9667 (39.5) & 5002 (20.4\%) & 9814 (40.1\%) & \\
Country (\%) &  &  &  & < 0.001\\
\hspace{1em}England & 5627 (58.2) & 2972 (59.4) & 5291 (53.9) & \\
\hspace{1em}NI & 1194 (12.4) & 527 (10.5) & 1400 (14.3) & \\
\hspace{1em}Scotland & 1527 (15.8) & 813 (16.3) & 1774 (18.1) & \\
\hspace{1em}Wales & 1318 (13.6) & 690 (13.8) & 1349 (13.7) & \\
Day of Week (\%) &  &  &  & < 0.001\\
\hspace{1em}Monday & 1303 (13.5) & 715 (14.3) & 1370 (14.0) & \\
\hspace{1em}Tuesday & 1266 (13.1) & 674 (13.5) & 1290 (13.1) & \\
\hspace{1em}Wednesday & 1225 (12.7) & 740 (14.8) & 1233 (12.6) & \\
\hspace{1em}Thursday & 1272 (13.2) & 752 (15.0) & 1425 (14.5) & \\
\hspace{1em}Friday & 1458 (15.1) & 797 (15.9) & 1479 (15.1) & \\
\hspace{1em}Saturday & 1537 (15.9) & 703 (14.1) & 1495 (15.2) & \\
\hspace{1em}Sunday & 1605 (16.6) & 621 (12.4) & 1522 (15.5) & \\
\hspace{1em}Weekend, Yes (\%) & 3142 (32.5) & 1324 (26.5) & 3017 (30.7) & < 0.001\\
Total energy (kJ) & 7539.98 (2875.87) & 7160.22 (2922.15) & 7439.68 (2978.91) & < 0.001\\
Carbohydrate (g) & 222.79 (89.84) & 209.70 (86.17) & 206.59 (84.42) & < 0.001\\
Protein (g) & 71.36 (29.79) & 69.55 (30.20) & 73.29 (32.94) & < 0.001\\
Fat (g) & 65.44 (33.27) & 63.94 (33.76) & 67.24 (34.73) & < 0.001\\
Alcohol (g) & 11.76 (27.31) & 8.85 (24.25) & 13.80 (33.00) & < 0.001\\
Total sugars (g) & 98.63 (56.03) & 88.03 (50.50) & 86.39 (50.96) & < 0.001\\
Starch (g) & 124.07 (55.84) & 121.59 (56.13) & 120.11 (54.62) & < 0.001\\
Non-milk extrinsic sugar\textsuperscript{\dag} & 59.45 (49.31) & 50.07 (43.41) & 50.41 (44.84) & < 0.001\\
Fruit (g) & 107.40 (137.97) & 103.15 (129.08) & 92.76 (126.02) & < 0.001\\
Yellow Red Green Vegetables (g) & 26.52 (46.44) & 26.84 (47.99) & 26.16 (45.99) & 0.681\\
\bottomrule
\multicolumn{5}{l}{\textsuperscript{*} P values were obtained from chi-square test for categorical variables, and one-way ANOVA comparing the}\\
\multicolumn{5}{l}{means in multiple groups for continuous variables;}\\
\multicolumn{5}{l}{\textsuperscript{\dag} Non-milk extrinsic sugar is defined as:}\\
\end{tabular}
\end{table}

\rowcolors{2}{white}{white}

\section{Social economic}\label{social-economic}

\rowcolors{2}{gray!6}{white}

\begin{table}

\caption{\label{tab:tab1}Means, row percentages, and 95 CIs of the characteristics by latent class memberships in the UK adults (NDNS RP 2008/09-2015/16, sample size = 6155).}
\centering
\fontsize{9}{11}\selectfont
\begin{tabular}[t]{lllll}
\hiderowcolors
\toprule
Variables & Individual class 1 (n = 1730) & Individual class = 2 (n = 1772) & Individual class = 3 (n = 2653) & P value \textsuperscript{*}\\
\midrule
\showrowcolors
Total (\%) & 28.4 (26.9, 29.9) & 28.7 (27.1, 30.3) & 43.0 (41.3, 44.7) & \\
Countries (\%) &  &  &  & 0.007\\
\hspace{1em}England & 84.5 (81.7, 86.9) & 82.0 (79.3, 84.5) & 84.7 (82.3, 86.8) & \\
\hspace{1em}Northern Ireland & 2.1 (1.6, 2.8) & 4.2 (3.2, 5.6) & 2.2 (1.7, 3.0) & \\
\hspace{1em}Scotland & 9.1 (7.0, 11.8) & 8.6 (6.7, 11.1) & 8.0 (6.3, 10.2) & \\
\hspace{1em}Wales & 4.3 (3.3, 5.6) & 5.1 (4.0, 6.4) & 5.1 (4.0, 6.4) & \\
Age (years) & 51.0 (49.9, 52.1) & 40.3 (39.1, 41.6) & 51.7 (50.7, 52.7) & < 0.001\\
Sex (\%) &  &  &  & 0.119\\
\hspace{1em}Men & 50.0 (46.9, 53.1) & 50.2 (47.0, 53.5) & 46.6 (44.0, 49.1) & \\
\hspace{1em}Women & 50.0 (46.9, 53.1) & 49.8 (46.5, 53.0) & 53.4 (50.9, 56.0) & \\
Survey years (\% in rows) &  &  &  & 0.015\\
\hspace{1em}1 & 32.5 (28.4, 36.9) & 26.3 (21.9, 31.2) & 41.2 (36.6, 46.0) & \\
\hspace{1em}2 & 26.8 (22.6, 31.3) & 22.6 (18.6, 27.3) & 50.6 (45.8, 55.4) & \\
\hspace{1em}3 & 22.6 (18.8, 26.9) & 33.7 (28.6, 39.2) & 43.6 (38.7, 48.7) & \\
\hspace{1em}4 & 27.9 (24.1, 32.2) & 27.6 (23.8, 31.8) & 44.4 (40.2, 48.7) & \\
\hspace{1em}5 & 27.9 (24.2, 32.0) & 28.7 (24.4, 33.5) & 43.3 (38.2, 48.6) & \\
\hspace{1em}6 & 28.0 (24.0, 32.4) & 31.5 (26.9, 36.6) & 40.5 (35.8, 45.3) & \\
\hspace{1em}7 & 29.1 (25.2, 33.4) & 29.0 (24.5, 34.0) & 41.8 (37.1, 46.7) & \\
\hspace{1em}8 & 31.1 (27.3, 35.3) & 30.5 (25.9, 35.5) & 38.4 (34.1, 42.8) & \\
Paid employment\textsuperscript{\dag} (\%) &  &  &  & 0.907\\
\hspace{1em}Yes & 40.3 (37.0, 43.6) & 40.8 (37.1, 44.5) & 39.8 (37.1, 42.6) & \\
\hspace{1em}No & 59.7 (56.4, 63.0) & 59.2 (55.5, 62.9) & 60.2 (57.4, 62.9) & \\
Live with partner\textsuperscript{\ddag} (\%) &  &  &  & < 0.001\\
\hspace{1em}Yes & 56.9 (53.6, 60.1) & 38.4 (35.2, 41.8) & 61.3 (58.7, 63.7) & \\
\hspace{1em}No & 43.1 (39.9, 46.4) & 61.6 (58.2, 64.8)) & 38.7 (36.3, 41,3) & \\
Household Income, GBP/year & 36558.53 (34800.21, 38316.84) & 27180.80 (25597.95, 28763.65) & 32171.58 (31024.96, 33318.2) & < 0.001\\
Ethnicity (\%) &  &  &  & \\
\hspace{1em}White & 94.2 (92.4, 95.6) & 79.5 (76.4, 82.3) & 91.9 (90.1, 93.4) & < 0.001\\
\hspace{1em}Non-White & 5.8 (4.4, 7.6) & 20.5 (17.7, 23.6) & 8.1 (6.6, 9.9) & \\
Education (\%) &  &  &  & \\
\hspace{1em}Degree or higher & 29.0 (26.1, 32.1) & 23.3 (20.5, 26.3) & 26.2 (24.1, 28.5) & 0.019\\
\hspace{1em}Lower than degree & 71.0 (67.9, 73.9) & 76.7 (73.7, 79.5) & 73.8 (71.5, 75.9) & \\
\bottomrule
\multicolumn{5}{l}{\textit{Note: }}\\
\multicolumn{5}{l}{Abbreviations: CI, confidence intervals; NDNS RP, national dietary and nutrition survey rolling programme;}\\
\multicolumn{5}{l}{\textsuperscript{*} For continuous variables, the F test was used to determine differences between latent classes with Bonferroni correction to account for multiple}\\
\multicolumn{5}{l}{testing across >2 classes. For categorical variables, differences between latent classes were assessed using the adjusted Pearson Chi-2 test for}\\
\multicolumn{5}{l}{survey data.}\\
\multicolumn{5}{l}{\textsuperscript{\dag} Paid employment was defined as being in paid employment during the last 4 weeks prior to the survey.}\\
\multicolumn{5}{l}{\textsuperscript{\ddag} Live with partner was defined as either living with a married husband/wife or a legally recognised civil partnership.}\\
\end{tabular}
\end{table}

\rowcolors{2}{white}{white}

\subsection{Anthropometric measurements and biochemical
analysis}\label{anthropometric-measurements-and-biochemical-analysis}

\rowcolors{2}{gray!6}{white}

\begin{table}

\caption{\label{tab:tab2}Means, percentages, and 95 CIs of the characteristics by latent class memberships in the UK adults (NDNS RP 2008/09-2015/16, sample size = 6155).}
\centering
\fontsize{9}{11}\selectfont
\begin{tabular}[t]{lllll}
\hiderowcolors
\toprule
Variables & Individual class 1 (n = 1730) & Individual class (n = 1772) & Individual class 3 (n = 2653) & P value\textsuperscript{*}\\
\midrule
\showrowcolors
BMI (kg/m2) & 27.8 (27.4, 28.2) & 27.2 (26.7, 27.7) & 27.3 (26.9, 27.6) & 0.006\\
WC (cm) & 94.6 (93.5, 95.6) & 92.3 (91.0, 93.5) & 92.2 (91.4, 93.1) & 0.001\\
Smoking status (\%) &  &  &  & \\
\hspace{1em}Current & 20.4 (18.0, 23.0) & 27.8 (25.0, 30.9) & 17.1 (15.4, 19.0) & < 0.001\\
\hspace{1em}Ex-smoker & 29.3 (26.5, 32.2) & 16.8 (14.6, 19.2) & 26.1 (24.9, 28.3) & \\
\hspace{1em}Never & 50.3 (47.2, 32.2) & 55.4 (52.2, 58.6) & 56.8 (54.3, 59.3) & \\
Current drinking status (\%) &  &  &  & \\
\hspace{1em}Yes & 11.8 (10.0, 13.8) & 24.0 (21.2, 27.1) & 20.6 (18.7, 22.8) & < 0.001\\
Hypertension\textsuperscript{\dag} (\%) &  &  &  & \\
\hspace{1em}Yes & 33.4 (29.9, 37.1) & 21.1 (17.7, 24.9) & 31.2 (28.3, 34.2) & < 0.001\\
Total energy intake (kJ) & 7985.8 (7823.3, 8146.295) & 7341.8 (7825.3, 8146.3) & 7677 (7555.8, 7799,8) & < 0.001\\
Carbohydrate percent\textsuperscript{\ddag} (\%) & 203.8 (199.8, 207.8) & 218.3 (212.9, 223.7) & 233.4 (229.6, 237.2) & < 0.001\\
Carbohydrate percent (\%) & 40.6 (40.2, 41.0) & 47.3 (46.8, 47.8) & 48.3 (47.9, 48.6) & < 0.001\\
Protein intake (g) & 79.9 (77.9, 81.8) & 69.3 (67.6, 71.0) & 73.7 (72.5, 74.8) & < 0.001\\
Protein percent (\%) & 17.2 (16.9, 17.5) & 16.3 (16.0, 16.6) & 16.5 (16.3, 16.6) & < 0.001\\
Fat intake (g) & 74.7 (73.1, 76.4) & 63.8 (62.1, 65.5) & 65.7 (64.4, 67.0) & < 0.001\\
Fat percent (\%) & 35.4 (34.9, 35.8) & 32.5 (32.1, 32.9) & 32.0 (31.7, 32.3) & < 0.001\\
Glucose (mmol/l) & 5.20 (5.15, 5.26) & 5.09 (5.02, 5.17) & 5.13 (5.09, 5.12) & 0.051\\
A1C (\%) & 5.49 (5.45, 5.52) & 5.44 (5.40, 5.48) & 5.52 (5.49, 5.55) & 0.012\\
DM \textsuperscript{\S} & 6.9 (5.0, 9.3) & 3.5 (2.3, 5.3) & 4.1 (2.9, 5.6) & 0.011\\
Physical
activity (hours/day) \textsuperscript{\P} & 1.63 (1.43, 1.83) & 1.32 (1.15, 1.49) & 1.65 (1.49, 1.81) & 0.016\\
\bottomrule
\multicolumn{5}{l}{\textit{Note: }}\\
\multicolumn{5}{l}{Abbreviations: CI, confidence intervals; NDNS RP, national dietary and nutrition survey rolling programme;}\\
\multicolumn{5}{l}{ BMI body mass index; WC, waist circumference; A1C, haemoglobin A1c; DM, diabetes mellitus.}\\
\multicolumn{5}{l}{Variables from the blood tests (glucose and A1C) are weighted by blood sample weights, the others are weighted by individual weights.}\\
\multicolumn{5}{l}{Glucose and A1C levels are estimated in subgroups of people without diabetes.}\\
\multicolumn{5}{l}{\textsuperscript{*} For continuous variables, the F test was used to determine differences between latent classes with Bonferroni correction to account}\\
\multicolumn{5}{l}{for multiple testing across >2 classes. For categorical variables, differences between latent classes were assessed using the adjusted}\\
\multicolumn{5}{l}{Pearson Chi-2 test for survey data.}\\
\multicolumn{5}{l}{\textsuperscript{\dag} Hypertension was defined as either systolic blood pressure >= 140 mmHg or diastolic blood pressure >= 90 mmHg, or under treatment for}\\
\multicolumn{5}{l}{hypertension.}\\
\multicolumn{5}{l}{\textsuperscript{\ddag} Carbohydrate percent indicates the percentage of energy from carbohydrate in total energy intake}\\
\multicolumn{5}{l}{\textsuperscript{\S} DM was defined by A1C > 6.5\%.}\\
\multicolumn{5}{l}{\textsuperscript{\P} Physical activity was calculated as mean time spent at moderate or vigorous physical activity including both work-related and}\\
\multicolumn{5}{l}{recreational activities during the survey.}\\
\end{tabular}
\end{table}

\rowcolors{2}{white}{white}

\end{document}