### 2018-08-17 paper/Appendices/AppendixD.tex

% Appendix Template

\chapter{R \& Stata codes for processing individual level data analysis} % Main appendix title

\label{AppendixD}
\linespread{0.8}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(plyr)}
\KeywordTok{library}\NormalTok{(epiDisplay)}
\KeywordTok{library}\NormalTok{(tidyverse)}
\KeywordTok{library}\NormalTok{(dplyr)}
\KeywordTok{library}\NormalTok{(readr)}
\KeywordTok{library}\NormalTok{(haven)}
\KeywordTok{library}\NormalTok{(naniar)}

\CommentTok{# read the individual level data sets ----------------------------------------}

\NormalTok{blood78 <-}\StringTok{ }\KeywordTok{read_dta}\NormalTok{(}\StringTok{"ndns_rp_yr7-8a_indiv.dta"}\NormalTok{)}
\NormalTok{blood56 <-}\StringTok{ }\KeywordTok{read_dta}\NormalTok{(}\StringTok{"ndns_rp_yr5-6a_indiv.dta"}\NormalTok{)}
\NormalTok{blood14 <-}\StringTok{ }\KeywordTok{read_dta}\NormalTok{(}\StringTok{"ndns_rp_yr1-4a_indiv_uk.dta"}\NormalTok{)}
\NormalTok{food14 <-}\StringTok{ }\KeywordTok{read_dta}\NormalTok{(}\StringTok{"ndns_rp_yr1-4a_personleveldietarydata_uk.dta"}\NormalTok{)}
\NormalTok{food56 <-}\StringTok{ }\KeywordTok{read_dta}\NormalTok{(}\StringTok{"ndns_rp_yr5-6a_personleveldietarydata.dta"}\NormalTok{)}
\NormalTok{food78 <-}\StringTok{ }\KeywordTok{read_dta}\NormalTok{(}\StringTok{"ndns_rp_yr7-8a_personleveldietarydata.dta"}\NormalTok{)}
\KeywordTok{names}\NormalTok{(blood78)[}\KeywordTok{names}\NormalTok{(blood78)}\OperatorTok{==}\StringTok{"seriali"}\NormalTok{] <-}\StringTok{ "ID"}
\KeywordTok{names}\NormalTok{(blood56)[}\KeywordTok{names}\NormalTok{(blood56)}\OperatorTok{==}\StringTok{"seriali"}\NormalTok{] <-}\StringTok{ "ID"}
\KeywordTok{names}\NormalTok{(blood14)[}\KeywordTok{names}\NormalTok{(blood14)}\OperatorTok{==}\StringTok{"seriali"}\NormalTok{] <-}\StringTok{ "ID"}
\KeywordTok{names}\NormalTok{(food78)[}\KeywordTok{names}\NormalTok{(food78)}\OperatorTok{==}\StringTok{"seriali"}\NormalTok{] <-}\StringTok{ "ID"}
\KeywordTok{names}\NormalTok{(food56)[}\KeywordTok{names}\NormalTok{(food56)}\OperatorTok{==}\StringTok{"seriali"}\NormalTok{] <-}\StringTok{ "ID"}
\KeywordTok{names}\NormalTok{(food14)[}\KeywordTok{names}\NormalTok{(food14)}\OperatorTok{==}\StringTok{"seriali"}\NormalTok{] <-}\StringTok{ "ID"}

\CommentTok{# Loading the data from Mplus output}
\NormalTok{CW3CB3 <-}\StringTok{ }\KeywordTok{read_table2}\NormalTok{(}\StringTok{"../NDNSslot_CW3CB3.txt"}\NormalTok{,  }\CommentTok{# change the path accordingly}
                      \DataTypeTok{col_names =} \OtherTok{FALSE}\NormalTok{)}

\KeywordTok{names}\NormalTok{(CW3CB3) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Breakfast"}\NormalTok{, }\StringTok{"Morning.snack"}\NormalTok{, }\StringTok{"Lunch"}\NormalTok{,}
                   \StringTok{"Afternoon.snack"}\NormalTok{, }\StringTok{"Dinner"}\NormalTok{, }\StringTok{"Before.bedtime.snack"}\NormalTok{,}
                   \StringTok{"Midnight.food"}\NormalTok{, }\StringTok{"ID_DAY"}\NormalTok{, }\StringTok{"AGE"}\NormalTok{, }\StringTok{"SEX"}\NormalTok{, }\StringTok{"CPROB1"}\NormalTok{,}
                   \StringTok{"CPROB2"}\NormalTok{, }\StringTok{"CPROB3"}\NormalTok{, }\StringTok{"CPROB4"}\NormalTok{, }\StringTok{"CPROB5"}\NormalTok{, }\StringTok{"CPROB6"}\NormalTok{, }
                   \StringTok{"CPROB7"}\NormalTok{, }\StringTok{"CPROB8"}\NormalTok{, }\StringTok{"CPROB9"}\NormalTok{, }
                   \StringTok{"CB"}\NormalTok{, }\StringTok{"CW"}\NormalTok{, }\StringTok{"MLCJOINT"}\NormalTok{, }\StringTok{"ID"}\NormalTok{)}

\NormalTok{CW3idday <-}\StringTok{ }\NormalTok{CW3CB3 }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(ID, ID_DAY, CW, CB, MLCJOINT, AGE, SEX)}

\NormalTok{Energy_slots <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"../Energy_slots.csv"}\NormalTok{) }\CommentTok{# change the path accordingly}
\NormalTok{Energy_slots <-}\StringTok{ }\NormalTok{Energy_slots }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{ID =}\NormalTok{ id)}

\CommentTok{# Recode day level classification to keep consistency}
\NormalTok{CW3idday}\OperatorTok{$}\NormalTok{CW_new <-}\StringTok{ }\DecValTok{0}
\NormalTok{CW3idday}\OperatorTok{$}\NormalTok{CW_new[CW3idday}\OperatorTok{$}\NormalTok{CW }\OperatorTok{==}\StringTok{ }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{3}
\NormalTok{CW3idday}\OperatorTok{$}\NormalTok{CW_new[CW3idday}\OperatorTok{$}\NormalTok{CW }\OperatorTok{==}\StringTok{ }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}
\NormalTok{CW3idday}\OperatorTok{$}\NormalTok{CW_new[CW3idday}\OperatorTok{$}\NormalTok{CW }\OperatorTok{==}\StringTok{ }\DecValTok{3}\NormalTok{] <-}\StringTok{ }\DecValTok{2}

\NormalTok{Energy_slots <-}\StringTok{ }\NormalTok{Energy_slots }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{left_join}\NormalTok{(CW3idday, }\DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"DayofWeek"}\NormalTok{))}

\NormalTok{Energy_slots}\OperatorTok{$}\NormalTok{TimeSlot <-}\StringTok{ }\KeywordTok{factor}\NormalTok{(Energy_slots}\OperatorTok{$}\NormalTok{TimeSlot, }
                                \DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{"[6,9)"}\NormalTok{, }\StringTok{"[9,12)"}\NormalTok{, }
                                           \StringTok{"[12,14)"}\NormalTok{, }\StringTok{"[14,17)"}\NormalTok{,}
                                           \StringTok{"[17,20)"}\NormalTok{, }\StringTok{"[20,22)"}\NormalTok{, }\StringTok{"[22, 6)"}\NormalTok{))}

\NormalTok{FoodbyCB_7slots <-}\StringTok{ }\NormalTok{Energy_slots }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(ID, TimeSlot) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{summarise}\NormalTok{(}\DataTypeTok{sumEnergy =} \KeywordTok{sum}\NormalTok{(Tot_Energ),}
            \DataTypeTok{sumCarb =} \KeywordTok{sum}\NormalTok{(Tot_Carb), }
            \DataTypeTok{sumSugar =} \KeywordTok{sum}\NormalTok{(Tot_Sugar), }
            \DataTypeTok{sumStarch =} \KeywordTok{sum}\NormalTok{(Tot_Starch),}
            \DataTypeTok{sumFibre =} \KeywordTok{sum}\NormalTok{(Tot_Fibre), }
            \DataTypeTok{sumNMES =} \KeywordTok{sum}\NormalTok{(Tot_NMES),}
            \DataTypeTok{sumFat =} \KeywordTok{sum}\NormalTok{(Tot_Fat),}
            \DataTypeTok{sumProt =} \KeywordTok{sum}\NormalTok{(Tot_Prot), }
            \DataTypeTok{sumAlc =} \KeywordTok{sum}\NormalTok{(Tot_Alc))}

\CommentTok{# Calculate the sum of each nutrients for each time slots}
\NormalTok{Carbsum <-}\StringTok{ }\NormalTok{FoodbyCB_7slots }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(ID, TimeSlot, sumCarb) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{spread}\NormalTok{(}\DataTypeTok{key =}\NormalTok{ TimeSlot, }
         \DataTypeTok{value =}\NormalTok{ sumCarb)}

\NormalTok{Carbsum[}\KeywordTok{is.na}\NormalTok{(Carbsum)] <-}\StringTok{ }\DecValTok{0}

\KeywordTok{names}\NormalTok{(Carbsum) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"Carb6_9"}\NormalTok{, }\StringTok{"Carb9_12"}\NormalTok{, }\StringTok{"Carb12_14"}\NormalTok{, }\StringTok{"Carb14_17"}\NormalTok{,}
                     \StringTok{"Carb17_20"}\NormalTok{, }\StringTok{"Carb20_22"}\NormalTok{, }\StringTok{"Carb22_6"}\NormalTok{)}

\NormalTok{Energysum <-}\StringTok{ }\NormalTok{FoodbyCB_7slots }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(ID, TimeSlot, sumEnergy) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{spread}\NormalTok{(}\DataTypeTok{key =}\NormalTok{ TimeSlot, }
         \DataTypeTok{value =}\NormalTok{ sumEnergy)}
\NormalTok{Energysum[}\KeywordTok{is.na}\NormalTok{(Energysum)] <-}\StringTok{ }\DecValTok{0}

\KeywordTok{names}\NormalTok{(Energysum) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"Energy6_9"}\NormalTok{, }\StringTok{"Energy9_12"}\NormalTok{, }\StringTok{"Energy12_14"}\NormalTok{, }
\StringTok{"Energy14_17"}\NormalTok{, }\StringTok{"Energy17_20"}\NormalTok{, }\StringTok{"Energy20_22"}\NormalTok{, }\StringTok{"Energy22_6"}\NormalTok{)}

\NormalTok{Starchsum <-}\StringTok{ }\NormalTok{FoodbyCB_7slots }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(ID, TimeSlot, sumStarch) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{spread}\NormalTok{(}\DataTypeTok{key =}\NormalTok{ TimeSlot, }
         \DataTypeTok{value =}\NormalTok{ sumStarch)}
\NormalTok{Starchsum[}\KeywordTok{is.na}\NormalTok{(Starchsum)] <-}\StringTok{ }\DecValTok{0}

\KeywordTok{names}\NormalTok{(Starchsum) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"Starch6_9"}\NormalTok{, }\StringTok{"Starch9_12"}\NormalTok{, }\StringTok{"Starch12_14"}\NormalTok{, }
\StringTok{"Starch14_17"}\NormalTok{, }\StringTok{"Starch17_20"}\NormalTok{, }\StringTok{"Starch20_22"}\NormalTok{, }\StringTok{"Starch22_6"}\NormalTok{)}

\NormalTok{Sugarsum <-}\StringTok{ }\NormalTok{FoodbyCB_7slots }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(ID, TimeSlot, sumSugar) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{spread}\NormalTok{(}\DataTypeTok{key =}\NormalTok{ TimeSlot, }
         \DataTypeTok{value =}\NormalTok{ sumSugar)}
\NormalTok{Sugarsum[}\KeywordTok{is.na}\NormalTok{(Sugarsum)] <-}\StringTok{ }\DecValTok{0}

\KeywordTok{names}\NormalTok{(Sugarsum) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"Sugar6_9"}\NormalTok{, }\StringTok{"Sugar9_12"}\NormalTok{, }\StringTok{"Sugar12_14"}\NormalTok{, }
\StringTok{"Sugar14_17"}\NormalTok{, }\StringTok{"Sugar17_20"}\NormalTok{, }\StringTok{"Sugar20_22"}\NormalTok{, }\StringTok{"Sugar22_6"}\NormalTok{)}

\NormalTok{Fibresum <-}\StringTok{ }\NormalTok{FoodbyCB_7slots }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(ID, TimeSlot, sumFibre) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{spread}\NormalTok{(}\DataTypeTok{key =}\NormalTok{ TimeSlot, }
         \DataTypeTok{value =}\NormalTok{ sumFibre)}
\NormalTok{Fibresum[}\KeywordTok{is.na}\NormalTok{(Fibresum)] <-}\StringTok{ }\DecValTok{0}

\KeywordTok{names}\NormalTok{(Fibresum) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"Fibre6_9"}\NormalTok{, }\StringTok{"Fibre9_12"}\NormalTok{, }\StringTok{"Fibre12_14"}\NormalTok{, }
\StringTok{"Fibre14_17"}\NormalTok{, }\StringTok{"Fibre17_20"}\NormalTok{, }\StringTok{"Fibre20_22"}\NormalTok{, }\StringTok{"Fibre22_6"}\NormalTok{)}

\NormalTok{NMESsum <-}\StringTok{ }\NormalTok{FoodbyCB_7slots }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(ID, TimeSlot, sumNMES) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{spread}\NormalTok{(}\DataTypeTok{key =}\NormalTok{ TimeSlot, }
         \DataTypeTok{value =}\NormalTok{ sumNMES)}
\NormalTok{NMESsum[}\KeywordTok{is.na}\NormalTok{(NMESsum)] <-}\StringTok{ }\DecValTok{0}

\KeywordTok{names}\NormalTok{(NMESsum) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"NMES6_9"}\NormalTok{, }\StringTok{"NMES9_12"}\NormalTok{, }\StringTok{"NMES12_14"}\NormalTok{, }
\StringTok{"NMES14_17"}\NormalTok{, }\StringTok{"NMES17_20"}\NormalTok{, }\StringTok{"NMES20_22"}\NormalTok{, }\StringTok{"NMES22_6"}\NormalTok{)}

\NormalTok{Fatsum <-}\StringTok{ }\NormalTok{FoodbyCB_7slots }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(ID, TimeSlot, sumFat) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{spread}\NormalTok{(}\DataTypeTok{key =}\NormalTok{ TimeSlot, }
         \DataTypeTok{value =}\NormalTok{ sumFat)}
\NormalTok{Fatsum[}\KeywordTok{is.na}\NormalTok{(Fatsum)] <-}\StringTok{ }\DecValTok{0}

\KeywordTok{names}\NormalTok{(Fatsum) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"Fat6_9"}\NormalTok{, }\StringTok{"Fat9_12"}\NormalTok{, }\StringTok{"Fat12_14"}\NormalTok{, }\StringTok{"Fat14_17"}\NormalTok{,}
                     \StringTok{"Fat17_20"}\NormalTok{, }\StringTok{"Fat20_22"}\NormalTok{, }\StringTok{"Fat22_6"}\NormalTok{)}

\NormalTok{Protsum <-}\StringTok{ }\NormalTok{FoodbyCB_7slots }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(ID, TimeSlot, sumProt) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{spread}\NormalTok{(}\DataTypeTok{key =}\NormalTok{ TimeSlot, }
         \DataTypeTok{value =}\NormalTok{ sumProt)}
\NormalTok{Protsum[}\KeywordTok{is.na}\NormalTok{(Protsum)] <-}\StringTok{ }\DecValTok{0}

\KeywordTok{names}\NormalTok{(Protsum) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"Prot6_9"}\NormalTok{, }\StringTok{"Prot9_12"}\NormalTok{, }\StringTok{"Prot12_14"}\NormalTok{, }\StringTok{"Prot14_17"}\NormalTok{,}
                    \StringTok{"Prot17_20"}\NormalTok{, }\StringTok{"Prot20_22"}\NormalTok{, }\StringTok{"Prot22_6"}\NormalTok{)}

\NormalTok{Alcsum <-}\StringTok{ }\NormalTok{FoodbyCB_7slots }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(ID, TimeSlot, sumAlc) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{spread}\NormalTok{(}\DataTypeTok{key =}\NormalTok{ TimeSlot, }
         \DataTypeTok{value =}\NormalTok{ sumAlc)}
\NormalTok{Alcsum[}\KeywordTok{is.na}\NormalTok{(Alcsum)] <-}\StringTok{ }\DecValTok{0}

\KeywordTok{names}\NormalTok{(Alcsum) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"Alc6_9"}\NormalTok{, }\StringTok{"Alc9_12"}\NormalTok{, }\StringTok{"Alc12_14"}\NormalTok{, }\StringTok{"Alc14_17"}\NormalTok{,}
                     \StringTok{"Alc17_20"}\NormalTok{, }\StringTok{"Alc20_22"}\NormalTok{, }\StringTok{"Alc22_6"}\NormalTok{)}

\CommentTok{# Extract number of days of diary completed}

\NormalTok{blood14 <-}\StringTok{ }\NormalTok{blood14 }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(ID, Ndays)}

\NormalTok{blood56 <-}\StringTok{ }\NormalTok{blood56 }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(ID, Ndays)}

\NormalTok{blood78 <-}\StringTok{ }\NormalTok{blood78 }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(ID, NDays) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{Ndays =}\NormalTok{ NDays)}

\NormalTok{NDAYS <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(blood14, blood56, blood78)}
\NormalTok{NDAYS}\OperatorTok{$}\NormalTok{ID <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{(NDAYS}\OperatorTok{$}\NormalTok{ID)}

\NormalTok{IntakeSlots <-}\StringTok{ }\NormalTok{Energysum }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{left_join}\NormalTok{(Carbsum, }\DataTypeTok{by =} \StringTok{"ID"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{left_join}\NormalTok{(Sugarsum, }\DataTypeTok{by =} \StringTok{"ID"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{left_join}\NormalTok{(Starchsum, }\DataTypeTok{by =} \StringTok{"ID"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{left_join}\NormalTok{(Fibresum, }\DataTypeTok{by =} \StringTok{"ID"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{left_join}\NormalTok{(Fatsum, }\DataTypeTok{by =} \StringTok{"ID"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{left_join}\NormalTok{(Protsum, }\DataTypeTok{by =} \StringTok{"ID"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{left_join}\NormalTok{(NMESsum, }\DataTypeTok{by =} \StringTok{"ID"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{left_join}\NormalTok{(Alcsum, }\DataTypeTok{by =} \StringTok{"ID"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{left_join}\NormalTok{(NDAYS, }\DataTypeTok{by =} \StringTok{"ID"}\NormalTok{)}

\NormalTok{IntakeSlots}\OperatorTok{$}\NormalTok{ID <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{(IntakeSlots}\OperatorTok{$}\NormalTok{ID)}

\NormalTok{IntakeSlots}\OperatorTok{$}\NormalTok{Energy6_}\DecValTok{9}\NormalTok{ <-}\StringTok{ }\NormalTok{IntakeSlots}\OperatorTok{$}\NormalTok{Energy6_}\DecValTok{9}\OperatorTok{/}\NormalTok{(IntakeSlots}\OperatorTok{$}\NormalTok{Ndays)}

\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{3}\OperatorTok{:}\DecValTok{57}\NormalTok{)\{}
\NormalTok{  IntakeSlots[, i] <-}\StringTok{ }\NormalTok{IntakeSlots[, i]}\OperatorTok{/}\NormalTok{(IntakeSlots}\OperatorTok{$}\NormalTok{Ndays)}
\NormalTok{\}}

\CommentTok{# select the variables needed, recoding }
\CommentTok{# NAs, renaming to the same ----------------------------------------}

\NormalTok{BMI78 <-}\StringTok{ }\NormalTok{blood78 }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(ID, Sex, age, bmival, wstval, Diabetes, bpmedc2, bpmedd2, }
\NormalTok{         hyper140_}\DecValTok{2}\NormalTok{, hibp140_}\DecValTok{2}\NormalTok{,  Glucose, A1C, cigsta3, dnoft3, }
\NormalTok{         dnnow, wti_Y78, wtn_Y78, wtb_Y78, ethgrp5, ethgrp2,}
\NormalTok{         cluster1, cluster2, cluster3, nssec8, paidemployment, }
\NormalTok{         qual7, eqvinc, MarSt2, cluster4, cluster5, area, gor, }
\NormalTok{         LDL, HDL, Chol, Trig) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{wti =}\NormalTok{ wti_Y78, }\DataTypeTok{wtn =}\NormalTok{ wtn_Y78, }\DataTypeTok{wtb =}\NormalTok{ wtb_Y78, }
         \DataTypeTok{drink =}\NormalTok{ dnoft3) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{Years =} \StringTok{"7-8"}\NormalTok{, }\DataTypeTok{MVPAtime =} \OtherTok{NA}\NormalTok{, }\DataTypeTok{MarStat =} \OtherTok{NA}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{replace_with_na}\NormalTok{(}\DataTypeTok{replace =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{bmival =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{qual7 =} \OperatorTok{-}\DecValTok{8}\NormalTok{,}
                                 \DataTypeTok{wstval =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{eqvinc =} \OperatorTok{-}\DecValTok{1}\NormalTok{,}
                                 \DataTypeTok{bpmedd2 =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{MarSt2 =} \OperatorTok{-}\DecValTok{1}\NormalTok{,}
                                 \DataTypeTok{bpmedc2 =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{hyper140_2 =} \OperatorTok{-}\DecValTok{7}\NormalTok{, }
                                 \DataTypeTok{hibp140_2 =} \OperatorTok{-}\DecValTok{7}\NormalTok{, }\DataTypeTok{Glucose =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }
                                 \DataTypeTok{A1C  =} \OperatorTok{-}\DecValTok{1}\NormalTok{,  }\DataTypeTok{LDL =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{HDL =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }
                                 \DataTypeTok{Chol =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{Trig =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{dnnow =} \OperatorTok{-}\DecValTok{1}\NormalTok{,}
                                 \DataTypeTok{drink =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{ethgrp5 =} \OperatorTok{-}\DecValTok{4}\NormalTok{, }
                                 \DataTypeTok{ethgrp2 =} \OperatorTok{-}\DecValTok{4}\NormalTok{, }\DataTypeTok{cigsta3 =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }
                                 \DataTypeTok{nssec8 =} \OperatorTok{-}\DecValTok{9}\NormalTok{,  }\DataTypeTok{paidemployment =} \OperatorTok{-}\DecValTok{1}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{replace_with_na}\NormalTok{(}\DataTypeTok{replace =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{hyper140_2 =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{hibp140_2 =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }
                                 \DataTypeTok{nssec8 =} \DecValTok{99}\NormalTok{, }\DataTypeTok{qual7 =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{drink =} \OperatorTok{-}\DecValTok{8}\NormalTok{, }
                                 \DataTypeTok{ethgrp2 =} \OperatorTok{-}\DecValTok{9}\NormalTok{, }\DataTypeTok{ethgrp5 =} \OperatorTok{-}\DecValTok{9}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{replace_with_na}\NormalTok{(}\DataTypeTok{replace =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{drink =} \OperatorTok{-}\DecValTok{9}\NormalTok{, }\DataTypeTok{ethgrp2 =} \OperatorTok{-}\DecValTok{8}\NormalTok{, }\DataTypeTok{ethgrp5 =} \OperatorTok{-}\DecValTok{8}\NormalTok{,}
                                 \DataTypeTok{cigsta3 =} \OperatorTok{-}\DecValTok{8}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{replace_with_na}\NormalTok{(}\DataTypeTok{replace =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{ethgrp2 =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{ethgrp5 =} \OperatorTok{-}\DecValTok{1}\NormalTok{))}

\NormalTok{BMI56 <-}\StringTok{ }\NormalTok{blood56 }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(ID, Sex, age, area, bmival, wstval, Diabetes, bpmedc2, bpmedd2, }
\NormalTok{         hyper140_}\DecValTok{2}\NormalTok{, hibp140_}\DecValTok{2}\NormalTok{, MVPAtime, Glucose, A1C, cigsta3, dnoft3, }
\NormalTok{         dnnow, wti_Y56, wtn_Y56, wtb_Y56, ethgrp5, ethgrp2, qual7,}
\NormalTok{         cluster1, cluster2, cluster3, nssec8, paidemployment, eqvinc, }
\NormalTok{         MarSt2, cluster4, cluster5, area, gor, LDL, HDL, Chol, Trig) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{Years =} \StringTok{"5-6"}\NormalTok{, }\DataTypeTok{MarStat =} \OtherTok{NA}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{wti =}\NormalTok{ wti_Y56, }\DataTypeTok{wtn =}\NormalTok{ wtn_Y56, }\DataTypeTok{wtb =}\NormalTok{ wtb_Y56, }\DataTypeTok{drink =}\NormalTok{ dnoft3) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{replace_with_na}\NormalTok{(}\DataTypeTok{replace =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{bmival =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{paidemployment =} \OperatorTok{-}\DecValTok{9}\NormalTok{,}
                                 \DataTypeTok{wstval =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{qual7 =} \OperatorTok{-}\DecValTok{8}\NormalTok{,}
                                 \DataTypeTok{bpmedd2 =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{eqvinc =} \OperatorTok{-}\DecValTok{1}\NormalTok{,}
                                 \DataTypeTok{bpmedc2 =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{MarSt2 =} \OperatorTok{-}\DecValTok{1}\NormalTok{,}
                                 \DataTypeTok{hyper140_2 =} \OperatorTok{-}\DecValTok{7}\NormalTok{, }\DataTypeTok{hibp140_2 =} \OperatorTok{-}\DecValTok{7}\NormalTok{, }
                                 \DataTypeTok{Glucose =} \OperatorTok{-}\DecValTok{1}\NormalTok{,  }\DataTypeTok{A1C  =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }
                                 \DataTypeTok{dnnow =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{drink =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{ethgrp5 =} \OperatorTok{-}\DecValTok{4}\NormalTok{,}
                                 \DataTypeTok{ethgrp2 =} \OperatorTok{-}\DecValTok{4}\NormalTok{, }\DataTypeTok{LDL =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{HDL =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }
                                 \DataTypeTok{Chol =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{Trig =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{cigsta3 =} \OperatorTok{-}\DecValTok{1}\NormalTok{,}
                                 \DataTypeTok{MVPAtime =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{nssec8 =} \OperatorTok{-}\DecValTok{9}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{replace_with_na}\NormalTok{(}\DataTypeTok{replace =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{hyper140_2 =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{hibp140_2 =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }
                                 \DataTypeTok{paidemployment =} \OperatorTok{-}\DecValTok{8}\NormalTok{, }
                                 \DataTypeTok{drink =} \OperatorTok{-}\DecValTok{8}\NormalTok{, }\DataTypeTok{nssec8 =} \DecValTok{99}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{    }\KeywordTok{replace_with_na}\NormalTok{(}\DataTypeTok{replace =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{drink =} \OperatorTok{-}\DecValTok{9}\NormalTok{, }\DataTypeTok{paidemployment =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }
                                   \DataTypeTok{qual7 =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{cigsta3 =} \OperatorTok{-}\DecValTok{8}\NormalTok{))}

\NormalTok{BMI14 <-}\StringTok{ }\NormalTok{blood14 }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(ID, Sex, age, bmival, wstval, Diabetes, bpmedc, bpmedd, hyper140, }
\NormalTok{         hibp140, MVPAtime, Glucose, A1C, cigsta3, dnoft3, dnnow, }
\NormalTok{         wti_CY1234, wtn_CY1234, wtb_CY1234, ethgr5, ethgr2, cluster, }
\NormalTok{         area, gor, nssec8, paidemployment, qual7, eqvinc, MarSt2, }
\NormalTok{         MarStat, LDL, HDL, Chol, Trig) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{hyper140_2 =}\NormalTok{ hyper140, }\DataTypeTok{hibp140_2 =}\NormalTok{ hibp140, }\DataTypeTok{bpmedd2 =}\NormalTok{ bpmedd, }
         \DataTypeTok{bpmedc2 =}\NormalTok{ bpmedc, }\DataTypeTok{cluster1 =}\NormalTok{ cluster, }\DataTypeTok{ethgrp5 =}\NormalTok{ ethgr5, }
         \DataTypeTok{ethgrp2 =}\NormalTok{ ethgr2, }\DataTypeTok{wti =}\NormalTok{ wti_CY1234, }\DataTypeTok{wtn =}\NormalTok{ wtn_CY1234, }
         \DataTypeTok{wtb =}\NormalTok{  wtb_CY1234, }\DataTypeTok{drink =}\NormalTok{ dnoft3) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{cluster2 =} \OtherTok{NA}\NormalTok{, }\DataTypeTok{cluster3 =} \OtherTok{NA}\NormalTok{, }\DataTypeTok{cluster4 =} \OtherTok{NA}\NormalTok{, }\DataTypeTok{cluster5 =} \OtherTok{NA}\NormalTok{, }
         \DataTypeTok{Years =} \StringTok{"1-4"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{replace_with_na}\NormalTok{(}\DataTypeTok{replace =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{bmival =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{paidemployment =} \OperatorTok{-}\DecValTok{9}\NormalTok{, }
                                 \DataTypeTok{wstval =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{qual7 =} \OperatorTok{-}\DecValTok{8}\NormalTok{,}
                                 \DataTypeTok{bpmedd2 =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{eqvinc =} \OperatorTok{-}\DecValTok{1}\NormalTok{,}
                                 \DataTypeTok{bpmedc2 =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{MarSt2 =} \OperatorTok{-}\DecValTok{4}\NormalTok{, }
                                 \DataTypeTok{hyper140_2 =} \OperatorTok{-}\DecValTok{7}\NormalTok{, }\DataTypeTok{MarStat =} \OperatorTok{-}\DecValTok{4}\NormalTok{,}
                                 \DataTypeTok{hibp140_2 =} \OperatorTok{-}\DecValTok{7}\NormalTok{,  }\DataTypeTok{Glucose =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }
                                 \DataTypeTok{A1C  =} \OperatorTok{-}\DecValTok{1}\NormalTok{,  }\DataTypeTok{dnnow =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{LDL =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }
                                 \DataTypeTok{HDL =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{Chol =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{Trig =} \OperatorTok{-}\DecValTok{1}\NormalTok{,}
                                 \DataTypeTok{drink =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{ethgrp5 =} \OperatorTok{-}\DecValTok{4}\NormalTok{, }\DataTypeTok{ethgrp2 =} \OperatorTok{-}\DecValTok{4}\NormalTok{,}
                                 \DataTypeTok{cigsta3 =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{MVPAtime =} \OperatorTok{-}\DecValTok{4}\NormalTok{,  }
                                 \DataTypeTok{nssec8 =} \OperatorTok{-}\DecValTok{8}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{replace_with_na}\NormalTok{(}\DataTypeTok{replace =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{hyper140_2 =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{hibp140_2 =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }
                                 \DataTypeTok{MVPAtime =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{paidemployment =} \OperatorTok{-}\DecValTok{8}\NormalTok{,}
                                 \DataTypeTok{drink =} \OperatorTok{-}\DecValTok{8}\NormalTok{, }\DataTypeTok{nssec8 =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }\DataTypeTok{qual7 =} \OperatorTok{-}\DecValTok{1}\NormalTok{, }
                                 \DataTypeTok{MarSt2 =} \OperatorTok{-}\DecValTok{1}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{replace_with_na}\NormalTok{(}\DataTypeTok{replace =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{drink =} \OperatorTok{-}\DecValTok{9}\NormalTok{, }\DataTypeTok{paidemployment =} \OperatorTok{-}\DecValTok{4}\NormalTok{,}
                                 \DataTypeTok{cigsta3 =} \OperatorTok{-}\DecValTok{8}\NormalTok{, }\DataTypeTok{nssec8 =} \DecValTok{99}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{replace_with_na}\NormalTok{(}\DataTypeTok{replace =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{paidemployment =} \OperatorTok{-}\DecValTok{1}\NormalTok{))}

\NormalTok{BMI <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(BMI14, BMI56, BMI78)}

\NormalTok{Energy14 <-}\StringTok{ }\NormalTok{food14 }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(ID, Country, SurveyYear, EnergykJ, Carbohydrateg, }
\NormalTok{         CHOpctotE, Proteing, ProteinpctotE, Alcoholg, AlcoholpctotE,}
\NormalTok{         Fatg, FatpctotE)}
\NormalTok{Energy14}\OperatorTok{$}\NormalTok{Country[Energy14}\OperatorTok{$}\NormalTok{Country }\OperatorTok{==}\StringTok{ "Northern Ireland"}\NormalTok{] <-}\StringTok{ "NI"}

\NormalTok{Energy56 <-}\StringTok{ }\NormalTok{food56 }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(ID, Country, Surveyyear, EnergykJ, Carbohydrateg, }
\NormalTok{         CHOpctotE, Proteing, ProteinpctotE, Alcoholg, AlcoholpctotE,}
\NormalTok{         Fatg, FatpctotE) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{SurveyYear =}\NormalTok{ Surveyyear)}

\NormalTok{Energy78 <-}\StringTok{ }\NormalTok{food78 }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(ID, Country, SurveyYear, EnergykJ, Carbohydrateg, }
\NormalTok{         CHOpctotE, Proteing, ProteinpctotE, Alcoholg, AlcoholpctotE,}
\NormalTok{         Fatg, FatpctotE)}

\NormalTok{Energy <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(Energy14, Energy56, Energy78)}

\NormalTok{dta_NDNS <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"dta_NDNS_Tslots.csv"}\NormalTok{) }\CommentTok{# extract the day of week data}

\NormalTok{CW3idday}\OperatorTok{$}\NormalTok{DayNo <-}\StringTok{ }\KeywordTok{ave}\NormalTok{(CW3idday}\OperatorTok{$}\NormalTok{ID_DAY, CW3idday}\OperatorTok{$}\NormalTok{ID, }
                      \DataTypeTok{FUN =}\NormalTok{ seq_along) }\CommentTok{# adding the day no }

\NormalTok{dta_DayofWeek <-}\StringTok{ }\NormalTok{dta_NDNS }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(id, id_dy, DayofWeek)}

\NormalTok{dta_DayofWeek}\OperatorTok{$}\NormalTok{DayNo <-}\StringTok{ }\KeywordTok{unlist}\NormalTok{(}\KeywordTok{strsplit}\NormalTok{(dta_DayofWeek}\OperatorTok{$}\NormalTok{id_dy,}
                                       \StringTok{"D"}\NormalTok{))[}\KeywordTok{c}\NormalTok{(}\OtherTok{FALSE}\NormalTok{, }\OtherTok{TRUE}\NormalTok{)] }\CommentTok{# creating a day No}
\KeywordTok{names}\NormalTok{(dta_DayofWeek)[}\DecValTok{1}\NormalTok{] <-}\StringTok{ "ID"}
\NormalTok{dta_DayofWeek}\OperatorTok{$}\NormalTok{DayNo <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{(dta_DayofWeek}\OperatorTok{$}\NormalTok{DayNo)}

\NormalTok{CW3idday <-}\StringTok{   }\NormalTok{CW3idday }\OperatorTok{%>%}\StringTok{  }
\StringTok{  }\KeywordTok{left_join}\NormalTok{(dta_DayofWeek, }\DataTypeTok{by=} \KeywordTok{c}\NormalTok{(}\StringTok{"ID"}\NormalTok{, }\StringTok{"DayNo"}\NormalTok{))}

\NormalTok{## Manually check the day of week}
\NormalTok{CW3idday}\OperatorTok{$}\NormalTok{DayofWeek[CW3idday}\OperatorTok{$}\NormalTok{ID_DAY }\OperatorTok{==}\StringTok{ }\DecValTok{40714261000}\NormalTok{] <-}\StringTok{ "Saturday"}
\NormalTok{CW3idday}\OperatorTok{$}\NormalTok{DayofWeek[CW3idday}\OperatorTok{$}\NormalTok{ID_DAY }\OperatorTok{==}\StringTok{ }\DecValTok{112050710000}\NormalTok{] <-}\StringTok{ "Tuesday"}
\NormalTok{CW3idday}\OperatorTok{$}\NormalTok{DayofWeek[CW3idday}\OperatorTok{$}\NormalTok{ID_DAY }\OperatorTok{==}\StringTok{ }\DecValTok{310122510000}\NormalTok{] <-}\StringTok{ "Wednesday"}
\NormalTok{CW3idday}\OperatorTok{$}\NormalTok{DayofWeek[CW3idday}\OperatorTok{$}\NormalTok{ID_DAY }\OperatorTok{==}\StringTok{ }\DecValTok{5050616100}\NormalTok{] <-}\StringTok{ "Monday"}
\NormalTok{CW3idday}\OperatorTok{$}\NormalTok{DayofWeek[CW3idday}\OperatorTok{$}\NormalTok{ID_DAY }\OperatorTok{==}\StringTok{ }\DecValTok{50506161000}\NormalTok{] <-}\StringTok{ "Tuesday"}
\NormalTok{CW3idday}\OperatorTok{$}\NormalTok{DayofWeek[CW3idday}\OperatorTok{$}\NormalTok{ID_DAY }\OperatorTok{==}\StringTok{ }\DecValTok{505061610000}\NormalTok{] <-}\StringTok{ "Wednesday"}
\NormalTok{CW3idday}\OperatorTok{$}\NormalTok{DayofWeek[CW3idday}\OperatorTok{$}\NormalTok{ID_DAY }\OperatorTok{==}\StringTok{ }\DecValTok{7090824100}\NormalTok{] <-}\StringTok{ "Monday"}
\NormalTok{CW3idday}\OperatorTok{$}\NormalTok{DayofWeek[CW3idday}\OperatorTok{$}\NormalTok{ID_DAY }\OperatorTok{==}\StringTok{ }\DecValTok{70908241000}\NormalTok{] <-}\StringTok{ "Tuesday"}
\NormalTok{CW3idday}\OperatorTok{$}\NormalTok{DayofWeek[CW3idday}\OperatorTok{$}\NormalTok{ID_DAY }\OperatorTok{==}\StringTok{ }\DecValTok{709082410000}\NormalTok{] <-}\StringTok{ "Wednesday"}
\NormalTok{CW3idday}\OperatorTok{$}\NormalTok{DayofWeek[CW3idday}\OperatorTok{$}\NormalTok{ID_DAY }\OperatorTok{==}\StringTok{ }\DecValTok{807021910000}\NormalTok{] <-}\StringTok{ "Monday"}
\NormalTok{CW3idday}\OperatorTok{$}\NormalTok{DayofWeek[CW3idday}\OperatorTok{$}\NormalTok{ID_DAY }\OperatorTok{==}\StringTok{ }\DecValTok{812101310000}\NormalTok{] <-}\StringTok{ "Sunday"}

\NormalTok{CW3CB3_reg <-}\StringTok{ }\NormalTok{CW3CB3[}\OperatorTok{!}\KeywordTok{duplicated}\NormalTok{(CW3CB3}\OperatorTok{$}\NormalTok{ID), ]}

\CommentTok{# extract only the CB variable (Between individual classes == 1 or 2)}
\NormalTok{CW3CB3_reg <-}\StringTok{ }\NormalTok{CW3CB3_reg }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(ID, AGE, SEX, CB) }

\KeywordTok{tab1}\NormalTok{(CW3CB3_reg}\OperatorTok{$}\NormalTok{CB, }\DataTypeTok{graph =} \OtherTok{FALSE}\NormalTok{)}

\NormalTok{## dataset for 3by3 multilevel latent classes }
\NormalTok{CW3CB3_regss <-}\StringTok{ }\NormalTok{CW3CB3_reg }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{left_join}\NormalTok{(BMI, }\DataTypeTok{by =} \StringTok{"ID"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{left_join}\NormalTok{(Energy, }\DataTypeTok{by =} \StringTok{"ID"}\NormalTok{) }

\CommentTok{# Rescale the weighting values}
\CommentTok{# individual weighting}

\NormalTok{a <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "1-4"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wti)}
\NormalTok{b <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "5-6"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wti)}
\NormalTok{c <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "7-8"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wti)}

\NormalTok{CW3CB3_regss}\OperatorTok{$}\NormalTok{wti1to8 <-}\StringTok{ }\NormalTok{CW3CB3_regss}\OperatorTok{$}\NormalTok{wti}

\NormalTok{CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "1-4"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wti1to8 <-}\StringTok{ }
\StringTok{  }\NormalTok{CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "1-4"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wti}\OperatorTok{*}\NormalTok{(a}\OperatorTok{+}\NormalTok{b}\OperatorTok{+}\NormalTok{c)}\OperatorTok{*}\NormalTok{(}\DecValTok{1}\OperatorTok{/}\DecValTok{2}\NormalTok{)}\OperatorTok{/}\NormalTok{a}
\NormalTok{CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "5-6"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wti1to8 <-}\StringTok{ }
\StringTok{  }\NormalTok{CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "5-6"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wti}\OperatorTok{*}\NormalTok{(a}\OperatorTok{+}\NormalTok{b}\OperatorTok{+}\NormalTok{c)}\OperatorTok{*}\NormalTok{(}\DecValTok{1}\OperatorTok{/}\DecValTok{4}\NormalTok{)}\OperatorTok{/}\NormalTok{b}
\NormalTok{CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "7-8"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wti1to8 <-}\StringTok{ }
\StringTok{  }\NormalTok{CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "7-8"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wti}\OperatorTok{*}\NormalTok{(a}\OperatorTok{+}\NormalTok{b}\OperatorTok{+}\NormalTok{c)}\OperatorTok{*}\NormalTok{(}\DecValTok{1}\OperatorTok{/}\DecValTok{4}\NormalTok{)}\OperatorTok{/}\NormalTok{c}
\KeywordTok{mean}\NormalTok{(CW3CB3_regss}\OperatorTok{$}\NormalTok{wti1to8)}

\NormalTok{CW3CB3_regss}\OperatorTok{$}\NormalTok{wti1to8 <-}\StringTok{ }\NormalTok{CW3CB3_regss}\OperatorTok{$}\NormalTok{wti1to8}\OperatorTok{/}\FloatTok{1.209816814}
\KeywordTok{summ}\NormalTok{(CW3CB3_regss}\OperatorTok{$}\NormalTok{wti1to8, }\DataTypeTok{graph =} \OtherTok{FALSE}\NormalTok{)}
\CommentTok{#Check if the weighting sum up to the sample size we have}
\KeywordTok{sum}\NormalTok{(CW3CB3_regss}\OperatorTok{$}\NormalTok{wti1to8, }\DataTypeTok{graph =} \OtherTok{FALSE}\NormalTok{) }

\CommentTok{# Nurse weights}

\NormalTok{a <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "1-4"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wtn)}
\NormalTok{b <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "5-6"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wtn)}
\NormalTok{c <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "7-8"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wtn)}

\NormalTok{CW3CB3_regss}\OperatorTok{$}\NormalTok{wtn1to8 <-}\StringTok{ }\NormalTok{CW3CB3_regss}\OperatorTok{$}\NormalTok{wtn}

\NormalTok{CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "1-4"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wtn1to8 <-}\StringTok{ }
\StringTok{  }\NormalTok{CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "1-4"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wtn}\OperatorTok{*}\NormalTok{(a}\OperatorTok{+}\NormalTok{b}\OperatorTok{+}\NormalTok{c)}\OperatorTok{*}\NormalTok{(}\DecValTok{1}\OperatorTok{/}\DecValTok{2}\NormalTok{)}\OperatorTok{/}\NormalTok{a}
\NormalTok{CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "5-6"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wtn1to8 <-}\StringTok{ }
\StringTok{  }\NormalTok{CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "5-6"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wtn}\OperatorTok{*}\NormalTok{(a}\OperatorTok{+}\NormalTok{b}\OperatorTok{+}\NormalTok{c)}\OperatorTok{*}\NormalTok{(}\DecValTok{1}\OperatorTok{/}\DecValTok{4}\NormalTok{)}\OperatorTok{/}\NormalTok{b}
\NormalTok{CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "7-8"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wtn1to8 <-}\StringTok{ }
\StringTok{  }\NormalTok{CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "7-8"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wtn}\OperatorTok{*}\NormalTok{(a}\OperatorTok{+}\NormalTok{b}\OperatorTok{+}\NormalTok{c)}\OperatorTok{*}\NormalTok{(}\DecValTok{1}\OperatorTok{/}\DecValTok{4}\NormalTok{)}\OperatorTok{/}\NormalTok{c}
\KeywordTok{mean}\NormalTok{(CW3CB3_regss}\OperatorTok{$}\NormalTok{wtn1to8)}

\NormalTok{CW3CB3_regss}\OperatorTok{$}\NormalTok{wtn1to8 <-}\StringTok{ }\NormalTok{CW3CB3_regss}\OperatorTok{$}\NormalTok{wtn1to8}\OperatorTok{/}\FloatTok{0.907003577}
\KeywordTok{summ}\NormalTok{(CW3CB3_regss}\OperatorTok{$}\NormalTok{wtn1to8, }\DataTypeTok{graph =} \OtherTok{FALSE}\NormalTok{)}
\CommentTok{#Check if the weighting sum up to the sample size we have}
\KeywordTok{sum}\NormalTok{(CW3CB3_regss}\OperatorTok{$}\NormalTok{wtn1to8, }\DataTypeTok{graph =} \OtherTok{FALSE}\NormalTok{) }

\CommentTok{# Blood weights}
\NormalTok{a <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "1-4"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wtb)}
\NormalTok{b <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "5-6"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wtb)}
\NormalTok{c <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "7-8"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wtb)}

\NormalTok{CW3CB3_regss}\OperatorTok{$}\NormalTok{wtb1to8 <-}\StringTok{ }\NormalTok{CW3CB3_regss}\OperatorTok{$}\NormalTok{wtb}

\NormalTok{CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "1-4"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wtb1to8 <-}\StringTok{ }
\StringTok{  }\NormalTok{CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "1-4"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wtb}\OperatorTok{*}\NormalTok{(a}\OperatorTok{+}\NormalTok{b}\OperatorTok{+}\NormalTok{c)}\OperatorTok{*}\NormalTok{(}\DecValTok{1}\OperatorTok{/}\DecValTok{2}\NormalTok{)}\OperatorTok{/}\NormalTok{a}
\NormalTok{CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "5-6"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wtb1to8 <-}\StringTok{ }
\StringTok{  }\NormalTok{CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "5-6"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wtb}\OperatorTok{*}\NormalTok{(a}\OperatorTok{+}\NormalTok{b}\OperatorTok{+}\NormalTok{c)}\OperatorTok{*}\NormalTok{(}\DecValTok{1}\OperatorTok{/}\DecValTok{4}\NormalTok{)}\OperatorTok{/}\NormalTok{b}
\NormalTok{CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "7-8"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wtb1to8 <-}\StringTok{ }
\StringTok{  }\NormalTok{CW3CB3_regss[CW3CB3_regss}\OperatorTok{$}\NormalTok{Years }\OperatorTok{==}\StringTok{ "7-8"}\NormalTok{,]}\OperatorTok{$}\NormalTok{wtb}\OperatorTok{*}\NormalTok{(a}\OperatorTok{+}\NormalTok{b}\OperatorTok{+}\NormalTok{c)}\OperatorTok{*}\NormalTok{(}\DecValTok{1}\OperatorTok{/}\DecValTok{4}\NormalTok{)}\OperatorTok{/}\NormalTok{c}
\KeywordTok{mean}\NormalTok{(CW3CB3_regss}\OperatorTok{$}\NormalTok{wtb1to8)}

\NormalTok{CW3CB3_regss}\OperatorTok{$}\NormalTok{wtb1to8 <-}\StringTok{ }\NormalTok{CW3CB3_regss}\OperatorTok{$}\NormalTok{wtb1to8}\OperatorTok{/}\FloatTok{0.4817444505}
\KeywordTok{summ}\NormalTok{(CW3CB3_regss}\OperatorTok{$}\NormalTok{wtb1to8, }\DataTypeTok{graph =} \OtherTok{FALSE}\NormalTok{)}
\CommentTok{#Check if the weighting sum up to the sample size we have}
\KeywordTok{sum}\NormalTok{(CW3CB3_regss}\OperatorTok{$}\NormalTok{wtb1to8, }\DataTypeTok{graph =} \OtherTok{FALSE}\NormalTok{)}

\NormalTok{weightings <-}\StringTok{ }\NormalTok{CW3CB3_regss }\OperatorTok{%>%}\StringTok{ }\KeywordTok{select}\NormalTok{(ID, wti1to8, wtn1to8, wtb1to8)}

\CommentTok{# Combine the data from nutrient intake at each time slot}
\NormalTok{CW3CB3_7regss <-}\StringTok{ }\NormalTok{CW3CB3_7regss }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{left_join}\NormalTok{(IntakeSlots, }\DataTypeTok{by =} \StringTok{"ID"}\NormalTok{)}

\NormalTok{## dta ready to be analysed in STATA}
 \CommentTok{# change the path accordingly}
\KeywordTok{write_dta}\NormalTok{(CW3CB3_regss, }\StringTok{"../CW3CB3_7regss.dta"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
************************************************
// Analysing NDNS survey data in stata
// for CW3CB3 survey data analysis 
// date created: 2018-08-01
// manipulation of the data was done in R
// import data from CW3CB3_7sregss.dta
// change the path accordingly
************************************************

use "../CW3CB3_7regss.dta", clear

label define smoking 1 "current" 2 "ex-smoker" 3 "Never"
label values cigsta3 smoking
label define gender 1 "Men" 2 "Women"
label values Sex gender
label define paid 1 "No" 2 "Yes"
label values paidemployment paid
label define ethnicity5 1 "White" 2 "Mixed" 3 "Black" 4 "Asian" 5 "Other"
label values ethgrp5 ethnicity5
label define ethnicity2 1 "White" 2 "non-White"
label values ethgrp2 ethnicity2

gen Married = 1 if MarStat == 2 | MarSt2 == 2 
replace Married = 1 if MarSt2 == 3
replace Married = 0 if Married !=1
tab Married
tab MarSt2
tab MarStat

label define Partner 0 "No" 1 "Yes"
label values Married Partner

gen Education = qual7 == 1
label define Ed 0 "lower than Degree" 1 "Degree or higher"
label values Education Ed

replace Education = . if qual7 >100
tab Educ

egen BMIcat = cut(bmival), at(10, 25, 30, 40, 60)
tab BMIcat

**********************************************************
// variables need to be log transfomred                 //
**********************************************************

gen logalc = ln(Alcoholg+1)
summ logalc, detail
gen logMVP = ln(MVPAtime+1)
summ logMVP, detail
gen logGlu = ln(Glucose)
summ logGlu, detail
gen logA1C = ln(A1C)
summ logA1C, detail
gen logChol = ln(Chol)
summ logChol, detail
gen logLDL = ln(LDL)
gen logHDL = ln(HDL)
gen logTG = ln(Trig)

*********************************************************
//  weighting use wti to see the individual results    //
//                                                     //
*********************************************************

// weighting with individual weights, area is primary sampling unit, 
// gor is the cluster variable

svyset area [pweight = wti1to8], strata(gor)

svydescribe wti // describe the weighted data set

svy: tabulate Sex CB, row se ci format(%7.3f)
svy: tabulate Sex CB, col se ci format(%7.3f)
svy: tabulate Country CB, col se ci format(%7.3f)
svy: tabulate Country CB, row se ci format(%7.3f)
svy: tabulate SurveyYear CB, col se ci format(%7.3f)
svy: tabulate SurveyYear CB, row se ci format(%7.3f)
svy: tabulate paid CB, col se ci format(%7.3f)

svy: tabulate MarSt2 CB
svy: tabulate MarStat CB
svy: tabulate Married CB, row se ci format(%7.3f)
svy: tabulate Married CB, col se ci format(%7.3f)
svy: mean eqvinc, over(CB)
test [eqvinc]1 = [eqvinc]2 = [eqvinc]3, mtest(b) 
// bonferroni-adjusted p-values for multiple groups using the mtest(b) option

svy: tabulate ethgrp2 CB, row se ci format(%7.3f)
svy: tabulate ethgrp2 CB, col se ci format(%7.3f)
svy: tabulate Education CB, row se ci format(%7.3f)
svy: tabulate Education CB, col se ci format(%7.3f)

*********************************************************
//  nutritional distribution                           //
//                                                     //
*********************************************************

svy: mean EnergykJ, over(CB)
test [EnergykJ]1 = [EnergykJ]2 = [EnergykJ]3, mtest(b)

svy: mean Energy6, over(CB)
test [Energy6_9]1 = [Energy6_9]2 = [Energy6_9]3, mtest(b)

svy: mean Energy9, over(CB) 
test [Energy9_12]1 = [Energy9_12]2 = [Energy9_12]3, mtest(b)

svy: mean Energy12, over(CB)
test [Energy12_14]1 = [Energy12_14]2 = [Energy12_14]3, mtest(b)

svy: mean Energy14, over(CB)
test [Energy14_17]1 = [Energy14_17]2 = [Energy14_17]3, mtest(b)

svy: mean Energy17, over(CB)
test [Energy17_20]1 = [Energy17_20]2 = [Energy17_20]3, mtest(b)

svy: mean Energy20, over(CB)
test [Energy20_22]1 = [Energy20_22]2 = [Energy20_22]3, mtest(b)

svy: mean Energy22, over(CB)
test [Energy22_6]1 = [Energy22_6]2 = [Energy22_6]3, mtest(b)

svy: mean Carbohydrateg, over(CB)
test [Carbohydrateg]1 = [Carbohydrateg]2 = [Carbohydrateg]3, mtest(b)

svy: mean Carb6, over(CB)
test [Carb6_9]1 = [Carb6_9]2 = [Carb6_9]3, mtest(b)

svy: mean Carb9, over(CB)
test [Carb9_12]1 = [Carb9_12]2 = [Carb9_12]3, mtest(b)

svy: mean Carb12, over(CB)
test [Carb12_14]1 = [Carb12_14]2 = [Carb12_14]3, mtest(b)

svy: mean Carb14, over(CB)
test [Carb14_17]1 = [Carb14_17]2 = [Carb14_17]3, mtest(b)

svy: mean Carb17, over(CB)
test [Carb17_20]1 = [Carb17_20]2 = [Carb17_20]3, mtest(b)

svy: mean Carb20, over(CB)
test [Carb20_22]1 = [Carb20_22]2 = [Carb20_22]3, mtest(b)

svy: mean Carb22, over(CB)
test [Carb22_6]1 = [Carb22_6]2 = [Carb22_6]3, mtest(b)

svy: mean Sugar6, over(CB)
test [Sugar6_9]1 = [Sugar6_9]2 = [Sugar6_9]3, mtest(b)

svy: mean Sugar9, over(CB)
test [Sugar9_12]1 = [Sugar9_12]2 = [Sugar9_12]3, mtest(b)

svy: mean Sugar12, over(CB)
test [Sugar9_12]1 = [Sugar9_12]2 = [Sugar9_12]3, mtest(b)

svy: mean Sugar14, over(CB)
test [Sugar14_17]1 = [Sugar14_17]2 = [Sugar14_17]3, mtest(b)

svy: mean Sugar17, over(CB)
test [Sugar17_20]1 = [Sugar17_20]2 = [Sugar17_20]3, mtest(b)

svy: mean Sugar20, over(CB)
test [Sugar20_22]1 = [Sugar20_22]2 = [Sugar20_22]3, mtest(b)

svy: mean Sugar22, over(CB)
test [Sugar22_6]1 = [Sugar22_6]2 = [Sugar22_6]3, mtest(b)

svy: mean Starch6, over(CB)
test [Starch6_9]1 = [Starch6_9]2 = [Starch6_9]3, mtest(b)

svy: mean Starch9, over(CB)
test [Sugar12_14]1 = [Sugar12_14]2 = [Sugar12_14]3, mtest(b)

svy: mean Starch12, over(CB)
test [Starch12_14]1 = [Starch12_14]2 = [Starch12_14]3, mtest(b)

svy: mean Starch14, over(CB)
test [Starch14_17]1 = [Starch14_17]2 = [Starch14_17]3, mtest(b)

svy: mean Starch17, over(CB)
test [Starch17_20]1 = [Starch17_20]2 = [Starch17_20]3, mtest(b)

svy: mean Starch20, over(CB)
test [Starch20_22]1 = [Starch20_22]2 = [Starch20_22]3, mtest(b)

svy: mean Starch22, over(CB)
test [Starch20_22]1 = [Starch20_22]2 = [Starch20_22]3, mtest(b)

svy: mean Fibre6, over(CB)
test [Starch22_6]1 = [Starch22_6]2 = [Starch22_6]3, mtest(b)

gen Fibreg = Fibre6 + Fibre9 + Fibre12 + Fibre14 + Fibre17 + Fibre20 + Fibre22
svy: mean Fibreg, over(CB)
test [Fibreg]1 = [Fibreg]2 = [Fibreg]3, mtest(b)

svy: mean Fibre9, over(CB)
test [Fibre9_12]1 = [Fibre9_12]2 = [Fibre9_12]3, mtest(b)

svy: mean Fibre12, over(CB)
test [Fibre12_14]1 = [Fibre12_14]2 = [Fibre12_14]3, mtest(b)

svy: mean Fibre14, over(CB)
test [Fibre14_17]1 = [Fibre14_17]2 = [Fibre14_17]3, mtest(b)

svy: mean Fibre17, over(CB)
test [Fibre17_20]1 = [Fibre17_20]2 = [Fibre17_20]3, mtest(b)

svy: mean Fibre20, over(CB)
test [Fibre20_22]1 = [Fibre20_22]2 = [Fibre20_22]3, mtest(b)

svy: mean Fibre22, over(CB)
test [Fibre22_6]1 = [Fibre22_6]2 = [Fibre22_6]3, mtest(b)

svy: mean NMES6, over(CB)
test [NMES6_9]1 = [NMES6_9]2 = [NMES6_9]3, mtest(b)

svy: mean NMES9, over(CB)
test [NMES9_12]1 = [NMES9_12]2 = [NMES9_12]3, mtest(b)

svy: mean NMES12, over(CB)
test [NMES12_14]1 = [NMES12_14]2 = [NMES12_14]3, mtest(b)

svy: mean NMES14, over(CB)
test [NMES14_17]1 = [NMES14_17]2 = [NMES14_17]3, mtest(b)

svy: mean NMES17, over(CB)
test [NMES17_20]1 = [NMES17_20]2 = [NMES17_20]3, mtest(b)

svy: mean NMES20, over(CB)
test [NMES20_22]1 = [NMES20_22]2 = [NMES20_22]3, mtest(b)

svy: mean NMES22, over(CB)

svy: mean CHO, over(CB)
test [CHOpctotE]1 = [CHOpctotE]2 = [CHOpctotE]3, mtest(b)

svy: mean Proteing, over(CB)
test [Proteing]1 = [Proteing]2 = [Proteing]3, mtest(b)

svy: mean Prot6, over(CB)
test [Prot6_9]1 = [Prot6_9]2 = [Prot6_9]3, mtest(b)

svy: mean Prot9, over(CB)
test [Prot9_12]1 = [Prot9_12]2 = [Prot9_12]3, mtest(b)

svy: mean Prot12, over(CB)
test [Prot12_14]1 = [Prot12_14]2 = [Prot12_14]3, mtest(b)

svy: mean Prot14, over(CB)
test [Prot14_17]1 = [Prot14_17]2 = [Prot14_17]3, mtest(b)

svy: mean Prot17, over(CB)
test [Prot17_20]1 = [Prot17_20]2 = [Prot17_20]3, mtest(b)

svy: mean Prot20, over(CB)
test [Prot20_22]1 = [Prot20_22]2 = [Prot20_22]3, mtest(b)

svy: mean Prot22, over(CB)
test [Prot22_6]1 = [Prot22_6]2 = [Prot22_6]3, mtest(b)

svy: mean Proteinp, over(CB)
test [ProteinpctotE]1 = [ProteinpctotE]2 = [ProteinpctotE]3, mtest(b)

svy: mean Fatg, over(CB)
test [Fatg]1 = [Fatg]2 = [Fatg]3, mtest(b)

svy: mean Fat6, over(CB)
test [Fat6_9]1 = [Fat6_9]2 = [Fat6_9]3, mtest(b)

svy: mean Fat9, over(CB)
test [Fat9_12]1 = [Fat9_12]2 = [Fat9_12]3, mtest(b)

svy: mean Fat12, over(CB)
test [Fat12_14]1 = [Fat12_14]2 = [Fat12_14]3, mtest(b)

svy: mean Fat14, over(CB)
test [Fat14_17]1 = [Fat14_17]2 = [Fat14_17]3, mtest(b)

svy: mean Fat17, over(CB)
test [Fat17_20]1 = [Fat17_20]2 = [Fat17_20]3, mtest(b)

svy: mean Fat20, over(CB)
test [Fat20_22]1 = [Fat20_22]2 = [Fat20_22]3, mtest(b)

svy: mean Fat22, over(CB)
test [Fat22_6]1 = [Fat22_6]2 = [Fat22_6]3, mtest(b)

svy: mean Fatp, over(CB)
test [FatpctotE]1 = [FatpctotE]2 = [FatpctotE]3, mtest(b)

svy: mean Alcoholg, over(CB)
test [Alcoholg]1 = [Alcoholg]2 = [Alcoholg]3, mtest(b)

svy: mean Alc6, over(CB)
test [Alc6_9]1 = [Alc6_9]2 = [Alc6_9]3, mtest(b)

svy: mean Alc9, over(CB)
test [Alc9_12]1 = [Alc9_12]2 = [Alc9_12]3, mtest(b)

svy: mean Alc12, over(CB)
test [Alc12_14]1 = [Alc12_14]2 = [Alc12_14]3, mtest(b)

svy: mean Alc14, over(CB)
test [Alc14_17]1 = [Alc14_17]2 = [Alc14_17]3, mtest(b)

svy: mean Alc17, over(CB)
test [Alc14_17]1 = [Alc14_17]2 = [Alc14_17]3, mtest(b)

svy: mean Alc20, over(CB)
test [Alc14_17]1 = [Alc14_17]2 = [Alc14_17]3, mtest(b)

svy: mean Alcoholp, over(CB)
test [AlcoholpctotE]1 = [AlcoholpctotE]2 = [AlcoholpctotE]3, mtest(b)

svy: tabulate cigsta3 CB, col se ci format(%7.3f)
svy: tabulate dnnow CB, col se ci format(%7.3f)
svy: tabulate hibp CB, col se ci format(%7.3f)

sum MVP [weight=wti1to8] if CB ==1 , det
sum MVP [weight=wti1to8] if CB ==2 , det
sum MVP [weight=wti1to8] if CB ==3 , det
svy: mean MVP, over(CB)

svy: mean logMVP, over(CB) eform
test [logMVP]1 = [logMVP]2 = [logMVP]3, mtest(b)

disp exp(.731059) - 1    
dis exp(.6768489) -1 
dis exp(.7852691) -1

disp exp(.6239265) - 1    
dis exp(.571165) -1 
dis exp( .6766879) -1

disp exp(.7273621) - 1    
dis exp(.684545) -1 
dis exp(.7701791) -1

svy: mean logalc, over(CB)

disp exp( 2.035795) - 1    
dis exp(1.933326) -1 
dis exp(2.138264) -1

*********************************************************
//  re-weighting use wtn to see the BMI,wc measurements //
//                                                     //
*********************************************************

svyset area [pweight = wtn1to8], strata(gor)
svy: mean wst, over(CB)
test [wstval]1 = [wstval]2 = [wstval]3, mtest(b)

gen Men = Sex == 1 
svy, subpop(Men): mean wst, over(CB)
test [wstval]1 = [wstval]2 = [wstval]3, mtest(b)

gen Women = Sex == 2
svy, subpop(Women): mean wst, over(CB)
test [wstval]1 = [wstval]2 = [wstval]3, mtest(b)

svy: mean bmi, over(CB)

test [bmival]1 = [bmival]2 = [bmival]3, mtest(b)

*********************************************************
//  re-weighting use wtb to see the blood test results //
//                                                     //
*********************************************************

svyset area [pweight = wtb1to8], strata(gor)

svy: mean HDL, over(CB)
test [HDL]1 = [HDL]2 = [HDL]3, mtest(b)

svy: mean Chol, over(CB)
test [Chol]1 = [Chol]2 = [Chol]3, mtest(b)

svy: mean LDL, over(CB)
test [LDL]1 = [LDL]2 = [LDL]3, mtest(b)

svy: mean Trig, over(CB)
test [Trig]1 = [Trig]2 = [Trig]3, mtest(b)

gen DM = A1C <= 6.5 if !missing(A1C)

svy, subpop(DM): mean Glucose, over(CB)
test [Glucose]1 = [Glucose]2
test [Glucose]1 = [Glucose]2 = [Glucose]3, mtest(b)

svy, subpop(DM): mean A1C, over(CB)
test [A1C]1 = [A1C]2
test [A1C]1 = [A1C]2 = [A1C]3, mtest(b)

svy: tabulate DM CB, col se ci format(%7.3f)

svy, subpop(DM): mean Glucose, over(CB)
test [Glucose]1 = [Glucose]2
test [Glucose]1 = [Glucose]2 = [Glucose]3, mtest(b)
svy, subpop(DM): mean A1C, over(C)
test [A1C]1 = [A1C]2
test [A1C]1 = [A1C]2 = [A1C]3, mtest(b)

svy: tabulate DM C, col se ci format(%7.3f)

svy, subpop(DM): mean logGlu, over(CB)
test [logGlu]1 = [logGlu]2 = [logGlu]3, mtest(b)

dis exp(1.642848)
dis exp(1.632226)
dis exp(1.653471)

dis exp(1.620347)
dis exp(1.606447)
dis exp(1.634246)

dis exp(1.629356)
dis exp(1.620271)
dis exp(1.63844)

svy, subpop(DM): mean logA1C, over(CB)
test [logA1C]1 = [logA1C]2 = [logA1C]3, mtest(b)

dis exp(1.699581)
dis exp(1.69296)
dis exp(1.706203)

dis exp(1.691608)
dis exp(1.683897)
dis exp(1.699318)

dis exp(1.705623)
dis exp(1.700665)
dis exp(1.710581)

svy: mean logChol, over(CB)
dis exp(1.598818)
dis exp(1.577698)
dis exp(1.619939)

dis exp(1.55251)  
dis exp(1.530408)
dis exp(1.574613)

dis exp(1.599389)
dis exp(1.583391)
dis exp(1.615388)

test [logChol]1 = [logChol]2 = [logChol]3, mtest(b)

svy: mean logHDL, over(CB)
dis exp(.3293169)
dis exp(.3026793)
dis exp(.3559545)

dis exp(.2749379)  
dis exp(.2476816)
dis exp(.3021941)

dis exp(.3269002)
dis exp(.3062623)
dis exp(.3475381)

test [logHDL]1 = [logHDL]2 = [logHDL]3, mtest(b)

svy: mean logLDL, over(CB)
dis exp(1.058635)
dis exp(1.028391)
dis exp(1.08888)

dis exp(1.018181)  
dis exp(.984431)
dis exp(1.051931)

dis exp(1.075229)
dis exp(1.051369)
dis exp(1.09909)

test [logLDL]1 = [logLDL]2 = [logLDL]3, mtest(b)

svy: mean logTG, over(CB)
dis exp(.1273876)
dis exp(.0777152)
dis exp(.17706)

dis exp(.1012169)  
dis exp(.0460972)
dis exp(.1563366)

dis exp(.0983298)
dis exp(.0607423)
dis exp(.1359172)

test [logTG]1 = [logTG]2 = [logTG]3, mtest(b)

***************************************************
// Analysing NDNS survey data in stata
// for CW3CB3 survey data analysis on hypertension
// date created: 2018-08-06
// manipulation of the data was done in R
// import data from CW3CB3_7sregss.dta
// change the path accordingly
**************************************************

use "../CW3CB3_7regss.dta", clear

label define smoking 1 "current" 2 "ex-smoker" 3 "Never"
label values cigsta3 smoking
label define gender 1 "Men" 2 "Women"
label values Sex gender
label define paid 1 "No" 2 "Yes"
label values paidemployment paid
label define ethnicity5 1 "White" 2 "Mixed" 3 "Black" 4 "Asian" 5 "Other"
label values ethgrp5 ethnicity5
label define ethnicity2 1 "White" 2 "non-White"
label values ethgrp2 ethnicity2

gen Married = 1 if MarStat == 2 | MarSt2 == 2 
replace Married = 1 if MarSt2 == 3
replace Married = 0 if Married !=1
tab Married
tab MarSt2
tab MarStat

label define Partner 0 "No" 1 "Yes"
label values Married Partner

gen Education = qual7 == 1
label define Ed 0 "lower than Degree" 1 "Degree or higher"
label values Education Ed

replace Education = . if qual7 >100
tab Educ

egen BMIcat = cut(bmival), at(10, 25, 30, 40, 60)
tab BMIcat

**********************************************************
// variables need to be log transfomred                 //
**********************************************************

gen logalc = ln(Alcoholg+1)
summ logalc, detail
gen logMVP = ln(MVPAtime+1)
summ logMVP, detail
gen logGlu = ln(Glucose)
summ logGlu, detail
gen logA1C = ln(A1C)
summ logA1C, detail
gen logChol = ln(Chol)
summ logChol, detail
gen logLDL = ln(LDL)
gen logHDL = ln(HDL)
gen logTG = ln(Trig)

*********************************************************
//  weighting use wti to see the individual results    //
*********************************************************

// weighting with individual weights, area is primary sampling unit, 
// gor is the cluster variable
svyset area [pweight = wti1to8], strata(gor)

svydescribe wti // describe the weighted data set

*********************************************************
//  re-weighting use wtn to see the BMI,wc measurements //
*********************************************************
svyset area [pweight = wtn1to8], strata(gor)

gen Men = Sex == 1  //  n of men = 2537
gen Women = Sex == 2 // n of women = 3618

svy, subpop(Men): tab hibp, se ci format(%7.3f)
svy, subpop(Women): tab hibp, se ci format(%7.3f)

svy, subpop(Men): mean age, over(hibp)
test [age]1 = [age]0
svy, subpop(Women): mean age, over(hibp)
test [age]1 = [age]0

svy, subpop(Men): mean wst, over(hibp)
test [wstval]1 = [wstval]0
svy, subpop(Women): mean wst, over(hibp)
test [wstval]1 = [wstval]0

svy, subpop(Men): tabulate CB hibp, col se ci format(%7.3f)

svy, subpop(Women): tabulate CB hibp, col se ci format(%7.3f)

svy, subpop(Men): tabulate Country hibp, col se ci format(%7.3f)

svy, subpop(Women): tabulate Country hibp, col se ci format(%7.3f)

svy, subpop(Men): tabulate SurveyYear hibp, col se ci format(%7.3f)

svy, subpop(Women): tabulate SurveyYear hibp, col se ci format(%7.3f)

svy, subpop(Men): tabulate ethgrp2 hibp, col se ci format(%7.3f)

svy, subpop(Women): tabulate ethgrp2 hibp, col se ci format(%7.3f)

svy, subpop(Men): tabulate Edu hibp, col se ci format(%7.3f)

svy, subpop(Women): tabulate Edu hibp, col se ci format(%7.3f)

svy, subpop(Men): tabulate cigsta3 hibp, col se ci format(%7.3f)

svy, subpop(Women): tabulate cigsta3 hibp, col se ci format(%7.3f)

svy, subpop(Men): tabulate Married hibp, col  se ci format(%7.3f)

svy, subpop(Men): mean logMVP, over(hibp)
test [logMVP]1 = [logMVP]0

disp exp(.9234363 ) - 1    
dis exp(.8457101) -1 
dis exp(1.001163) -1

disp exp(.828635) - 1    
dis exp(.730244) -1 
dis exp( .9270261) -1

svy, subpop(Women): mean logMVP, over(hibp) 
test [logMVP]1 = [logMVP]0

disp exp(.5916676) - 1    
dis exp(.5473043) -1 
dis exp(.6360309) -1

disp exp(.4231103) - 1    
dis exp(.3536885) -1 
dis exp(.4925322) -1

svy: mean logalc, over(CB)

disp exp( 2.035795) - 1    
dis exp(1.933326) -1 
dis exp(2.138264) -1

svy, subpop(Men): mean bmi, over(hibp)
test [bmival]1 = [bmival]0
svy, subpop(Women): mean bmi, over(hibp)
test [bmival]1 = [bmival]0

svy, subpop(Men): mean EnergykJ, over(hibp)
test [EnergykJkJ]1 = [EnergykJkJ]0
svy, subpop(Women): mean EnergykJ, over(hibp)
test [EnergykJkJ]1 = [EnergykJkJ]0

svy, subpop(Men): mean Carbo, over(hibp)
test [Carbohydrateg]1 = [Carbohydrateg]0

svy, subpop(Women): mean Carbohydrateg, over(hibp)
test [Carbohydrateg]1 = [Carbohydrateg]0

svy, subpop(Men): mean Proteing, over(hibp)
test [Proteing]1 = [Proteing]0

svy, subpop(Women): mean Carbohydrateg, over(hibp)
test [Carbohydrateg]1 = [Carbohydrateg]0

svy: tabulate Sex hibp, col se ci format(%7.3f)

svy: tabulate paid hibp, col se ci format(%7.3f)

gen DM = A1C > 6.5 if !missing(A1C)

svy, subpop(Men): tabulate Married hibp, col se ci format(%7.3f)
svy, subpop(Women): tabulate Married hibp, col se ci format(%7.3f)

svy, subpop(Men): mean eqvinc, over(hibp)
test [eqvinc]1 = [eqvinc]0

svy, subpop(Women): mean eqvinc, over(hibp)
test [eqvinc]1 = [eqvinc]0

********************************************************
**   Building the GLM model 
**   date: 07/08/2018
********************************************************

svyset area [pweight = wtn1to8], strata(gor)

// crude association between CB and hypertension

svy, subpop(Men): logistic hibp i.CB

svy, subpop(Women): logistic hibp i.CB

// in non DM 
svy, subpop(Men if DM != 1): logistic hibp i.CB
svy, subpop(Women if DM != 1): logistic hibp i.CB

// looking for confounders one by one
// Age: -> confounder
svy, subpop(Men): logistic hibp i.CB age
test age
svy, subpop(Women): logistic hibp i.CB age
test age

svy, subpop(Men): logistic hibp i.CB##c.age
svy, subpop(Women): logistic hibp i.CB##c.age
test 2.CB#c.age 3.CB#c.age // no interaction

// Partner -> confounder
svy, subpop(Men): logistic hibp i.CB i.Married
test 1.Married
svy, subpop(Women): logistic hibp i.CB i.Married
test 1.Married

svy, subpop(Men): logistic hibp i.CB##i.Married
svy, subpop(Women): logistic hibp i.CB##i.Married

test 2.CB#1.Married 3.CB#1.Married // -> no interaction

// Income -> not confounder for men but confounder for women
svy, subpop(Men): logistic hibp i.CB eqvinc 
test eqvinc
svy, subpop(Women): logistic hibp i.CB eqvinc 
test eqvinc

svy, subpop(Men): logistic hibp i.CB##c.eqvinc 
svy, subpop(Women): logistic hibp i.CB##c.eqvinc

test 2.CB#c.eqvinc 3.CB#c.eqvinc // -> (probably) no interaction

// Education -> confounder
svy, subpop(Men): logistic hibp i.CB i.Edu 
test 1.Edu
svy, subpop(Women): logistic hibp i.CB i.Edu 
test 1.Edu

svy, subpop(Men): logistic hibp i.CB##i.Edu 
svy, subpop(Women): logistic hibp i.CB##i.Edu

test 2.CB#1.Edu 3.CB#1.Edu // no interaction

// BMI -> confounder
svy, subpop(Men): logistic hibp i.CB bmi
test bmi
svy, subpop(Women): logistic hibp i.CB bmi
test bmi

svy, subpop(Men): logistic hibp i.CB##c.bmi
svy, subpop(Women): logistic hibp i.CB##c.bmi

test 2.CB#c.bmival 3.CB#c.bmival // no ineraction

// paid employment -> not confounder
svy, subpop(Men): logistic hibp i.CB i.paid
test 2.paid
svy, subpop(Women): logistic hibp i.CB i.paid
test 2.paid

svy, subpop(Men): logistic hibp i.CB##i.paid
svy, subpop(Women): logistic hibp i.CB##i.paid

test 2.CB#2.paid 3.CB#2.paid

// Smoking -> confounder
svy, subpop(Men): logistic hibp i.CB i.cigsta3
test 2.cigsta3 3.cigsta3
svy, subpop(Women): logistic hibp i.CB i.cigsta3
test 2.cigsta3 3.cigsta3

svy, subpop(Men): logistic hibp i.CB##i.cigsta3
svy, subpop(Women): logistic hibp i.CB##i.cigsta3

test 2.CB#2.cigsta3 2.CB#3.cigsta3 3.CB#2.cigsta3 3.CB#3.cigsta3 // no interaction

// Total energy intake -> confounder
svy, subpop(Men): logistic hibp i.CB EnergykJ
test EnergykJ
svy, subpop(Women): logistic hibp i.CB EnergykJ
test EnergykJ

svy, subpop(Men): logistic hibp i.CB##c.EnergykJ
svy, subpop(Women): logistic hibp i.CB##c.EnergykJ
test 2.CB#c.EnergykJkJ 3.CB#c.EnergykJkJ // no interaction

// ethnicity -> not confounder
svy, subpop(Men): logistic hibp i.CB i.ethgrp2
test 2.eth
svy, subpop(Women): logistic hibp i.CB i.ethgrp2
test 2.eth

svy, subpop(Men): logistic hibp i.CB##i.ethgrp2
svy, subpop(Women): logistic hibp i.CB##i.ethgrp2
test 2.CB#2.ethgrp2 // no interaction

// Alcohol -> not confounder for men but confounder for women
svy, subpop(Men): logistic hibp i.CB Alcoholg
test Alcoholg
svy, subpop(Women): logistic hibp i.CB Alcoholg
test Alcoholg

svy, subpop(Men): logistic hibp i.CB##c.Alcoholg
svy, subpop(Women): logistic hibp i.CB##c.Alcoholg
test 2.CB#c.Alcoholg 3.CB#c.Alcoholg // no interaction

// logMVP -> not confounder 
svy, subpop(Men): logistic hibp i.CB logMVP
test logMVP
svy, subpop(Women): logistic hibp i.CB logMVP
test logMVP

svy, subpop(Men): logistic hibp i.CB##c.logMVP
svy, subpop(Women): logistic hibp i.CB##c.logMVP
test 2.CB#c.logMVP 3.CB#c.logMVP // no interaction

//  Model includes all possible confounders in Men
svy, subpop(Men): logistic hibp i.CB age i.Married i.Edu bmi i.cig EnergykJ 
linktest
svy, subpop(Men): logistic hibp i.CB age i.Married i.Edu wst i.cig EnergykJ 
linktest
svy, subpop(if Men & DM != 1): logistic hibp i.CB age i.Married ///
 i.Edu bmi i.cig EnergykJ 
linktest
svy, subpop(if Men & DM != 1): logistic hibp i.CB age i.Married ///
 i.Edu wst i.cig EnergykJ 
linktest

// use the model above to see whether any other factors are confounders 
// conditional on the other variables

// income -> not confounder
svy, subpop(Men): logistic hibp i.CB age i.Married i.Edu bmi i.cig /// 
 EnergykJ eqvinc
test eqvinc

// ethnicity -> not confounder
svy, subpop(Men): logistic hibp i.CB age i.Married i.Edu bmi i.cig /// 
 EnergykJ i.ethgrp2
test 2.ethgrp2

// alcohol -> not confounder
svy, subpop(Men): logistic hibp i.CB age i.Married i.Edu bmi i.cig /// 
 EnergykJ Alcoholg
test Alcoholg

// MVP(physical activity) -> not confounder
svy, subpop(Men): logistic hibp i.CB age i.Married i.Edu bmi i.cig /// 
 EnergykJ logMVP
test logMVP

//  Model includes all possible confounders in Women
svy, subpop(Women): logistic hibp i.CB age i.Married eqvinc i.Edu ///
 bmi i.cig EnergykJ Alcoholg 
linktest
svy, subpop(Women): logistic hibp i.CB age i.Married eqvinc i.Edu ///
 wst i.cig EnergykJ Alcoholg 
linktest

// use the model above to see whether any other factors are confounders 
// conditional on the other variables

// ethnicity -> not confounder
svy, subpop(Women): logistic hibp i.CB age i.Married eqvinc i.Edu /// 
 bmi i.cig EnergykJ Alcoholg i.ethgrp2
test 2.ethgrp2

// MVP (physical activity) -> not confounder
svy, subpop(Women): logistic hibp i.CB age i.Married eqvinc i.Edu /// 
 bmi i.cig EnergykJ Alcoholg logMVP
test logMVP

svy, subpop(if Women & DM != 1): logistic hibp i.CB age i.Married /// 
 eqvinc i.Edu bmi i.cig EnergykJ Alcoholg
linktest

svy, subpop(if Women & DM != 1): logistic hibp i.CB age i.Married /// 
 eqvinc i.Edu wst i.cig EnergykJ Alcoholg
linktest
\end{verbatim}