### 2018-08-17 paper/Appendices/AppendixE.tex

% Appendix Template

\chapter{Example of a food diary for one day} % Main appendix title

\label{AppendixE} % Change X to a consecutive letter; for referencing this appendix elsewhere, use \ref{AppendixX}
%\begin{verbatim}

\begin{figure}[H]
%\vspace*{13cm}
\centering
\includegraphics{Figures/Diary1.png}
\decoRule
\caption[One day food diary example 6 am to 2 pm]{NATIONAL DIET AND NUTRITION SURVEY -- Food and Drink Diary Example, from 6 am to 2 pm.}
\label{fig:diary1}
\end{figure}
%\end{verbatim}

\begin{figure}[H]
%\vspace*{13cm}
\centering
\includegraphics{Figures/Diary2.png}
\decoRule
\caption[One day food diary example 2 pm to 6 am]{NATIONAL DIET AND NUTRITION SURVEY -- Food and Drink Diary Example, from 2 pm to 6 am.}
\label{fig:diary2}
\end{figure}
%\end{verbatim}

\begin{figure}[H]
%\vspace*{13cm}
\centering
\includegraphics{Figures/Diary3.png}
\decoRule
\caption[Food diary example of home made food recipes]{NATIONAL DIET AND NUTRITION SURVEY -- Food and Drink Diary Example, home made food recipes.}
\label{fig:diary3}
\end{figure}