### 2018-08-20 paper/Appendices/AppendixE.tex
abstract almost done, proof reading first round done

% Appendix Template

\chapter{Example of a food diary for one day} % Main appendix title

\label{AppendixE} % Change X to a consecutive letter; for referencing this appendix elsewhere, use \ref{AppendixX}
%\begin{verbatim}

\begin{figure}[H]
%\vspace*{13cm}
\centering
\includegraphics{Figures/Diary1.png}
\decoRule
\caption[One day food diary example 6 am to 12 noon.]{NATIONAL DIET AND NUTRITION SURVEY -- Food and Drink Diary Example, from 6 am to 12 noon.}
\label{fig:diary1}
\end{figure}
%\end{verbatim}

\begin{figure}
%\vspace*{13cm}
\centering
\includegraphics{Figures/Diary2.png}
\decoRule
\caption[One day food diary example 12 noon to 5 pm]{NATIONAL DIET AND NUTRITION SURVEY -- Food and Drink Diary Example, from 12 noon to 5 pm.}
\label{fig:diary2}
\end{figure}
%\end{verbatim}

\begin{figure}
	%\vspace*{13cm}
	\centering
	\includegraphics{Figures/Diary03.png}
	\decoRule
	\caption[One day food diary example 5 pm to 6 am]{NATIONAL DIET AND NUTRITION SURVEY -- Food and Drink Diary Example, from 5 pm to 6 am.}
	\label{fig:diary2}
\end{figure}
%\end{verbatim}

\begin{figure}
%\vspace*{13cm}
\centering
\includegraphics{Figures/Diary3.png}
\decoRule
\caption[Food diary example of home made food recipes]{NATIONAL DIET AND NUTRITION SURVEY -- Food and Drink Diary Example, home made food recipes.}
\label{fig:diary3}
\end{figure}